### 2016-12-04 doc/ms.tex
Document

\documentclass[a4paper, 12pt]{article}
\usepackage{amsmath}
\usepackage{dsfont}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[left=2cm, right=2cm, bottom=3cm, top=2cm]{geometry}
\usepackage{natbib}
\usepackage{microtype}

\title{Crystals}
\author{Brendon J. Brewer}
\date{}

\begin{document}
\maketitle

%\abstract{\noindent Abstract}

% Need this after the abstract
\setlength{\parindent}{0pt}
\setlength{\parskip}{8pt}

\section{The prior information}

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}

### 2016-12-04 doc/ms.tex
Hierarchical model stuff

@@
 \usepackage{microtype}
+
+\newcommand{\hypers}{\boldsymbol{\alpha}}
+\newcommand{\params}{\boldsymbol{\theta}}
+\newcommand{\data}{\boldsymbol{x}}
+\newcommand{\info}{\boldsymbol{M}}
 
@@
 
+The joint prior distribution for the hyperparameters, parameters, and data,
+is written $p(\hypers, \params, \data | \info)$, and is usually factorised
+in this way:
+\begin{align}
+p(\hypers, \params, \data | \info) &=
+    p(\hypers | \info)p(\params | \hypers, \info)
+    p(\data | \params, \hypers, \info)\\
+    &= p(\hypers | \info)p(\params | \hypers, \info)
+    p(\data | \params, \info)
+\end{align}
+where the first step is true in general by the product rule, and the second
+step assumes that knowing the parameters would make the hyperparameters
+irrelevant when predicting what data would be observed.
+
 \bibliographystyle{plainnat}

### 2016-12-04 src/MyModel.cpp
Wider prior for asymmetry

@@ -30,7 +30,7 @@ void MyModel::from_prior(DNest4::RNG& rng)
     rc = rng.rand();
     power = 0.1 + 1.9*rng.rand();
 
-    asymmetry = exp(0.2*rng.randn());
+    asymmetry = exp(0.5*rng.randn());
     r_asymmetry = rng.rand();
 
     spikes.from_prior(rng);
@@ -98,9 +98,9 @@ double MyModel::perturb(DNest4::RNG& rng)
         else if(which == 6)
         {
             asymmetry = log(asymmetry);
-            logH -= -0.5*pow(asymmetry/0.2, 2);
-            asymmetry += 0.2*rng.randh();
-            logH += -0.5*pow(asymmetry/0.2, 2);
+            logH -= -0.5*pow(asymmetry/0.5, 2);
+            asymmetry += 0.5*rng.randh();
+            logH += -0.5*pow(asymmetry/0.5, 2);
             asymmetry = exp(asymmetry);
         }
         else

### 2016-12-04 src/Lookup.cpp
Use lookup table

@@ -1,0 +1,33 @@
+#include "Lookup.h"
+#include <cmath>
+#include <cassert>
+
+namespace Crystals
+{
+
+Lookup Lookup::instance;
+
+Lookup::Lookup()
+:num(5001)
+,xMin(0.0)
+,xMax(100.0)
+,dx((xMax - xMin)/(num-1))
+,one_over_dx(1.0/dx)
+,_exp(num)
+{
+	for(int i=0; i<num; i++)
+		_exp[i] = exp(-i*dx);
+}
+
+double Lookup::evaluate(double x)
+{
+	int i = static_cast<int>(x/instance.dx);
+    double frac = (x - i*instance.dx)*instance.one_over_dx;
+
+	if(i < 0 || i >= (static_cast<int>(instance._exp.size()) - 1))
+		return 0.0;
+	return frac*instance._exp[i+1] + (1.0 - frac)*instance._exp[i];
+}
+
+} // namespace Crystals
+

### 2016-12-04 src/Lookup.h

@@ -1,0 +1,34 @@
+#ifndef Crystals_Lookup
+#define Crystals_Lookup
+
+#include <vector>
+
+/*
+* Lookup tables for speeding things up
+* Singleton pattern
+*/
+
+namespace Crystals
+{
+
+class Lookup
+{
+	private:
+		int num;
+		double xMin, xMax, dx, one_over_dx;
+		std::vector<double> _exp; // exp(-x) for x >= 0
+
+		Lookup();
+		Lookup(const Lookup& other);
+
+		static Lookup instance;
+
+	public:
+		static double evaluate(double x);
+
+};
+
+} // namespace Crystals
+
+#endif
+

### 2016-12-04 src/MyModel.cpp

@@ -1,5 +1,6 @@
 #include "MyModel.h"
 #include "DNest4/code/DNest4.h"
+#include "Lookup.h"
 #include <sstream>
 #include <cmath>
 
@@ -197,7 +198,7 @@ void MyModel::compute_the_spikes(bool update)
         {
             rsq = pow(data_x[j] - c, 2);
             if(rsq*tau < 100.0)
-                the_spikes[j] += a*exp(-0.5*rsq*tau);
+                the_spikes[j] += a*Lookup::evaluate(0.5*rsq*tau);
         }
     }
 }

### 2016-12-04 src/MyModel.cpp
Replaced more exp calls with lookup table

@@ -153,7 +153,7 @@ void MyModel::compute_wide_component()
     for(size_t i=0; i<wide_component.size(); ++i)
     {
         rsq = pow(data_x[i] - center, 2) + pow(rc*width, 2);
-        wide_component[i] = exp(-pow(rsq*tau, power));
+        wide_component[i] = Lookup::evaluate(pow(rsq*tau, power));
     }
 
     // Multiply by a sigmoid for asymmetry
@@ -163,7 +163,7 @@ void MyModel::compute_wide_component()
     for(size_t i=0; i<wide_component.size(); ++i)
     {
         n = (data_x[i] - center)/(r_asymmetry*width);
-        wide_component[i] *= start + (end - start)/(1.0 + exp(-n));
+        wide_component[i] *= start + (end - start)/(1.0 + Lookup::evaluate(n));
     }
 
     // Normalise to amplitude

### 2016-12-04 src/display.py
if name == main

@@ -61,3 +61,6 @@ def display():
            b=wide_fraction.std()))
     plt.show()
 
+if __name__ == "__main__":
+    display()
+

### 2016-12-05 doc/ms.tex
More work on the table
Bit more in the table

@@
 
+All of the assumptions are given in Table~\ref{tab:priors}.
+
+\begin{table}
+\centering
+\begin{tabular}{|lll|}
+\hline
+{\bf Quantity}      &   {\bf Meaning}   &  {\bf Prior distribution}\\
+\hline
+{\em Hyperparameters} & &\\
+$N$     &   Number of spikes    &  $p(N) \propto 1/(N+1), N \in \{0, 1, ..., 300\}$\\
+\hline
+{\em Parameters}& &\\
+$\sigma_0$ &    Constant noise level  &   $\ln(\sigma_0) \sim \textnormal{Laplace}(0,5)$\\
+$\sigma_1$ &    Noise proportionality   &  $\ln(\sigma_1) \sim \textnormal{Laplace}(0,5)$ \\
+$\nu$     &   Shape parameter for residuals   &   Loguniform$(1, 1000)$\\
+\hline
+{\em Data}&&\\
+\hline
+$\{y_1, y_2, ..., y_n\}$  &   Measurements    & $t(f(x_i; \params), ,\nu)$\\
+\hline
+{\em Prior information}&&\\
+\hline
+$n$ & Number of measurements & given\\
+$\{x_1, x_2, ..., x_n\}$  & $x$-values of measurements & given \\
+
+\hline
+\end{tabular}
+\caption{\label{tab:priors}}
+\end{table}
+
 \bibliographystyle{plainnat}

### 2016-12-26 src/display.py
Plot wide component and spikes

@@ -48,6 +48,9 @@ def display():
         # Plot the model
         if i < 50:
             plt.plot(data[:,0], model, "g", linewidth=2, alpha=0.1)
+            plt.plot(data[:,0], wide_component, "b", linewidth=2, alpha=0.1)
+            plt.plot(data[:,0], the_spikes, "r", linewidth=2, alpha=0.1)
+            plt.ylim(0)
 
     plt.xlabel("$x$", fontsize=16)
     plt.ylabel("$y$", fontsize=16)

### 2016-12-26 src/MyModel.cpp
Bugfix

@@ -163,7 +163,7 @@ void MyModel::compute_wide_component()
     for(size_t i=0; i<wide_component.size(); ++i)
     {
         n = (data_x[i] - center)/(r_asymmetry*width);
-        wide_component[i] *= start + (end - start)/(1.0 + Lookup::evaluate(n));
+        wide_component[i] *= start + (end - start)/(1.0 + exp(-n));
     }
 
     // Normalise to amplitude

### 2016-12-26 src/display.py
plot flat component

@@ -50,6 +50,8 @@ def display():
             plt.plot(data[:,0], model, "g", linewidth=2, alpha=0.1)
             plt.plot(data[:,0], wide_component, "b", linewidth=2, alpha=0.1)
             plt.plot(data[:,0], the_spikes, "r", linewidth=2, alpha=0.1)
+            plt.plot(data[:,0], posterior_sample[i, 0]*np.ones(len(model)),\
+                                                "y", linewidth=2, alpha=0.1)
             plt.ylim(0)
 
     plt.xlabel("$x$", fontsize=16)

### 2016-12-26 src/MyConditionalPrior.cpp
Less diverse widths

@@ -22,8 +22,8 @@ void MyConditionalPrior::from_prior(DNest4::RNG& rng)
     location_log_amplitude = laplace.generate(rng);
     scale_log_amplitude = 5*rng.rand();
 
-    K = rng.rand();
-    max_width = exp(log(0.001*x_range) + log(100.0)*rng.rand());
+    location_log_width = log(0.001*x_range) + log(100.0)*rng.rand();
+    scale_log_width = 0.2*rng.rand();
 }
 
 double MyConditionalPrior::perturb_hyperparameters(DNest4::RNG& rng)
@@ -42,15 +42,13 @@ double MyConditionalPrior::perturb_hyperparameters(DNest4::RNG& rng)
     }
     else if(which == 2)
     {
-        K += rng.randh();
-        DNest4::wrap(K, 0.0, 1.0);
+        location_log_width += log(100.0)*rng.randh();
+        DNest4::wrap(location_log_width, log(0.001*x_range), log(0.1*x_range));
     }
     else
     {
-        max_width = log(max_width);
-        max_width += log(100.0)*rng.randh();
-        DNest4::wrap(max_width, log(0.001*x_range), log(0.1*x_range));
-        max_width = exp(max_width);
+        scale_log_width += 0.2*rng.randh();
+        DNest4::wrap(scale_log_width, 0.0, 0.2);
     }
 
     return logH;
@@ -60,12 +58,10 @@ double MyConditionalPrior::perturb_hyperparameters(DNest4::RNG& rng)
 
 double MyConditionalPrior::log_pdf(const std::vector<double>& vec) const
 {
-    double min_width = K*max_width;
-
     if(vec[0] < x_min || vec[0] > x_max)
         return -1E300;
 
-    if(vec[2] < min_width || vec[2] > max_width)
+    if(vec[2] <= 0.0)
         return -1E300;
 
     DNest4::Laplace l(location_log_amplitude, scale_log_amplitude);
@@ -73,35 +69,37 @@ double MyConditionalPrior::log_pdf(const std::vector<double>& vec) const
     double logp = 0.0;
 
     logp += l.log_pdf(vec[1]);
-    logp += -log(max_width - min_width);
+
+    l = DNest4::Laplace(location_log_width, scale_log_width);
+    logp += -log(vec[2]) + l.log_pdf(log(vec[2]));
 
     return logp;
 }
 
[…]

### 2016-12-26 src/MyConditionalPrior.h

@@ -16,7 +16,7 @@ class MyConditionalPrior:public DNest4::ConditionalPrior
         double location_log_amplitude, scale_log_amplitude;
 
         // For widths
-        double K, max_width;
+        double location_log_width, scale_log_width;
 
         double perturb_hyperparameters(DNest4::RNG& rng);
 

### 2016-12-26 src/MyModel.cpp
Updated header

@@ -259,7 +259,7 @@ std::string MyModel::description() const
     s<<"asymmetry, r_asymmetry, ";
     s<<"dim_spikes, max_num_spikes, ";
     s<<"location_log_amplitude, scale_log_amplitude, ";
-    s<<"K, max_width, num_spikes, ";
+    s<<"location_log_width, scale_log_width, num_spikes, ";
     for(size_t i=0; i<max_num_spikes; ++i)
         s<<"center["<<i<<"], ";
     for(size_t i=0; i<max_num_spikes; ++i)

### 2016-12-26 src/display.py
Transparent data

@@ -24,7 +24,7 @@ def display():
             plt.clf()
 
             # Plot the data
-            plt.plot(data[:,0], data[:,1], "ko", markersize=3)
+            plt.plot(data[:,0], data[:,1], "ko", markersize=3, alpha=0.2)
             plt.hold(True)
 
         # Extract the wide component

### 2017-01-26 doc/ms.tex
Crystallinity
A bit of writing about the shape of the model curve

@@
 \usepackage{amsmath}
+\usepackage{color}
 \usepackage{dsfont}
@@
 
+\definecolor{orange}{rgb}{1, 0.5, 0}
+\definecolor{green}{rgb}{0, 0.5, 0}
+
 \newcommand{\hypers}{\boldsymbol{\alpha}}
@@
 \newcommand{\info}{\boldsymbol{M}}
+\newcommand{\x}{x}
+\newcommand{\todo}{\color{orange} \bf}
 
@@
 \section{The prior information}
+
+\subsection{The shape of the model curve}
+The function $f(\x)$, defined from $\x_{\rm min}$ to $\x_{\rm max}$,
+is assumed to be a sum of the following components:
+\begin{enumerate}
+\item A background component
+\item $N_{\rm wide}$ wide gaussians
+\item $N_{\rm narrow}$ narrow gaussians, to model the spikes
+\end{enumerate}
+
+The background component is assumed to be linear
+{\todo between $x_{\rm min}$ and A, between A and B, and between B
+and $x_{\rm max}$}.
+
+\subsection{Prior probability distributions}
 
@@
 
+The shape of the model function $f(x)$ is therefore
+\begin{align}
+f(x) &= A + f_{\rm wide}(x) + \sum_{i=1}^N a_i f_i(x)
+\end{align}
+and the crystallinity is
+\begin{align}
+C &= \frac{\int f_{\rm wide}(x) \, dx}
+          {\int f_{\rm wide}(x) \, dx + \int f_{\rm narrow}(x) \, dx}
+\end{align}
+
 \bibliographystyle{plainnat}

### 2017-01-26 doc/ms.tex
Bit more in the table

@@
 {\em Hyperparameters} & &\\
-$N$     &   Number of spikes    &  $p(N) \propto 1/(N+1), N \in \{0, 1, ..., 300\}$\\
+$N_{\rm narrow}$   &   Number of narrow spikes    &  $\propto 1/(N_{\rm narrow}+1)$, $N_{\rm narrow} \in \{0, 1, ..., 300\}$ \\
+$N_{\rm wide}$   &   Number of wide spikes    &  $\propto 1/(N_{\rm wide}+1)$,
+$N_{\rm wide} \in \{0, 1, ..., 300\}$ \\
 \hline
@@
 $\sigma_1$ &    Noise proportionality   &  $\ln(\sigma_1) \sim \textnormal{Laplace}(0,5)$ \\
-$\nu$     &   Shape parameter for residuals   &   Loguniform$(1, 1000)$\\
+$\nu$     &   Shape parameter for residuals   &   LogUniform$(1, 1000)$\\
+$b$       & Mean background level       & $\ln(b) \sim$ Laplace$(0, 5)$\\
+$\{n_1, ..., n_4\}$  & 
 \hline

### 2017-01-26 doc/ms.tex
Wrote some stuff that needs to be written...

@@
 $N_{\rm wide} \in \{0, 1, ..., 300\}$ \\
+
 \hline
@@
 $\nu$     &   Shape parameter for residuals   &   LogUniform$(1, 1000)$\\
-$b$       & Mean background level       & $\ln(b) \sim$ Laplace$(0, 5)$\\
-$\{n_1, ..., n_4\}$  & 
+$b$       & Mean background level       & Laplace$(0, 5)$\\
+$\{n_1, ..., n_4\}$  & Background deviations & iid Normal$(0,1)$\\
 \hline
@@
 \hline
-$\{y_1, y_2, ..., y_n\}$  &   Measurements    & $t(f(x_i; \params), ,\nu)$\\
+$\{y_1, y_2, ..., y_n\}$  &   Measurements    & Student $t(f(x_i; \params), ,\nu)$\\
 \hline
@@
 $\{x_1, x_2, ..., x_n\}$  & $x$-values of measurements & given \\
-
 \hline
@@
 
+The Laplace, or biexponential, distribution with location parameter $a$ and
+scale parameter $b$ has probability density
+\begin{align}
+p(x | a, b) &= \frac{1}{2b}\exp\left[-\frac{|x - a|}{b}\right].
+\end{align}
+
 The shape of the model function $f(x)$ is therefore
 \begin{align}
-f(x) &= A + f_{\rm wide}(x) + \sum_{i=1}^N a_i f_i(x)
+f(x) &= f_{\rm bg}(x) + f_{\rm wide}(x) + f_{\rm narrow}(x)
 \end{align}

### 2017-01-27 doc/ms.tex
Wrote a tad

@@
 \subsection{The shape of the model curve}
-The function $f(\x)$, defined from $\x_{\rm min}$ to $\x_{\rm max}$,
-is assumed to be a sum of the following components:
+The algorithm fits the data with a model curve
+$f(\x)$, defined between $\x=\x_{\rm min}$ and $\x=\x_{\rm max}$,
+which is assumed to be a sum of the following components:
 \begin{enumerate}
@@
 \item $N_{\rm wide}$ wide gaussians
-\item $N_{\rm narrow}$ narrow gaussians, to model the spikes
+\item $N_{\rm narrow}$ narrow gaussians, to model the spikes.
 \end{enumerate}
+
+We compute the posterior distribution for the unknown parameters of the
+components, and use \citep{dnest4} to generate samples from that posterior
+distribution, representing plausible parameter values. From each sample,
+it is possible to compute the crystallinity, making it straightforward to
+calculate the marginal posterior distribution of the crystallinity, describing
+uncertainty about its value in the light of the data.
 

### 2017-01-27 doc/ms.tex
Background component model

@@
 
-The background component is assumed to be linear
-{\todo between $x_{\rm min}$ and A, between A and B, and between B
-and $x_{\rm max}$}.
+The background component is assumed to be piecewise linear
+{\todo between $\x=x_{\rm min}$ and $\x=10$, between $\x=10$ and $\x=40$,
+and between $\x=40$ and $\x=\x_{\rm max}$}.
+We parameterise the background component using
+five parameters. Firstly, there is a mean level $b$, which sets the
+typical amplitude of the background component. There are also four parameters
+$n_1, ..., n_4$, which set the positions of four control points:
+\begin{align}
+\left(\x_{\rm min}, be^{n_1}\right) \nonumber\\
+\left(10, be^{n_2}\right) \nonumber\\
+\left(40, be^{n_3}\right) \nonumber\\
+\left(\x_{\rm max}, be^{n_4}\right) \nonumber
+\end{align}
+The value of the background component at any other point is found by
+linear interpolation.
 

### 2017-01-31 doc/figures.py
A script to make figures

@@ -1,0 +1,34 @@
+import numpy as np
+import numpy.random as rng
+import matplotlib.pyplot as plt
+
+# Set rng seed
+rng.seed(3)
+
+# x coordinate
+x = np.linspace(5.0, 50.0, 1001)
+
+# Background level
+b = 1.0
+
+# Create figure
+plt.figure(1)
+plt.hold(True)
+
+for i in range(0, 3):
+    # Control point parameters
+    n = rng.randn(4)
+
+    # The four control points
+    xx = np.array([x.min(), 10.0, 40.0, x.max()])
+    yy = b * np.exp(n)
+
+    # Plot the background component shape
+    plt.plot(xx, yy, "ko-", alpha=0.2)
+
+plt.xlabel("$x$", fontsize=20)
+plt.ylabel("$y$", fontsize=20)
+plt.title("Background component", fontsize=14)
+plt.savefig("figures/background.pdf", bbox_inches="tight")
+plt.show()
+

### 2017-01-31 doc/ms.tex
Background figure
Use \x throughout

@@
 The background component is assumed to be piecewise linear
-{\todo between $\x=x_{\rm min}$ and $\x=10$, between $\x=10$ and $\x=40$,
+{\todo between $\x=\x_{\rm min}$ and $\x=10$, between $\x=10$ and $\x=40$,
 and between $\x=40$ and $\x=\x_{\rm max}$}.
@@
 The value of the background component at any other point is found by
-linear interpolation.
+linear interpolation. Figure~\ref{fig:background} shows three example
+background curves generated from this parameterisation, with typical
+values of the parameters.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.7\textwidth]{figures/background.pdf}
+\caption{\label{fig:background}}
+\end{figure}
 
@@
 $n$ & Number of measurements & given\\
-$\{x_1, x_2, ..., x_n\}$  & $x$-values of measurements & given \\
+$\{\x_1, \x_2, ..., \x_n\}$  & $\x$-values of measurements & given \\
 \hline
@@
 \begin{align}
-f(x) &= f_{\rm bg}(x) + f_{\rm wide}(x) + f_{\rm narrow}(x)
+f(\x) &= f_{\rm bg}(\x) + f_{\rm wide}(\x) + f_{\rm narrow}(\x)
 \end{align}
-and the crystallinity is
+and the crystallinity is defined as the proportion of the area in the
+wide and narrow components which is accounted for by the wide component:
 \begin{align}
-C &= \frac{\int f_{\rm wide}(x) \, dx}
-          {\int f_{\rm wide}(x) \, dx + \int f_{\rm narrow}(x) \, dx}
+C &= \frac{\int f_{\rm wide}(\x) \, d\x}
+          {\int f_{\rm wide}(\x) \, d\x + \int f_{\rm narrow}(\x) \, d\x}
 \end{align}

### 2017-02-10 src/display.py
Compute skewness; removed hold calls

@@ -24,6 +24,8 @@ def display():
     wide_integral = np.zeros(posterior_sample.shape[0])
     wide_center_of_mass = np.zeros(posterior_sample.shape[0])
     wide_width = np.zeros(posterior_sample.shape[0])
+    wide_skewness = np.zeros(posterior_sample.shape[0])
+    wide_nongaussianity = np.zeros(posterior_sample.shape[0])
 
     spikes_integral = np.zeros(posterior_sample.shape[0])
     max_num_spikes = int(posterior_sample[0, indices["max_num_gaussians1"]])
@@ -35,7 +37,6 @@ def display():
 
             # Plot the data
             plt.plot(data[:,0], data[:,1], "ko", markersize=3, alpha=0.2)
-            plt.hold(True)
 
         # Extract the background
         start = indices["bg[0]"]
@@ -63,6 +64,8 @@ def display():
         wide_width[i] = np.sqrt(\
                         np.sum((x - wide_center_of_mass[i])**2*wide_component)\
                             / wide_integral[i])
+        wide_skewness[i] = np.sum(((x - wide_center_of_mass[i])\
+                                    /wide_width[i])**3) / wide_integral[i]
 
         spikes_integral[i] = np.sum(the_spikes)
 
@@ -96,6 +99,12 @@ def display():
     plt.xlabel("Width of wide component")
     print("Width = {a} +- {b}".format(a=wide_width.mean(),
            b=wide_width.std()))
+    plt.show()
+
+    plt.hist(wide_skewness, 100, color=[0.8, 0.8, 0.8])
+    plt.xlabel("Skewness of wide component")
+    print("Skewness = {a} +- {b}".format(a=wide_skewness.mean(),
+           b=wide_skewness.std()))
     plt.show()
 
 if __name__ == "__main__":

### 2017-02-10 src/MyModel.cpp
Get control points from file

@@ -10,7 +10,7 @@ namespace Crystals
 // Static things
 Data MyModel::data;
 const DNest4::Laplace MyModel::laplace(0.0, 5.0);
-const std::vector<double> MyModel::x_knots{data.get_x_min(),
+std::vector<double> MyModel::x_knots{data.get_x_min(),
                                             10.0, 40.0, data.get_x_max()};
 
 // Constructor
@@ -242,6 +242,17 @@ double MyModel::log_likelihood() const
         logL = -1E300;
 
     return logL;
+}
+
+void MyModel::load_control_points(const char* filename)
+{
+    std::fstream fin(filename, std::ios::in);
+    if(!fin)
+        std::cerr<<"# ERROR couldn't open "<<filename<<"."<<std::endl;
+
+    fin>>x_knots[1];
+    fin>>x_knots[2];
+    fin.close();
 }
 
 void MyModel::print(std::ostream& out) const

### 2017-02-10 src/MyModel.h

@@ -63,13 +63,14 @@ class MyModel
         static const DNest4::Laplace laplace;
 
         // x-positions of background knots
-        static const std::vector<double> x_knots;
+        static std::vector<double> x_knots;
 
         // Maximum number of spikes
         static constexpr size_t max_num_spikes = 300;
 
     public:
         static void load_data(const char* filename);
+        static void load_control_points(const char* filename);
 };
 
 } // namespace Crystals

### 2017-02-10 src/main.cpp

@@ -19,6 +19,9 @@ int main(int argc, char** argv)
         std::cin >> data_file;
     }
 
+    // Load the control points
+    Crystals::MyModel::load_control_points("control_points.in");
+
     // Save the data filename
     std::fstream fout("run_data.txt", std::ios::out);
     fout<<data_file;

### 2017-02-10 src/display.py
Calculate nongaussianity (needs testing)

@@ -67,6 +67,15 @@ def display():
         wide_skewness[i] = np.sum(((x - wide_center_of_mass[i])\
                                     /wide_width[i])**3) / wide_integral[i]
 
+        # Best fitting gaussian to the wide component
+        gaussian = np.exp(-0.5*(x - wide_center_of_mass[i])**2 \
+                            / wide_width[i]**2)
+        gaussian /= gaussian.sum()
+        f = wide_component / wide_component.sum()
+
+        # Nongaussianity based on KL divergence
+        wide_nongaussianity[i] = np.sum(f*np.log(f / gaussian + 1E-300))
+
         spikes_integral[i] = np.sum(the_spikes)
 
         # Plot the model
@@ -105,6 +114,12 @@ def display():
     plt.xlabel("Skewness of wide component")
     print("Skewness = {a} +- {b}".format(a=wide_skewness.mean(),
            b=wide_skewness.std()))
+    plt.show()
+
+    plt.hist(wide_nongaussianity, 100, color=[0.8, 0.8, 0.8])
+    plt.xlabel("Nongaussianity of wide component")
+    print("Nongaussianity = {a} +- {b}".format(a=wide_nongaussianity.mean(),
+           b=wide_nongaussianity.std()))
     plt.show()
 
 if __name__ == "__main__":

### 2017-02-10 src/display.py
Fixed and simplified some formulae

@@ -60,18 +60,17 @@ def display():
 
         # Compute some integrals
         wide_integral[i] = np.sum(wide_component)
-        wide_center_of_mass[i] = np.sum(x * wide_component) / wide_integral[i]
-        wide_width[i] = np.sqrt(\
-                        np.sum((x - wide_center_of_mass[i])**2*wide_component)\
-                            / wide_integral[i])
-        wide_skewness[i] = np.sum(((x - wide_center_of_mass[i])\
-                                    /wide_width[i])**3) / wide_integral[i]
+        f = wide_component / wide_integral[i] # Normalised
+
+        wide_center_of_mass[i] = np.sum(x * f)
+        wide_width[i] = np.sqrt(np.sum((x - wide_center_of_mass[i])**2*f))
+        wide_skewness[i] = np.sum(f *\
+                            ((x - wide_center_of_mass[i]) / wide_width[i])**3)
 
         # Best fitting gaussian to the wide component
         gaussian = np.exp(-0.5*(x - wide_center_of_mass[i])**2 \
                             / wide_width[i]**2)
         gaussian /= gaussian.sum()
-        f = wide_component / wide_component.sum()
 
         # Nongaussianity based on KL divergence
         wide_nongaussianity[i] = np.sum(f*np.log(f / gaussian + 1E-300))

### 2017-02-10 src/display.py
Positive quantity so start histogram at zero

@@ -117,6 +117,7 @@ def display():
 
     plt.hist(wide_nongaussianity, 100, color=[0.8, 0.8, 0.8])
     plt.xlabel("Nongaussianity of wide component")
+    plt.xlim(0.0)
     print("Nongaussianity = {a} +- {b}".format(a=wide_nongaussianity.mean(),
            b=wide_nongaussianity.std()))
     plt.show()

### 2017-02-12 doc/figures.py
Niceties

@@ -2,6 +2,10 @@ import numpy as np
 import numpy.random as rng
 import matplotlib.pyplot as plt
 
+# Set up fonts
+plt.rc("font", size=14, family="serif", serif="Computer Sans")
+plt.rc("text", usetex=True)
+
 # Set rng seed
 rng.seed(3)
 
@@ -13,8 +17,6 @@ b = 1.0
 
 # Create figure
 plt.figure(1)
-plt.hold(True)
-
 for i in range(0, 3):
     # Control point parameters
     n = rng.randn(4)
@@ -26,9 +28,9 @@ for i in range(0, 3):
     # Plot the background component shape
     plt.plot(xx, yy, "ko-", alpha=0.2)
 
-plt.xlabel("$x$", fontsize=20)
-plt.ylabel("$y$", fontsize=20)
-plt.title("Background component", fontsize=14)
+plt.xlabel("$x$", fontsize=16)
+plt.ylabel("$y$", fontsize=16)
+plt.title("Background component")
 plt.savefig("figures/background.pdf", bbox_inches="tight")
 plt.show()
 

### 2017-02-12 doc/ms.tex
Minor work

@@
 \includegraphics[width=0.7\textwidth]{figures/background.pdf}
-\caption{\label{fig:background}}
+\caption{Three example functions showing possible shapes of the
+background component.\label{fig:background}}
 \end{figure}
@@
 \hline
-{\em Parameters}& &\\
+{\em Background parameters}\\
+$b$       & Mean background level       & Laplace$(0, 5)$\\
+$\{n_1, ..., n_4\}$  & Background deviations & iid Normal$(0,1)$\\
+\hline 
+{\em Noise parameters}\\
 $\sigma_0$ &    Constant noise level  &   $\ln(\sigma_0) \sim \textnormal{Laplace}(0,5)$\\
@@
 $\nu$     &   Shape parameter for residuals   &   LogUniform$(1, 1000)$\\
-$b$       & Mean background level       & Laplace$(0, 5)$\\
-$\{n_1, ..., n_4\}$  & Background deviations & iid Normal$(0,1)$\\
 \hline
@@
 \hline
-$\{y_1, y_2, ..., y_n\}$  &   Measurements    & Student $t(f(x_i; \params), ,\nu)$\\
+$\{y_1, y_2, ..., y_n\}$  &   Measurements    & Student-$t(f(x_i; \params), ,\nu)$\\
 \hline
@@
 
-The Laplace, or biexponential, distribution with location parameter $a$ and
-scale parameter $b$ has probability density
+The Laplace, or biexponential, distribution with location parameter
+(central position) $a$ and scale parameter (width) $b$ has probability
+density given by
 \begin{align}
@@
 \end{align}
+We used a Laplace distribution for informative priors, rather than
+a more conventional normal distribution, because it assigns higher
+prior probability to values far from the center, and can be computed conveniently, i.e., no special functions such as {\tt erf} are needed
+to compute the cumulative distribution function.
 

### 2017-02-12 doc/ms.tex
Entered more hyperparameters

@@
 \begin{table}
+\footnotesize
 \centering
@@
 \hline
+{\em Prior information}&&\\
+\hline
+$n$ & Number of measurements & given\\
+$\{\x_1, \x_2, ..., \x_n\}$  & $\x$-values of measurements & given \\
+\hline
 {\em Hyperparameters} & &\\
@@
 $N_{\rm wide} \in \{0, 1, ..., 300\}$ \\
-
+$a_{\ln A, \rm narrow}$ & Typical log-amplitude of narrow spikes & Laplace(0, 5)\\
+$a_{\ln A, \rm wide}$ & Typical log-amplitude of wide spikes & Laplace(0, 5)\\
+$b_{\ln A, \rm narrow}$ & Diversity of log-amplitude of narrow spikes & Uniform(0, 5)\\
+$b_{\ln A, \rm wide}$ & Diversity of log-amplitude of wide spikes & Uniform(0, 5)\\
+$a_{\ln W, \rm narrow}$ & Typical log-width of narrow spikes & LogUniform$(10^{-3}x_{\rm range}, 0.05x_{\rm range})$\\
+$a_{\ln W, \rm wide}$ & Typical log-width of wide spikes & LogUniform$(0.05x_{\rm range}, x_{\rm range})$\\
+$b_{\ln W, \rm narrow}$ & Diversity of log-width of narrow spikes & Uniform(0, 0.1)\\
+$b_{\ln W, \rm wide}$ & Diversity of log-width of wide spikes & Uniform(0, 0.2)\\
 \hline
-{\em Background parameters}\\
+{\em Background parameters}&&\\
 $b$       & Mean background level       & Laplace$(0, 5)$\\
@@
 \hline 
-{\em Noise parameters}\\
-$\sigma_0$ &    Constant noise level  &   $\ln(\sigma_0) \sim \textnormal{Laplace}(0,5)$\\
-$\sigma_1$ &    Noise proportionality   &  $\ln(\sigma_1) \sim \textnormal{Laplace}(0,5)$ \\
+{\em Noise parameters}&&\\
+$\sigma_0$ &    Constant noise level  &   $\ln \sigma_0 \sim \textnormal{Laplace}(0,5)$\\
+$\sigma_1$ &    Noise proportionality   &  $\ln \sigma_1 \sim \textnormal{Laplace}(0,5)$ \\
 $\nu$     &   Shape parameter for residuals   &   LogUniform$(1, 1000)$\\
@@
 \hline
-$\{y_1, y_2, ..., y_n\}$  &   Measurements    & Student-$t(f(x_i; \params), ,\nu)$\\
-\hline
-{\em Prior information}&&\\
-\hline
-$n$ & Number of measurements & given\\
-$\{\x_1, \x_2, ..., \x_n\}$  & $\x$-values of measurements & given \\
+$\{y_1, y_2, ..., y_n\}$  &   Measurements    & iid Student-$t(f(x_i; \params), ,\nu)$\\
 \hline

### 2017-02-13 doc/ms.tex
more work on that big table

@@
 \usepackage[left=2cm, right=2cm, bottom=3cm, top=2cm]{geometry}
+\usepackage{microtype}
 \usepackage{natbib}
-\usepackage{microtype}
+\usepackage{pdflscape}
 
@@
 
+\begin{landscape}
+
 \begin{table}
@@
 \hline
+{\em Narrow spike parameters}&&\\
+$X_{i, \rm narrow}$ & Positions of narrow spikes &
+                            Uniform$(x_{\rm min}, x_{\rm max})$ \\
+$A_{i, \rm narrow}$ & Amplitudes of narrow spikes &
+ $\ln A_{i, \rm narrow} \sim \textnormal{Laplace}\left(a_{\ln A, \rm narrow}, b_{\ln A, \rm narrow}\right)$ \\
+$W_{i, \rm narrow}$ & Widths of narrow spikes &
+ $\ln W_{i, \rm narrow} \sim \textnormal{Laplace}\left(a_{\ln W, \rm narrow}, b_{\ln W, \rm narrow}\right)$ \\
+\hline
+{\em Wide spike parameters}&&\\
+$X_{i, \rm wide}$ & Positions of wide spikes &
+          Uniform$(x_{\rm min}+0.3x_{\rm range}, x_{\rm max}-0.3x_{\rm range})$ \\
+$A_{i, \rm wide}$ & Amplitudes of wide spikes &
+ $\ln A_{i, \rm wide} \sim \textnormal{Laplace}\left(a_{\ln A, \rm wide}, b_{\ln A, \rm wide}\right)$ \\
+$W_{i, \rm wide}$ & Widths of wide spikes &
+ $\ln W_{i, \rm wide} \sim \textnormal{Laplace}\left(a_{\ln W, \rm wide}, b_{\ln W, \rm wide}\right)$ \\
+\hline
 {\em Noise parameters}&&\\
@@
 \end{table}
+
+\end{landscape}
 

### 2017-02-13 doc/ms.tex
Fixed mistake

@@
 {\em Background parameters}&&\\
-$b$       & Mean background level       & Laplace$(0, 5)$\\
+$b$       & Mean background level       & $\ln b \sim $ Laplace$(0, 5)$\\
 $\{n_1, ..., n_4\}$  & Background deviations & iid Normal$(0,1)$\\

### 2017-02-19 doc/ms.tex
Bayes paragraph

@@
 \section{The prior information}
+In the Bayesian framework, probability distributions are used to model
+states of uncertainty about the values of unknown quantities.
+Typically, one starts with the ``prior distribution''
+for unknown quantities (`parameters') $\params$, written $p(\params | \info)$
+(the probability distribution for $\params$ {\em given} prior information
+$\info$). This is supplemented with $p(\data | \params, \info)$,
+which describes the uncertainty about the data which we {\em would} have,
+if we knew the parameters (and the prior information). The
+product of these is the joint prior
+\begin{align}
+p(\params, \data | \info) &= p(\params | \info)p(\data | \params, \info).
+\end{align}
+Once a specific observed dataset $\data_o$ is known, knowledge about
+$\params$ is updated from the prior to the posterior
+\begin{align}
+p(\params | \data, \info) &\propto
+    p(\params | \info)p(\data_o | \params, \info).
+\end{align}
+which takes the data into account.
 

### 2017-03-01 doc/ms.tex
Working on the document

@@
 \newcommand{\params}{\boldsymbol{\theta}}
-\newcommand{\data}{\boldsymbol{x}}
-\newcommand{\info}{\boldsymbol{M}}
+\newcommand{\data}{\boldsymbol{d}}
+\newcommand{\info}{I}
 \newcommand{\x}{x}
@@
 
-\section{The prior information}
+\section{Bayesian inference}
 In the Bayesian framework, probability distributions are used to model
@@
 \begin{align}
-p(\params | \data, \info) &\propto
+p(\params | \data_0, \info) &\propto
     p(\params | \info)p(\data_o | \params, \info).
@@
 
-\subsection{The shape of the model curve}
+\section{The shape of the model curve}
 The algorithm fits the data with a model curve
@@
 \begin{enumerate}
-\item A background component
-\item $N_{\rm wide}$ wide gaussians
-\item $N_{\rm narrow}$ narrow gaussians, to model the spikes.
+\item A background component, described by parameters $\params_{\rm bg}$
+\item A wide component, described by parameters $\params_{\rm wide}$
+\item Narrow spikes, described by parameters $\params_{\rm narrow}$
 \end{enumerate}
+
+The total model function $f_{\rm tot}(x)$ is therefore
+\begin{align}
+f_{\rm tot}(\x) &= f_{\rm bg}(\x; \params_{\rm bg})
+       + f_{\rm wide}(\x; \params_{\rm wide})
+       + f_{\rm narrow}(\x; \params_{\rm narrow})
+\end{align}
+and the crystallinity is defined as the proportion of the area in the
+wide and narrow components which is accounted for by the wide component:
+\begin{align}
+C &= \frac{\int_{\x_{\rm min}}^{\x_{\rm max}}
+            f_{\rm wide}(\x; \params_{\rm wide}) \, d\x}
+          {\int_{\x_{\rm min}}^{\x_{\rm max}}
+            \left[f_{\rm wide}(\x; \params_{\rm wide})
+            + f_{\rm narrow}(\x; \params_{\rm narrow}) \right] \, d\x}
+\end{align}
 
@@
 calculate the marginal posterior distribution of the crystallinity, describing
-uncertainty about its value in the light of the data.
+the remaining uncertainty about its value in the light of the data.
 
@@
 
-The shape of the model function $f(x)$ is therefore
-\begin{align}
-f(\x) &= f_{\rm bg}(\x) + f_{\rm wide}(\x) + f_{\rm narrow}(\x)
-\end{align}
-and the crystallinity is defined as the proportion of the area in the
-wide and narrow components which is accounted for by the wide component:
-\begin{align}
-C &= \frac{\int f_{\rm wide}(\x) \, d\x}
-          {\int f_{\rm wide}(\x) \, d\x + \int f_{\rm narrow}(\x) \, d\x}
-\end{align}
-
 \bibliographystyle{plainnat}

### 2017-03-06 doc/ms.tex
Narrow spikes formula

@@
 which describes the uncertainty about the data which we {\em would} have,
-if we knew the parameters (and the prior information). The
+if we knew the parameters (and the prior information), as a function of
+the parameters.
+
+The
 product of these is the joint prior
@@
 \end{align}
-which takes the data into account.
+which takes the specific data $\data_0$ into account.
 
-\section{The shape of the model curve}
+\section{The model curve}
 The algorithm fits the data with a model curve
@@
 We compute the posterior distribution for the unknown parameters of the
-components, and use \citep{dnest4} to generate samples from that posterior
-distribution, representing plausible parameter values. From each sample,
+components, using \citep{dnest4} to generate samples from that posterior
+distribution. The output is a collection of plausible values of the
+parameters which can be used to approximate posterior probabilities.
+For example, the probability some parameter $\theta$ is greater than 3.4
+is approximately the proportion of the posterior samples that have
+$\theta > 3.4$.
+
+From each sample,
 it is possible to compute the crystallinity, making it straightforward to
@@
 
+In the following subsections, I describe the specific parameterisations of the
+three parts of the model curve.
+
+\subsection{The background component}
 The background component is assumed to be piecewise linear
@@
 \end{figure}
+
+\subsection{The narrow spikes}
+I assumed that there is some number $N_{\rm narrow}$ of narrow spikes,
+with central positions $X_1, X_2, ..., X_{N, \rm narrow}$, amplitudes
+$A_1, A_2, ..., A_{N, \rm narrow}$, and widths
+$W_1, W_2, ..., W_{N, \rm narrow}$. The total shape contributed by the
+narrow spikes is a sum of $N_{\rm narrow}$ gaussians:
+\begin{align}
+f_{\rm narrow}(\x; \params_{\rm narrow}) &=
+    \sum_{i=1}^{N_{\rm narrow}} A_i
+                        \exp\left[-\frac{1}{2W_i^2}\left(x - X_i\right)^2\right]
+\end{align}
+
+\subsection{The wide `spikes'}
 

### 2017-03-08 doc/ms.tex
Narrow and wide gaussian sections

@@
 I assumed that there is some number $N_{\rm narrow}$ of narrow spikes,
-with central positions $X_1, X_2, ..., X_{N, \rm narrow}$, amplitudes
-$A_1, A_2, ..., A_{N, \rm narrow}$, and widths
-$W_1, W_2, ..., W_{N, \rm narrow}$. The total shape contributed by the
-narrow spikes is a sum of $N_{\rm narrow}$ gaussians:
+each of which has a gaussian shape. The $i$th narrow spike is parameterised
+by its central position $X_{i, \rm narrow}$,
+amplitude $A_{i, \rm narrow}$, and width $A_{i, \rm narrow}$.
+The parameter vector for narrow spikes is
+\begin{align}
+\params_{\rm narrow} &=
+  \{N_{\rm narrow}, X_1, X_2, ..., X_{N, \rm narrow},
+    A_1, A_2, ..., A_{N, \rm narrow},
+    W_1, W_2, ..., W_{N, \rm narrow}\}
+\end{align}
+and the total shape contributed by the
+narrow spikes is a sum of the $N_{\rm narrow}$ gaussians:
 \begin{align}
 f_{\rm narrow}(\x; \params_{\rm narrow}) &=
-    \sum_{i=1}^{N_{\rm narrow}} A_i
-                        \exp\left[-\frac{1}{2W_i^2}\left(x - X_i\right)^2\right]
+    \sum_{i=1}^{N_{\rm narrow}} A_{i, \rm narrow}
+ \exp\left[-\frac{1}{2W_{i, \rm narrow}^2}\left(x - X_{i, \rm narrow}\right)^2\right].
 \end{align}
 
-\subsection{The wide `spikes'}
+\subsection{The wide component}
+The wide component is also composed of gaussians, like the narrow spikes.
+In fact, the only real difference is the prior distributions for the positions,
+amplitudes, and widths, given in Section~\ref{sec:priors}.
 
-\subsection{Prior probability distributions}
+There is some number $N_{\rm wide}$ of gaussians in the wide component.
+The $i$th gaussian is parameterised
+by its central position $X_{i, \rm wide}$,
+amplitude $A_{i, \rm wide}$, and width $A_{i, \rm wide}$.
+The parameter vector for wide gaussians is
+\begin{align}
+\params_{\rm wide} &=
+  \{N_{\rm wide}, X_1, X_2, ..., X_{N, \rm wide},
+    A_1, A_2, ..., A_{N, \rm wide},
+    W_1, W_2, ..., W_{N, \rm wide}\}
+\end{align}
+and the total shape contributed by the
+wide component is a sum of the $N_{\rm wide}$ gaussians:
+\begin{align}
+f_{\rm wide}(\x; \params_{\rm wide}) &=
+    \sum_{i=1}^{N_{\rm wide}} A_{i, \rm wide}
+ \exp\left[-\frac{1}{2W_{i, \rm wide}^2}\left(x - X_{i, \rm wide}\right)^2\right].
+\end{align}
 
+\subsection{Prior probability distributions}\label{sec:priors}
 The joint prior distribution for the hyperparameters, parameters, and data,

### 2017-05-04 doc/ms.tex
Copy of dnest4 Python stuff, so it's bundled

@@
 
+\section{Installation and usage}
+
+\subsection*{Windows}
+The entire repository, including C++ and Python source code, and a Windows
+executable, is hosted on Bitbucket. The URL for the git repository is
+
+{\tt https://bitbucket.org/eggplantbren/crystals}
+
+If you do not use git, the repository can simply be downloaded from
+
+{\tt https://bitbucket.org/eggplantbren/crystals/get/c78decf570f0.zip}
+
+and unzipped into any location on your computer.
+
 \bibliographystyle{plainnat}

### 2017-05-04 src/display.py

@@ -1,4 +1,4 @@
-import dnest4.classic as dn4
+import dnest_functions as dn4
 import numpy as np
 import matplotlib.pyplot as plt
 

### 2017-05-04 src/dnest_functions.py

@@ -1,0 +1,576 @@
+import copy
+import numpy as np
+import matplotlib.pyplot as plt
+
+def my_loadtxt(filename, single_precision=False, delimiter=" "):
+    """
+    Load quickly
+    """
+    # Open the file
+    f = open(filename, "r")
+
+    # Storage
+    results = []
+
+    # Rows and columns
+    nrow = 0
+    ncol = None
+
+    while(True):
+        # Read the line and split by whitespace
+        line = f.readline()
+        if delimiter != " ":
+            cells = line.split(delimiter)
+        else:
+            cells = line.split()
+
+        # Quit when you see a different number of columns
+        if ncol is not None and len(cells) != ncol:
+            break
+
+        # Non-comment lines
+        if cells[0][0] != "#":
+            # If it's the first one, get the number of columns
+            if ncol is None:
+                ncol = len(cells)
+
+            # Otherwise, include in results
+            if single_precision:
+                results.append(np.array([float(cell) for cell in cells],\
+                                                          dtype="float32"))
+            else:
+                results.append(np.array([float(cell) for cell in cells]))
+            nrow += 1
+
+    results = np.vstack(results)
+    return results
+
+def loadtxt_rows(filename, rows, single_precision=False):
+    """
+    Load only certain rows
+    """
+    # Open the file
+    f = open(filename, "r")
+
+    # Storage
+    results = {}
+
+    # Row number
+    i = 0
+
+    # Number of columns
+    ncol = None
+
+    while(True):
+        # Read the line and split by whitespace
+        line = f.readline()
+        cells = line.split()
+
+        # Quit when you see a different number of columns
+        if ncol is not None and len(cells) != ncol:
+            break
+
+        # Non-comment lines
+        if cells[0] != "#":
+            # If it's the first one, get the number of columns
+            if ncol is None:
[…]

### 2017-05-04 src/showresults.py

@@ -1,4 +1,4 @@
-import dnest4.classic as dn4
+import dnest_functions as dn4
 dn4.postprocess()
 
 import display

### 2017-05-04 src/display.py
Try proper DNest4 first

@@ -1,4 +1,10 @@
-import dnest_functions as dn4
+# Uses actual DNest4 stuff if it's installed - otherwise
+# uses dnest_functions.py as a fallback
+try:
+    import dnest4.classic as dn4
+except:
+    import dnest_functions as dn4
+
 import numpy as np
 import matplotlib.pyplot as plt
 

### 2017-05-04 src/showresults.py

@@ -1,4 +1,10 @@
-import dnest_functions as dn4
+# Uses actual DNest4 stuff if it's installed - otherwise
+# uses dnest_functions.py as a fallback
+try:
+    import dnest4.classic as dn4
+except:
+    import dnest_functions as dn4
+
 dn4.postprocess()
 
 import display

### 2017-05-05 doc/ms.tex
skewness

@@
 
-\subsection{Prior probability distributions}\label{sec:priors}
+\subsection{Derived properties}
+Some derived properties of the wide component are used to summarise its shape.
+We have included `skewness', which describes asymmetry, and `nongaussianity'
+which quantifies the departure from gaussianity. These are not the same
+concept, as the wide component could be non-gaussian while still being symmetric
+(and thus having zero skewness).
+
+Defining the normalised version of the wide component as
+\begin{align}
+g(\x) &= \frac{f_{\rm wide}(\x)}
+             {\int_{-\infty}^\infty f_{\rm wide}(\x') \, d\x'},
+\end{align}
+its center of mass by
+\begin{align}
+\textnormal{center} &= \int_{-\infty}^\infty g(\x) \, d\x,
+\end{align}
+and its width (second central moment) by
+\begin{align}
+\textnormal{width} &= \sqrt{\int_{-\infty}^\infty
+                            g(\x) \left(\x - \textnormal{center}\right)^2
+                            \, d\x},
+\end{align}
+the skewness is
+\begin{align}
+\textnormal{skewness} &= \int_{-\infty}^\infty
+                           g(\x) 
+                           \left(
+                             \frac{x - \textnormal{center}}{\textnormal{width}}
+                           \right)^3
+                         \, d\x.
+\end{align}
+
+\section{Prior probability distributions}\label{sec:priors}
 The joint prior distribution for the hyperparameters, parameters, and data,

### 2017-05-13 doc/figures.py
Fixed normalisation

@@ -46,8 +46,9 @@ end   = indices["wide[1000]"] + 1
 for i in range(0, 8):
     k = rng.randint(sample.shape[0])
     y = sample[k, start:end]
-    y /= y.max() + 0.00001
+    y /= y.max()# + 0.00001
     y *= np.exp(0.1*rng.randn())
+    
     plt.plot(x, y, "-", alpha=0.4)
     plt.xlabel("$x$")
     plt.ylabel("$y$")

### 2017-05-13 doc/ms.tex
Nongaussianity
Figure showing wide components

@@
 \includegraphics[width=0.7\textwidth]{figures/background.pdf}
-\caption{Three example functions showing possible shapes of the
+\caption{\it Three example functions showing possible shapes of the
 background component.\label{fig:background}}
@@
 
+Figure~\ref{fig:wide_component} shows several examples of plausible
+wide component shapes.
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[scale=0.7]{figures/wide_component.pdf}
+\caption{\it Some examples of plausible shapes for the wide component,
+generated from the prior distribution for its parameters
+(Section~\ref{sec:priors}). For visual purposes only, these
+were scaled to amplitudes $\sim$ 1. \label{fig:wide_component}}
+\end{figure}
+
 \subsection{Derived properties}
@@
 \end{align}
+
+To quantify the non-gaussianity, we construct a gaussian
+$g'(\x)$
+with the same center and width as $g(\x)$, and compute the
+Kullback-Leibler divergence from $g'$ to $g$:
+\begin{align}
+\textnormal{non-gaussianity} &= 
+    \int_{-\infty}^\infty g(\x) \log\left[\frac{g(\x)}{g'(\x)}\right] \, d\x.
+\end{align}
+where
+\begin{align}
+g'(\x) &= \frac{1}{\textnormal{width} \times \sqrt{2\pi}}
+            \exp\left[-\frac{1}{2\times (\textnormal{width})^2}
+                    \left(\x - \textnormal{center}\right)^2\right].
+\end{align}
+This is zero when $g(\x)$ is a gaussian.
 
@@
 \end{tabular}
-\caption{\label{tab:priors}}
+\caption{\it The prior distributions for all hyperparameters,
+parameters, and the data.\label{tab:priors}}
 \end{table}

### 2017-05-13 doc/figures.py

@@ -53,5 +53,7 @@ for i in range(0, 8):
     plt.xlabel("$x$")
     plt.ylabel("$y$")
     plt.title("Wide component")
+
+plt.savefig("figures/wide_component.pdf", bbox_inches="tight")
 plt.show()
 

### 2017-05-13 doc/figures.py
Don't show figures

@@ -33,7 +33,8 @@ plt.xlabel("$x$", fontsize=16)
 plt.ylabel("$y$", fontsize=16)
 plt.title("Background component")
 plt.savefig("figures/background.pdf", bbox_inches="tight")
-plt.show()
+print("figures/background.pdf done")
+plt.clf()
 
 # Load some prior samples
 sample = dn4.my_loadtxt("figures/sample.txt")
@@ -55,5 +56,7 @@ for i in range(0, 8):
     plt.title("Wide component")
 
 plt.savefig("figures/wide_component.pdf", bbox_inches="tight")
-plt.show()
+print("figures/wide_component.pdf done")
+plt.clf()
+
 

### 2017-05-14 doc/figures.py
Skewness figure

@@ -59,4 +59,34 @@ plt.savefig("figures/wide_component.pdf", bbox_inches="tight")
 print("figures/wide_component.pdf done")
 plt.clf()
 
+# Demonstrate skewness and nongaussianity
+def properties(f):
+    f /= f.sum()
+    com = np.sum(f*x)
+    wsq = np.sum(f*(x - com)**2)
+    w = np.sqrt(wsq)
+    skewness = np.sum(f*((x - com)/w)**3)
+    gaussian = np.exp(-0.5*(x - com)**2/wsq)
+    gaussian /= np.sum(gaussian)
+    KL = np.sum(f*np.log(f/gaussian))
+    return (skewness, KL)
+
+width = 3*np.ones(len(x))
+f0 = np.exp(-0.5*(x - 25)**2/width**2)
+
+width = 3 + 1.15*(x > 25)
+f1 = np.exp(-0.5*(x - 25)**2/width**2)
+
+width = 3 + 3.05*(x > 25)
+f2 = np.exp(-0.5*(x - 25)**2/width**2)
+
+plt.plot(x, f0/f0.max(), alpha=0.6, label="Skewness = 0.0")
+plt.plot(x, f1/f1.max(), alpha=0.6, label="Skewness = 0.25")
+plt.plot(x, f2/f2.max(), alpha=0.6, label="Skewness = 0.50")
+plt.xlabel("$x$")
+plt.ylabel("$y$")
+plt.title("Skewness")
+plt.legend(loc="upper left")
+plt.savefig("figures/skewness.pdf", bbox_inches="tight")
+print("figures/skewness.pdf done")
 

### 2017-05-15 doc/figures.py
non-gaussianity figure

@@ -89,4 +89,30 @@ plt.title("Skewness")
 plt.legend(loc="upper left")
 plt.savefig("figures/skewness.pdf", bbox_inches="tight")
 print("figures/skewness.pdf done")
+plt.close()
+
+nu = 100.0
+f0 = (1.0 + (x - 25.0)**2/nu/3**2)**(-0.5*(nu + 1.0))
+f0 /= f0.max()
+
+nu = 3.0
+f1 = (1.0 + (x - 25.0)**2/nu/3**2)**(-0.5*(nu + 1.0))
+f1 /= f1.max()
+
+nu = 1.0
+f2 = (1.0 + (x - 25.0)**2/nu/3**2)**(-0.5*(nu + 1.0))
+f2 /= f2.max()
+
+plt.figure(figsize=(9, 6))
+plt.plot(x, f0, alpha=0.6, label="non-gaussianity=0")
+plt.plot(x, f1, alpha=0.6, label="non-gaussianity=0.08")
+plt.plot(x, f2, alpha=0.6, label="non-gaussianity=0.15")
+plt.xlabel("$x$")
+plt.ylabel("$y$")
+plt.title("non-gaussianity")
+plt.legend(loc="upper left")
+plt.savefig("figures/nongaussianity.pdf", bbox_inches="tight")
+print("figures/nongaussianity.pdf done")
+
+
 

### 2017-05-15 doc/ms.tex
Skewness figure
The nongaussianity figure

@@
 $g'(\x)$
-with the same center and width as $g(\x)$, and compute the
+with the same centre, width, and total integral as $g(\x)$, and compute the
 Kullback-Leibler divergence from $g'$ to $g$:
@@
 \end{align}
-This is zero when $g(\x)$ is a gaussian.
+The Kullback-Leibler divergence is a standard way of quantifying how
+different one measure is from another \citep{knuth2012foundations}.
+It is zero when the two measures are the same, in our case,
+if $g(\x)$ is a gaussian.
+
+See Figures~\ref{fig:skewness} and~\ref{fig:nongaussianity} for example
+shapes for the wide component, and corresponding values of the skewness
+and nongaussianity.
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[scale=0.7]{figures/skewness.pdf}
+\caption{\it Examples of wide component shapes with different values
+for the skewness.\label{fig:skewness}}
+\end{figure}
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[scale=0.7]{figures/nongaussianity.pdf}
+\caption{\it Examples of wide component shapes with different values
+for the nongaussianity.\label{fig:nongaussianity}}
+\end{figure}
 

### 2017-06-11 doc/ms.tex
Fixed error in center of mass equation
MOre verbosity

@@
 \section{Bayesian inference}
-In the Bayesian framework, probability distributions are used to model
+In the Bayesian framework
+\citep{gregory2005bayesian, o2004kendall, sivia2006data},
+probability distributions are used to model
 states of uncertainty about the values of unknown quantities.
@@
 if we knew the parameters (and the prior information), as a function of
-the parameters.
+the parameters. This describes uncertainty about what data will be observed,
+but knowledge about the fact that the data has something to do with the
+parameters $\params$.
+The
+product of these two distributions is the joint prior
+\begin{align}
+p(\params, \data | \info) &= p(\params | \info)p(\data | \params, \info)
+\end{align}
+which describes uncertainty about the value of both the parameters and the
+dataset.
 
-The
-product of these is the joint prior
-\begin{align}
-p(\params, \data | \info) &= p(\params | \info)p(\data | \params, \info).
-\end{align}
 Once a specific observed dataset $\data_o$ is known, knowledge about
@@
 \end{align}
-which takes the specific data $\data_0$ into account.
+which takes the specific data $\data_0$ into account. With a specific dataset
+$\data_0$ being specified, $p(\data_o | \params,\info)$ is a function of the
+unknown parameters $\params$ and is called the {\em likelihood function}.
+The posterior distribution is therefore proportional to the prior distribution
+times the likelihood function.
+
+When $\params$ is a large collection of parameters, the posterior distribution
+is a probability distribution over a high-dimensional parameter space. To
+represent it computationally, Markov Chain Monte Carlo (MCMC) methods are
+often used to generate samples from the parameter space, according to the
+posterior distribution. The output is a collection of plausible values of the
+parameters which can be used to approximate posterior probabilities.
+For example, the probability some parameter $\theta$ is greater than 3.4
+is approximately the proportion of the posterior samples that have
+$\theta > 3.4$.
+
+We compute the posterior distribution for the unknown parameters of the
+components, using \citep{dnest4} to generate samples from that posterior
+distribution.
 
@@
 \end{align}
-
-We compute the posterior distribution for the unknown parameters of the
-components, using \citep{dnest4} to generate samples from that posterior
-distribution. The output is a collection of plausible values of the
-parameters which can be used to approximate posterior probabilities.
-For example, the probability some parameter $\theta$ is greater than 3.4
-is approximately the proportion of the posterior samples that have
-$\theta > 3.4$.
 
@@
 \begin{align}
-\textnormal{center} &= \int_{-\infty}^\infty g(\x) \, d\x,
+\textnormal{center} &= \int_{-\infty}^\infty x g(\x) \, d\x,
 \end{align}

### 2017-06-11 doc/ms.tex
Editing the appendix

@@
 
-\subsection*{Windows}
 The entire repository, including C++ and Python source code, and a Windows
 executable, is hosted on Bitbucket. The URL for the git repository is
+For most users, we suggest simply using the pre-compiled Windows executable
+{\tt main.exe} from the repository.
 
-{\tt https://bitbucket.org/eggplantbren/crystals}
-
-If you do not use git, the repository can simply be downloaded from
+The repository can simply be downloaded from
 
@@
 
-and unzipped into any location on your computer.
+and unzipped into any location on your computer. To execute the program,
+run {\tt main.exe}. You will be prompted to enter the filename of the text
+file containing the data.
 

### 2017-09-01 src/display.py
Updated axis labels

@@ -98,8 +98,8 @@ def display():
             plt.plot(data[:,0], bg, "y", linewidth=2, alpha=0.1)
             plt.ylim(0)
 
-    plt.xlabel("$x$", fontsize=16)
-    plt.ylabel("$y$", fontsize=16)
+    plt.xlabel("$2\\theta$ (degrees)", fontsize=16)
+    plt.ylabel("$Intensity$", fontsize=16)
     plt.show()
 
     # Remove any nans before plotting
@@ -108,8 +108,8 @@ def display():
 
     plt.hist(wide_fraction, 100, color=[0.8, 0.8, 0.8])
     plt.xlim([0, 1])
-    plt.xlabel("(wide component flux)/(total spike flux + wide component flux)")
-    print("Wide fraction = {a} +- {b}".format(a=wide_fraction.mean(),\
+    plt.xlabel("(amporph)/(amorph + crystal)")
+    print("(amporph)/(amorph + crystal) = {a} +- {b}".format(a=wide_fraction.mean(),\
            b=wide_fraction.std()))
     plt.show()
 

### 2017-09-06 doc/figures.py
Wide -> Amorphous

@@ -53,7 +53,7 @@ for i in range(0, 8):
     plt.plot(x, y, "-", alpha=0.4)
     plt.xlabel("$x$")
     plt.ylabel("$y$")
-    plt.title("Wide component")
+    plt.title("Amorphous component")
 
 plt.savefig("figures/wide_component.pdf", bbox_inches="tight")
 print("figures/wide_component.pdf done")

### 2017-09-07 doc/figures.py
Consistent font sizes

@@ -29,8 +29,8 @@ for i in range(0, 3):
     # Plot the background component shape
     plt.plot(xx, yy, "ko-", alpha=0.2)
 
-plt.xlabel("$x$", fontsize=16)
-plt.ylabel("$y$", fontsize=16)
+plt.xlabel("$x$")
+plt.ylabel("$y$")
 plt.title("Background component")
 plt.savefig("figures/background.pdf", bbox_inches="tight")
 print("figures/background.pdf done")
@@ -92,24 +92,25 @@ print("figures/skewness.pdf done")
 plt.close()
 
 nu = 100.0
-f0 = (1.0 + (x - 25.0)**2/nu/3**2)**(-0.5*(nu + 1.0))
+f0 = (1.0 + (x - 30.0)**2/nu/3**2)**(-0.5*(nu + 1.0))
 f0 /= f0.max()
 
 nu = 3.0
-f1 = (1.0 + (x - 25.0)**2/nu/3**2)**(-0.5*(nu + 1.0))
+f1 = (1.0 + (x - 30.0)**2/nu/3**2)**(-0.5*(nu + 1.0))
 f1 /= f1.max()
 
 nu = 1.0
-f2 = (1.0 + (x - 25.0)**2/nu/3**2)**(-0.5*(nu + 1.0))
+f2 = (1.0 + (x - 30.0)**2/nu/3**2)**(-0.5*(nu + 1.0))
 f2 /= f2.max()
 
-plt.figure(figsize=(9, 6))
-plt.plot(x, f0, alpha=0.6, label="non-gaussianity=0")
-plt.plot(x, f1, alpha=0.6, label="non-gaussianity=0.08")
-plt.plot(x, f2, alpha=0.6, label="non-gaussianity=0.15")
+#plt.figure(figsize=(9, 6))
+plt.plot(x, f0, alpha=0.6, label="Non-gaussianity=0")
+plt.plot(x, f1, alpha=0.6, label="Non-gaussianity=0.08")
+plt.plot(x, f2, alpha=0.6, label="Non-gaussianity=0.15")
 plt.xlabel("$x$")
 plt.ylabel("$y$")
-plt.title("non-gaussianity")
+plt.xlim([-5, 55])
+plt.title("Non-gaussianity")
 plt.legend(loc="upper left")
 plt.savefig("figures/nongaussianity.pdf", bbox_inches="tight")
 print("figures/nongaussianity.pdf done")

### 2017-09-08 doc/ms.tex
Elsevier format
Version downloaded from Overleaf

@@
-\documentclass[a4paper, 12pt]{article}
+\documentclass[review]{elsarticle}
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{color}
@@
 \usepackage{graphicx}
-\usepackage[left=2cm, right=2cm, bottom=3cm, top=2cm]{geometry}
+% \usepackage[left=2cm, right=2cm, bottom=3cm, top=2cm]{geometry}
 \usepackage{microtype}
@@
 \usepackage{pdflscape}
+\usepackage{gensymb} % For degree symbol
+\usepackage{lineno,hyperref}
+\modulolinenumbers[5]
+
+\journal{Computers and Geosciences}
+
+\setcitestyle{authoryear,open={(},close={)}}
 
@@
 
-\title{Crystals}
-\author{Brendon J. Brewer}
-\date{}
+% http://support.river-valley.com/wiki/index.php?title=Elsarticle.cls
 
 \begin{document}
-\maketitle
 
-%\abstract{\noindent Abstract}
+\begin{frontmatter}
 
-% Need this after the abstract
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{8pt}
+\title{AMORPH: A statistical program for characterizing amorphous materials by X-ray diffraction}
 
-\section{Bayesian inference}
-In the Bayesian framework
-\citep{gregory2005bayesian, o2004kendall, sivia2006data},
+\author[rowe]{Michael C. Rowe\corref{cor1}}
+\ead{michael.rowe@auckland.ac.nz}
+\cortext[cor1]{Corresponding author}
+\address[rowe]{School of Environment, The University of Auckland, Private Bag 92019, Auckland 1142, New Zealand}
+
+\author[brewer]{Brendon J. Brewer}
+\ead{bj.brewer@auckland.ac.nz}
+\address[brewer]{Department of Statistics, The University of Auckland, Private Bag 92019, Auckland 1142, New Zealand}
+
+\begin{abstract}
+AMORPH utilizes a new Bayesian statistical approach to interpreting X-ray diffraction results of partially amorphous materials. AMORPH fits X-ray diffraction patterns with a mixture of gaussians,
+which allows us to simultaneously infer all of the model parameters and quantify their
+uncertainties.
+The program simulates background patterns previously manually applied, providing reproducible results, and significantly reducing inter- and intra-user biases.
+This approach allows for the quantification of amorphous and crystalline materials and for the characterization of the amorphous component, including parameters of centre of mass, width, skewness, and nongaussianity of the amorphous component.
+Results demonstrate the applicability of this program for calculating amorphous contents of volcanic materials and independently modeling their properties in compositionally variable materials.
+\end{abstract}
+
+\begin{keyword}
+  Amorphous materials \sep X-ray diffraction \sep Bayesian inference \sep Markov Chain Monte Carlo \sep Mixture models \sep crystallinity
+\end{keyword}
+
+\end{frontmatter}
+
+\linenumbers
+
+\section{Introduction}
+Quantification and characterization of amorphous materials is an important area of research in
+terrestrial and planetary sciences \citep[e.g.,][]{schmidt2009, dehouck2014, wall2014, morris2016, zorn2017}.
+Powder X-ray diffraction (XRD), although primarily a tool for characterization of crystalline materials,
+also provides a means to investigate amorphous materials \citep[e.g.,][]{rowe2012, achilles2013}. X-ray diffraction of rocks/soils containing amorphous materials (e.g. volcanic
+glass, amorphous silica, and organics) produces a diffraction pattern which is a combination of the
+crystalline and amorphous phases. Recent studies have debated the nature of amorphous materials in
+planetary sciences \citep[e.g.][]{schmidt2009, dehouck2014}, and a growing number of
+applications to terrestrial volcanics have presented an avenue of expansion for XRD methodologies
+for amorphous material analysis \citep{ellis2015, kanakiya2017, zorn2017}.
+Here we present a new statistical approach and computer program, AMORPH, which utilizes Bayesian inference to determine
+the crystallinity, and other characteristics, of partially amorphous materials from X-ray diffraction patterns, as well as the uncertainty in these quantities.
+
+Several techniques currently exist for the quantification of amorphous materials in heterogeneous (amorphous + crystalline) materials. However, these
+each present their own limitations. The two most commonly used approaches in recent literature
+either rely on a calibration curve for different crystallinities \citep[calibration method;][]{ rowe2012} or a combined Rietveld-Reference Intensity Ratio (RIR) method \citep{gualtieri1996, gualtieri2000}.
+While the combined Rietveld-RIR procedure allows for quantification of both the amorphous
+and crystalline materials, for proper characterization it requires a library of pure components,
+spiked with a reference material (e.g. corundum). Sample spiking presents additional complications
+as it reduces the observed intensity of the amorphous material thus reducing the accuracy of
+pattern matching to determine abundances. An Excel based adaptation of this pattern matching
+technique \citep[FULLPAT;][]{chipera2002} is routinely utilized for characterizing remotely acquired X-ray
+diffraction patterns, such as those recovered from the Mars Science Laboratory (MSL) Curiosity rover \citep{bish2013}. The problem faced by RIR approaches is they are based on modeling of an observed 
+diffraction pattern compared to library measurements, and therefore do not allow an independent
+assessment of the characteristics of the amorphous material. In comparison, the calibration method
+focuses solely on the quantification of the amorphous material, utilizing the integrated counts
+associated with the amorphous and crystalline componentry.
+
+\begin{align}
+C_{\rm m} \% = [C_{\rm pa}]/[A_{\rm pa} + C_{\rm pa}] \times 100
+\end{align}
+
+where $C_{\rm m}$ is the measured crystallinity, and $C_{\rm pa}$ and $A_{\rm pa}$ are the integrated peak areas for the
+crystalline and amorphous components, respectively (see Figure 5 of \citet{wall2014}). 
+While its simplicity has made it advantageous, the calibration methodology has been limited in that
+the quality of the results largely depend on the ability of the user to manually assign
+backgrounds for count integration, resulting in potentially large inter-user variations. Although the
+point of the calibration curve is to reduce the inter-user variation, determinations still tend to
+be associated with moderate user error (3-5 vol \%). As every user must independently practice
+background fittings using the calibration method, the results ultimately are only as good as the
+user’s ability to manually model the data. This technique, while not constrained by a
+necessary library of diffraction patterns, also does not characterize the XRD properties of the
+amorphous material. This study presents a new approach using Bayesian inference to automate
+background determination for the calibration method, removing inter-user and intra-user biases in
+the determination of crystallinity and peak characteristics for amorphous/partially amorphous
+materials, regardless of what they are.
+
+This paper is structured as follows. In Section~\ref{sec:bayes}, we outline the Bayesian
+framework of inference in general terms, and in Section~\ref{sec:model} we describe the
+specific model assumptions made by AMORPH (i.e., the mixture of gaussians), and
+Section~\ref{sec:priors} lists the probabilistic assumptions (i.e., the prior distributions) of the program. Section~\ref{sec:applications} presents some
+example applications and comparisons with standard techniques, and finally we conclude
+in Section~\ref{sec:conclusions}. Basic instructions for installation and use of AMORPH are given in~\ref{sec:program}.
+
+\section{Bayesian inference}\label{sec:bayes}
+AMORPH is founded upon Bayesian inference
+\citep{gregory2005bayesian, o2004kendall, sivia2006data}, where
 probability distributions are used to model
 states of uncertainty about the values of unknown quantities.
-Typically, one starts with the ``prior distribution''
+Bayesian inference is often implemented computationally with
+Markov Chain Monte Carlo (MCMC) methods \citep{mackay2003}, and AMORPH is no exception.
+Loosely speaking, the goal is to fit a dataset with a model by exploring
+those regions of the model's parameter space that are consistent with the
+data. The model assumptions are described in detail in
+Section~\ref{sec:model}, and the computation is performed with
+Diffusive Nested Sampling, an advanced MCMC method \citep{dns, dnest4}.
+
+Typically, one starts with the `prior distribution'
 for unknown quantities (`parameters') $\params$, written $p(\params | \info)$
 (the probability distribution for $\params$ {\em given} prior information
-$\info$). This is supplemented with $p(\data | \params, \info)$,
+$\info$). Throughout this paper, $\params$ is used to denote a
+collection of unknown parameters, whereas $2\theta$ (with the factor of two, and the
+non-bold $\theta$) is later used to
+refer to the incident angle plus reflected angle, which is the
+standard notation in X-ray diffraction.
+
+The prior describes an initial state of uncertainty about the
+parameter values, and is often a wide probability distribution.
+The prior is supplemented with $p(\data | \params, \info)$, sometimes
+called a `sampling distribution',
 which describes the uncertainty about the data which we {\em would} have,
@@
 the parameters. This describes uncertainty about what data will be observed,
-but knowledge about the fact that the data has something to do with the
-parameters $\params$.
+but encodes some knowledge about the fact that the data has something to do with the
+parameters $\params$. For example, $p(\data | \params, \info)$ often encodes
+the assumption that the data has noise in it, and the specific noise values
+are unknown a priori.
+
 The
@@
 which describes uncertainty about the value of both the parameters and the
-dataset.
-
-Once a specific observed dataset $\data_o$ is known, knowledge about
-$\params$ is updated from the prior to the posterior
+dataset. Once a specific observed dataset $\data_{\rm obs}$ is known, knowledge about
+$\params$ is updated from the prior $p(\params | \info)$ to the posterior
 \begin{align}
-p(\params | \data_0, \info) &\propto
-    p(\params | \info)p(\data_o | \params, \info).
+p(\params | \data_{\rm obs}, \info) &\propto
+    p(\params | \info)p(\data_{\rm obs} | \params, \info).
 \end{align}
-which takes the specific data $\data_0$ into account. With a specific dataset
-$\data_0$ being specified, $p(\data_o | \params,\info)$ is a function of the
-unknown parameters $\params$ and is called the {\em likelihood function}.
+which takes the specific data $\data_{\rm obs}$ into account. With the specific dataset,
+$p(\data_{\rm obs} | \params,\info)$ becomes a function of the
+unknown parameters $\params$ only and is called the {\em likelihood function}.
 The posterior distribution is therefore proportional to the prior distribution
@@
 parameters which can be used to approximate posterior probabilities.
-For example, the probability some parameter $\theta$ is greater than 3.4
+For example, the probability some parameter $\phi$ is greater than 3.4
 is approximately the proportion of the posterior samples that have
-$\theta > 3.4$.
+$\phi > 3.4$. While many data analysis procedures work by finding the
+best fitting values of the parameters (i.e., a maximum likelihood estimate),
+Bayesian inference via MCMC involves finding a sampling of plausible values
+for the parameters.
 
-We compute the posterior distribution for the unknown parameters of the
-components, using \citep{dnest4} to generate samples from that posterior
-distribution.
+\section{The model curve}\label{sec:model}
 
-\section{The model curve}
-The algorithm fits the data with a model curve
-$f(\x)$, defined between $\x=\x_{\rm min}$ and $\x=\x_{\rm max}$,
-which is assumed to be a sum of the following components:
+AMORPH fits the data with a model curve defined over some range from
+$x_{\rm min}$ to $x_{\rm max}$ (Fig.~\ref{fig:example_data2}). Throughout this section, we use
+$x$ as a synonym for $2\theta$, since the problem is basically a curve-fitting
+problem where $x$ and $y$ are typically used.
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=.7\textwidth]{figures/example_data2.jpg}
+\caption{\it Powder X-ray diffraction results for a basalt standard with 70\% amorphous material.
+Annotation corresponds to the model components: background, amorphous (wide) component, and crystalline component (narrow peaks).\label{fig:example_data2}}
+\end{figure}
+
+The model curve is a function $y = f(\x; \params)$,
+which is assumed to be a sum of the following components identified in Figure~\ref{fig:example_data2}:
 \begin{enumerate}
 \item A background component, described by parameters $\params_{\rm bg}$
-\item A wide component, described by parameters $\params_{\rm wide}$
-\item Narrow spikes, described by parameters $\params_{\rm narrow}$
+\item An amorphous component, described by parameters $\params_{\rm amorph}$
+\item A crystalline component, described by parameters $\params_{\rm xtal}$
 \end{enumerate}
-
 The total model function $f_{\rm tot}(x)$ is therefore
@@
 f_{\rm tot}(\x) &= f_{\rm bg}(\x; \params_{\rm bg})
-       + f_{\rm wide}(\x; \params_{\rm wide})
-       + f_{\rm narrow}(\x; \params_{\rm narrow})
+       + f_{\rm amorph}(\x; \params_{\rm amorph})
+       + f_{\rm xtal}(\x; \params_{\rm xtal})
 \end{align}
-and the crystallinity is defined as the proportion of the area in the
-wide and narrow components which is accounted for by the wide component:
+and the measured amorphous content ($A_{\rm m}$ or $1-C_{\rm m}$) is defined, similar to Equation 1, as the proportion of the area in the
+amorphous and crystalline components which is accounted for by the amorphous component:
 \begin{align}
-C &= \frac{\int_{\x_{\rm min}}^{\x_{\rm max}}
-            f_{\rm wide}(\x; \params_{\rm wide}) \, d\x}
+{A_{\rm m} = 1-C_{\rm m}} &= \frac{\int_{\x_{\rm min}}^{\x_{\rm max}}
+            f_{\rm amorph}(\x; \params_{\rm amorph}) \, d\x}
           {\int_{\x_{\rm min}}^{\x_{\rm max}}
-            \left[f_{\rm wide}(\x; \params_{\rm wide})
-            + f_{\rm narrow}(\x; \params_{\rm narrow}) \right] \, d\x}
+            f_{\rm amorph}(\x; \params_{\rm amorph}) \, dx
+            + \int_{\x_{\rm min}}^{\x_{\rm max}} f_{\rm xtal}(\x; \params_{\rm xtal}) \, d\x} \label{eqn:crystallinity}
 \end{align}
 
-From each sample,
-it is possible to compute the crystallinity, making it straightforward to
-calculate the marginal posterior distribution of the crystallinity, describing
+From each sample from the posterior distribution for all of the parameters,
+it is possible to compute the proportion of amorphous material, making it straightforward to
+calculate the marginal posterior distribution of the amorphous content, describing
 the remaining uncertainty about its value in the light of the data.
-
-In the following subsections, I describe the specific parameterisations of the
-three parts of the model curve.
+In the following subsections, we describe the specific parameterizations of the
+three parts of the model curve, that is, the functional forms of
+$f_{\rm bg}$, $f_{\rm amorph}$, and $f_{\rm xtal}$.
 
 \subsection{The background component}
-The background component is assumed to be piecewise linear
-{\todo between $\x=\x_{\rm min}$ and $\x=10$, between $\x=10$ and $\x=40$,
-and between $\x=40$ and $\x=\x_{\rm max}$}.
-We parameterise the background component using
+The background component is assumed to be piecewise linear, with constant
+slope
+{between $\x=\x_{\rm min}$ and $\x=10$\degree, between $\x=10$\degree~and $\x=40$\degree,
+and between $\x=40$\degree~and $\x=\x_{\rm max}$}.
+These values are selected to maintain a linear background over the interval from 10--40\degree, the predominant range of amorphous materials analyzed using a Cu K$\alpha$ x-ray source, while still allowing for some flexibility in the structure of the background outside this range.
+
+We parameterized the background component using
 five parameters. Firstly, there is a mean level $b$, which sets the
 typical amplitude of the background component. There are also four parameters
-$n_1, ..., n_4$, which set the positions of four control points:
+$n_1, ..., n_4$, which determine the offset of four `control points'
+from the mean level. The positions of the four control points are then:
 \begin{align}
@@
 \left(40, be^{n_3}\right) \nonumber\\
-\left(\x_{\rm max}, be^{n_4}\right) \nonumber
+\left(\x_{\rm max}, be^{n_4}\right). \nonumber
 \end{align}
 The value of the background component at any other point is found by
-linear interpolation. Figure~\ref{fig:background} shows three example
-background curves generated from this parameterisation, with typical
+linear interpolation of these control points.
+Figure~\ref{fig:background} shows three example
+background curves generated from this parameterization, with typical
 values of the parameters.
 
-\begin{figure}
+\begin{figure}[!ht]
 \centering
-\includegraphics[width=0.7\textwidth]{figures/background.pdf}
+\includegraphics[width=0.6\textwidth]{figures/background.pdf}
 \caption{\it Three example functions showing possible shapes of the
@@
 
-\subsection{The narrow spikes}
-I assumed that there is some number $N_{\rm narrow}$ of narrow spikes,
-each of which has a gaussian shape. The $i$th narrow spike is parameterised
-by its central position $X_{i, \rm narrow}$,
-amplitude $A_{i, \rm narrow}$, and width $A_{i, \rm narrow}$.
-The parameter vector for narrow spikes is
+\subsection{The crystalline component}
+We assume that there is some number $N_{\rm xtal}$ of narrow peaks,
+each of which has a gaussian shape. The $i$th peak is parameterized
+by its central position $X_{i, \rm xtal}$,
+amplitude $A_{i, \rm xtal}$, and width $W_{i, \rm xtal}$.
+The parameter vector for crystalline component is
 \begin{align}
-\params_{\rm narrow} &=
-  \{N_{\rm narrow}, X_1, X_2, ..., X_{N, \rm narrow},
-    A_1, A_2, ..., A_{N, \rm narrow},
-    W_1, W_2, ..., W_{N, \rm narrow}\}
+\params_{\rm xtal} &=
+  \{N_{\rm xtal}, X_1, X_2, ..., X_{N, \rm xtal},
+    A_1, A_2, ..., A_{N, \rm xtal},
+    W_1, W_2, ..., W_{N, \rm xtal}\}
 \end{align}
 and the total shape contributed by the
-narrow spikes is a sum of the $N_{\rm narrow}$ gaussians:
+crystalline component is a sum of the $N_{\rm xtal}$ gaussians:
 \begin{align}
-f_{\rm narrow}(\x; \params_{\rm narrow}) &=
-    \sum_{i=1}^{N_{\rm narrow}} A_{i, \rm narrow}
- \exp\left[-\frac{1}{2W_{i, \rm narrow}^2}\left(x - X_{i, \rm narrow}\right)^2\right].
+f_{\rm xtal}(\x; \params_{\rm xtal}) &=
+    \sum_{i=1}^{N_{\rm xtal}} A_{i, \rm xtal}
+ \exp\left[-\frac{1}{2W_{i, \rm xtal}^2}\left(x - X_{i, \rm xtal}\right)^2\right].
 \end{align}
 
-\subsection{The wide component}
-The wide component is also composed of gaussians, like the narrow spikes.
+\subsection{The amorphous component}
+The amorphous component is also composed of gaussians, like the crystalline component.
 In fact, the only real difference is the prior distributions for the positions,
@@
 
-There is some number $N_{\rm wide}$ of gaussians in the wide component.
-The $i$th gaussian is parameterised
-by its central position $X_{i, \rm wide}$,
-amplitude $A_{i, \rm wide}$, and width $A_{i, \rm wide}$.
+There is some number $N_{\rm amorph}$ of gaussians in the amorphous component.
+The $i$th gaussian is parameterized
+by its central position $X_{i, \rm amorph}$,
+amplitude $A_{i, \rm amorph}$, and width $W_{i, \rm amorph}$.
 The parameter vector for wide gaussians is
 \begin{align}
-\params_{\rm wide} &=
-  \{N_{\rm wide}, X_1, X_2, ..., X_{N, \rm wide},
-    A_1, A_2, ..., A_{N, \rm wide},
-    W_1, W_2, ..., W_{N, \rm wide}\}
+\params_{\rm amorph} &=
+  \{N_{\rm amorph}, X_1, X_2, ..., X_{N, \rm amorph},
+    A_1, A_2, ..., A_{N, \rm amorph},
+    W_1, W_2, ..., W_{N, \rm amorph}\}
 \end{align}
 and the total shape contributed by the
-wide component is a sum of the $N_{\rm wide}$ gaussians:
+amorphous component is a sum of the $N_{\rm amorph}$ gaussians:
 \begin{align}
-f_{\rm wide}(\x; \params_{\rm wide}) &=
-    \sum_{i=1}^{N_{\rm wide}} A_{i, \rm wide}
- \exp\left[-\frac{1}{2W_{i, \rm wide}^2}\left(x - X_{i, \rm wide}\right)^2\right].
+f_{\rm amorph}(\x; \params_{\rm amorph}) &=
+    \sum_{i=1}^{N_{\rm amorph}} A_{i, \rm amorph}
+ \exp\left[-\frac{1}{2W_{i, \rm amorph}^2}\left(x - X_{i, \rm amorph}\right)^2\right].
 \end{align}
 
-Figure~\ref{fig:wide_component} shows several examples of plausible
-wide component shapes.
+Figure~\ref{fig:wide_component} shows several examples of plausible amorphous component shapes. The prior distributions (specified in
+Section~\ref{sec:priors})
+constrain the central positions of these gaussians much
+more strongly than the central positions of the narrow peaks (crystalline component), which is how AMORPH is able to distinguish the crystalline
+component from the amorphous component.
 
@@
 \includegraphics[scale=0.7]{figures/wide_component.pdf}
-\caption{\it Some examples of plausible shapes for the wide component,
+\caption{\it Some examples of plausible shapes for the amorphous component,
 generated from the prior distribution for its parameters
@@
 \subsection{Derived properties}
-Some derived properties of the wide component are used to summarise its shape.
+Some derived properties of the amorphous component are used to summarize its shape.
 We have included `skewness', which describes asymmetry, and `nongaussianity'
 which quantifies the departure from gaussianity. These are not the same
-concept, as the wide component could be non-gaussian while still being symmetric
+concept, as the amorphous component could be non-gaussian while still being symmetric
 (and thus having zero skewness).
 
-Defining the normalised version of the wide component as
+Defining the normalized version of the amorphous component as
 \begin{align}
-g(\x) &= \frac{f_{\rm wide}(\x)}
-             {\int_{-\infty}^\infty f_{\rm wide}(\x') \, d\x'},
+g(\x) &= \frac{f_{\rm amorph}(\x)}
+             {\int_{-\infty}^\infty f_{\rm amorph}(\x') \, d\x'},
 \end{align}
@@
 \end{align}
-and its width (second central moment) by
+and its width (second central moment, perhaps better interpreted
+as a half-width) by
 \begin{align}
@@
 The Kullback-Leibler divergence is a standard way of quantifying how
-different one measure is from another \citep{knuth2012foundations}.
+different one `measure' ($\equiv$ density function, for our purposes here) is from another \citep{knuth2012foundations}.
 It is zero when the two measures are the same, in our case,
-if $g(\x)$ is a gaussian.
-
-See Figures~\ref{fig:skewness} and~\ref{fig:nongaussianity} for example
-shapes for the wide component, and corresponding values of the skewness
+if $g(\x)$ is a gaussian. See Figures~\ref{fig:skewness} and~\ref{fig:nongaussianity} for example
+shapes for the amorphous component, and corresponding values of the skewness
 and nongaussianity.
@@
 \centering
-\includegraphics[scale=0.7]{figures/skewness.pdf}
-\caption{\it Examples of wide component shapes with different values
+\includegraphics[width=0.6\textwidth]{figures/skewness.pdf}
+\caption{\it Examples of amorphous component shapes with different values
 for the skewness.\label{fig:skewness}}
@@
 \centering
-\includegraphics[scale=0.7]{figures/nongaussianity.pdf}
-\caption{\it Examples of wide component shapes with different values
+\includegraphics[width=0.6\textwidth]{figures/nongaussianity.pdf}
+\caption{\it Examples of amorphous component shapes with different values
 for the nongaussianity.\label{fig:nongaussianity}}
@@
 \section{Prior probability distributions}\label{sec:priors}
-The joint prior distribution for the hyperparameters, parameters, and data,
-is written $p(\hypers, \params, \data | \info)$, and is usually factorised
+The joint prior distribution for all of the unknown parameters and
+the data must be specified.
+When the number of parameters is large, this is often done `hierarchically', by
+specifying the priors conditional on some `hyperparameters', and then assigning
+priors to the hyperparameters.
+
+The joint prior distribution for
+the hyperparameters, parameters, and data,
+is written $p(\hypers, \params, \data | \info)$, and is usually factorized
 in this way:
@@
 step assumes that knowing the parameters would make the hyperparameters
-irrelevant when predicting what data would be observed.
+irrelevant when predicting what data would be observed. In the AMORPH model,
+we introduced hyperparameters describing the typical amplitude
+and width of the gaussians for the amorphous component, and
+the degree to which the actual amplitudes and widths are scattered
+around that typical value. This was also done for the crystalline
+component.
 
-All of the assumptions are given in Table~\ref{tab:priors}.
+A directed acyclic graph, also known as a probabilistic graphical model (PGM),
+which depicts the dependence structure of the entire joint
+prior probability distribution,
+is shown in Figure~\ref{fig:pgm-edited}. The prior distributions themselves
+are specified in Table~\ref{tab:priors}.
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[scale=0.7]{figures/pgm-edited.pdf}
+\caption{The PGM. The hyperparameters controlling the prior distributions
+for the amorphous and crystalline gaussians are denoted collectively
+by $\boldsymbol{\alpha}_{\rm amorph}$ and
+$\boldsymbol{\alpha}_{\rm xtal}$ respectively.
+The noise-related parameters are collected under $\boldsymbol{\sigma}$.
+The data is $\boldsymbol{y}$.\label{fig:pgm-edited}}
+\end{figure}
 
@@
 {\em Hyperparameters} & &\\
-$N_{\rm narrow}$   &   Number of narrow spikes    &  $\propto 1/(N_{\rm narrow}+1)$, $N_{\rm narrow} \in \{0, 1, ..., 300\}$ \\
-$N_{\rm wide}$   &   Number of wide spikes    &  $\propto 1/(N_{\rm wide}+1)$,
-$N_{\rm wide} \in \{0, 1, ..., 300\}$ \\
-$a_{\ln A, \rm narrow}$ & Typical log-amplitude of narrow spikes & Laplace(0, 5)\\
-$a_{\ln A, \rm wide}$ & Typical log-amplitude of wide spikes & Laplace(0, 5)\\
-$b_{\ln A, \rm narrow}$ & Diversity of log-amplitude of narrow spikes & Uniform(0, 5)\\
-$b_{\ln A, \rm wide}$ & Diversity of log-amplitude of wide spikes & Uniform(0, 5)\\
-$a_{\ln W, \rm narrow}$ & Typical log-width of narrow spikes & LogUniform$(10^{-3}x_{\rm range}, 0.05x_{\rm range})$\\
-$a_{\ln W, \rm wide}$ & Typical log-width of wide spikes & LogUniform$(0.05x_{\rm range}, x_{\rm range})$\\
-$b_{\ln W, \rm narrow}$ & Diversity of log-width of narrow spikes & Uniform(0, 0.1)\\
-$b_{\ln W, \rm wide}$ & Diversity of log-width of wide spikes & Uniform(0, 0.2)\\
+$N_{\rm xtal}$   &   Number of narrow peaks    &  $\propto 1/(N_{\rm xtal}+1)$, $N_{\rm xtal} \in \{0, 1, ..., 300\}$ \\
+$N_{\rm amorph}$   &   Number of wide peaks    &  $\propto 1/(N_{\rm amorph}+1)$,
+$N_{\rm amorph} \in \{0, 1, ..., 300\}$ \\
+$a_{\ln A, \rm xtal}$ & Typical log-amplitude of narrow peaks & Laplace(0, 5)\\
+$a_{\ln A, \rm amorph}$ & Typical log-amplitude of wide peaks & Laplace(0, 5)\\
+$b_{\ln A, \rm xtal}$ & Diversity of log-amplitude of narrow peaks & Uniform(0, 5)\\
+$b_{\ln A, \rm amorph}$ & Diversity of log-amplitude of wide peaks & Uniform(0, 5)\\
+$a_{\ln W, \rm xtal}$ & Typical log-width of narrow peaks & LogUniform$(10^{-3}x_{\rm range}, 0.05x_{\rm range})$\\
+$a_{\ln W, \rm amorph}$ & Typical log-width of wide peaks & LogUniform$(0.05x_{\rm range}, x_{\rm range})$\\
+$b_{\ln W, \rm xtal}$ & Diversity of log-width of narrow peaks & Uniform(0, 0.1)\\
+$b_{\ln W, \rm amorph}$ & Diversity of log-width of wide peaks & Uniform(0, 0.2)\\
 \hline
@@
 $b$       & Mean background level       & $\ln b \sim $ Laplace$(0, 5)$\\
-$\{n_1, ..., n_4\}$  & Background deviations & iid Normal$(0,1)$\\
+$\{n_1, ..., n_4\}$  & Background deviations & independent Normal$(0,1)$\\
 \hline
-{\em Narrow spike parameters}&&\\
-$X_{i, \rm narrow}$ & Positions of narrow spikes &
+{\em Crystalline parameters}&&\\
+$X_{i, \rm xtal}$ & Positions of narrow peaks &
                             Uniform$(x_{\rm min}, x_{\rm max})$ \\
-$A_{i, \rm narrow}$ & Amplitudes of narrow spikes &
- $\ln A_{i, \rm narrow} \sim \textnormal{Laplace}\left(a_{\ln A, \rm narrow}, b_{\ln A, \rm narrow}\right)$ \\
-$W_{i, \rm narrow}$ & Widths of narrow spikes &
- $\ln W_{i, \rm narrow} \sim \textnormal{Laplace}\left(a_{\ln W, \rm narrow}, b_{\ln W, \rm narrow}\right)$ \\
+$A_{i, \rm xtal}$ & Amplitudes of narrow peaks &
+ $\ln A_{i, \rm xtal} \sim \textnormal{Laplace}\left(a_{\ln A, \rm xtal}, b_{\ln A, \rm xtal}\right)$ \\
+$W_{i, \rm xtal}$ & Widths of narrow peaks &
+ $\ln W_{i, \rm xtal} \sim \textnormal{Laplace}\left(a_{\ln W, \rm xtal}, b_{\ln W, \rm xtal}\right)$ \\
 \hline
-{\em Wide spike parameters}&&\\
-$X_{i, \rm wide}$ & Positions of wide spikes &
+{\em Amorphous parameters}&&\\
+$X_{i, \rm amorph}$ & Positions of wide peaks &
           Uniform$(x_{\rm min}+0.3x_{\rm range}, x_{\rm max}-0.3x_{\rm range})$ \\
-$A_{i, \rm wide}$ & Amplitudes of wide spikes &
- $\ln A_{i, \rm wide} \sim \textnormal{Laplace}\left(a_{\ln A, \rm wide}, b_{\ln A, \rm wide}\right)$ \\
-$W_{i, \rm wide}$ & Widths of wide spikes &
- $\ln W_{i, \rm wide} \sim \textnormal{Laplace}\left(a_{\ln W, \rm wide}, b_{\ln W, \rm wide}\right)$ \\
+$A_{i, \rm amorph}$ & Amplitudes of wide peaks &
+ $\ln A_{i, \rm amorph} \sim \textnormal{Laplace}\left(a_{\ln A, \rm amorph}, b_{\ln A, \rm amorph}\right)$ \\
+$W_{i, \rm amorph}$ & Widths of wide peaks &
+ $\ln W_{i, \rm amorph} \sim \textnormal{Laplace}\left(a_{\ln W, \rm amorph}, b_{\ln W, \rm amorph}\right)$ \\
 \hline
@@
 $\sigma_0$ &    Constant noise level  &   $\ln \sigma_0 \sim \textnormal{Laplace}(0,5)$\\
-$\sigma_1$ &    Noise proportionality   &  $\ln \sigma_1 \sim \textnormal{Laplace}(0,5)$ \\
-$\nu$     &   Shape parameter for residuals   &   LogUniform$(1, 1000)$\\
+$\sigma_1$ &    Noise proportional to $\sqrt{\textnormal{model}}$   &  $\ln \sigma_1 \sim \textnormal{Laplace}(0,5)$ \\
+$\nu$     &   Shape parameter for $t$-distribution for noise   &   LogUniform$(1, 1000)$\\
 \hline
@@
 \hline
-$\{y_1, y_2, ..., y_n\}$  &   Measurements    & iid Student-$t(f(x_i; \params), ,\nu)$\\
+$\{y_1, y_2, ..., y_n\}$  &   Measurements    & independent Student-$t(f(x_i; \params), \sqrt{\sigma_0^2 + \sigma_1f(x_i; \params)},\nu)$\\
 \hline
@@
 
+Most of the chosen distributions were intended to be reasonably
+conservative and wide.
+Some of the prior distributions in AMORPH use the log-uniform distribution,
+sometimes (erroneously) called a Jeffreys prior, whose density for a quantity
+$z$ is proportional to $1/z$.
+
+We used a Laplace distribution for some informative priors, rather than
+a more conventional normal distribution, because it assigns higher
+prior probability to values far from the center, and can be computed conveniently, i.e., no special functions such as the inverse of the
+error function are needed.
 The Laplace, or biexponential, distribution with location parameter
@@
 \end{align}
-We used a Laplace distribution for informative priors, rather than
-a more conventional normal distribution, because it assigns higher
-prior probability to values far from the center, and can be computed conveniently, i.e., no special functions such as {\tt erf} are needed
-to compute the cumulative distribution function.
 
-\section{Installation and usage}
+The priors for the positions and widths of the `amorphous' and `crystalline'
+gaussians are different, to account for the prior knowledge that they
+should result in wide gaussians near the centre of the domain, and
+narrow gaussians throughout the domain respectively.
 
-The entire repository, including C++ and Python source code, and a Windows
-executable, is hosted on Bitbucket. The URL for the git repository is
+The prior for the data given all the parameters is a student-$t$ distribution,
+which allows for potentially heavier tails than a normal distribution would.
+Thus, analyses with AMORPH should be able to cope with discrepant datapoints
+as long as they are not locally correlated (which would cause them to be
+mistaken for peaks).
+
+AMORPH does not assume a known number of gaussians for either the
+amorphous or the crystalline component. The appropriate number of
+gaussians is inferred from the data. The maximum number of gaussians
+in each component is set to 300.
+
+\section{Applications}\label{sec:applications}
+\subsection{Crystallinity}
+The AMORPH program calculates a fraction of amorphous material ($A_{\rm m}$) for each posterior sample of the parameters,
+using Equation~\ref{eqn:crystallinity}.Thus measured crystallinity ($C_{\rm m}$) is easily determined from the relationship $C_{\rm m}$ = 1-$A_{\rm m}$, as defined in Equation 5.
+The measured crystallinity is not equivalent to an "actual" or "true" crystallinity, although the values are often very close to one another.
+It should therefore be noted that this procedure does not negate the need for calibration standards
+to translate measured crystallinity to an actual crystallinity.
+Transmission and absorption of X-rays in the material will vary as the ratio of mineral to glass changes in the sample, causing a deviation from a 1:1 line.
+In addition to theoretical considerations, practical challenges also exist for background fitting.
+At low crystallinity ($\lesssim 10\%$) diffraction patterns are dominated by the low amplitude amorphous peak such that removing the effects of background noise becomes problematic, resulting in overestimated crystallinities.
+At high crystallinity ($\gtrsim 90\%$) the problem arises that peak overlaps in multi-mineral assemblages cause broadening effects, and may be subsequently modeled as ``amorphous'' and thus crystallinities  underestimated.
+The calibration curve therefore is utilized to remove these effects.
+This is also an important step as
+different X-ray diffractometers will provide slightly different calibration curves (Fig.~\ref{fig:intruments}).
+Differences between instruments largely reflect changes in background intensity and shape,
+particularly at low values of $2\theta$.
+
+	Calibration standards can be easily created by mixing
+powdered minerals (e.g. olivine, pyroxene, plagioclase) with powdered glass in known
+proportions to simulate volcanic material of varying crystallinities \citep[e.g.,][]{rowe2012, wall2014}.
+Results of the AMORPH calculations show good agreement to calibration curves using the manual
+calibration approach (Fig.~\ref{fig:Comparison}). The offset between the AMORPH and manual calibration methods
+in Figure~\ref{fig:Comparison} is consistent with the change in analyzed $2\theta$ range, 10--40\degree~and 5--50\degree,
+respectively (Fig.~\ref{fig:Outputs}). Since the wide peak for the amorphous material does not extend
+appreciably out of the range from 10--40\degree~$2\theta$ (for a Cu K$\alpha$ source), peak integration
+outside of this range will decrease the relative contribution of the amorphous material
+(only peaks for crystalline materials are outside this range).
+Experimentation demonstrates also that data processing only over the 10--40\degree~$2\theta$ range is also significantly faster, as this reduces the total number of data points for processing.
+Thus we recommend that the optimal range for data processing in terms of time and accuracy is from 10--40\degree~$2\theta$.
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=0.7\textwidth]{figures/instruments.jpg}
+\caption{\it Two calibration curves of the same standard material (rhyolite) are compared from different instrumentation, including a Siemens D500 (Washington State University; red circles) and a PANalytical Empyrean (University of Auckland; blue diamonds).\label{fig:intruments}}
+\end{figure}
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=0.7\textwidth]{figures/Comparison.jpg}
+\caption{\it Comparison of calibration results for rhyolite calculated from AMORPH compared to a calibration calculated using the manual procedure described by \citet{rowe2012}.\label{fig:Comparison}}
+\end{figure}
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=1\textwidth]{figures/Outputs.jpg}
+\caption{\it Model results for a basalt standard containing 80\% amorphous glass.
+Data was processed from 10--40\degree~$2\theta$ (upper left panel) with a calculated amorphous content of 76.4\% (upper right panel).
+The same data was processed instead from 5--50\degree~$2\theta$ (lower left panel) with a calculated amorphous content of 71.5\% (lower right panel).
+Rationale for differences are discussed in the text.\label{fig:Outputs}}
+\end{figure}
+
+\subsection{Amorphous Characteristics}
+As discussed above, one of the primary advantages of the AMORPH program is that it can
+independently model characteristics of the amorphous component.  In particular, centre of mass,
+skewness, and nongaussianity are distinctive amongst different amorphous materials. To test
+this, we present amorphous characteristics of rhyolite and basalt glass picked from a
+Taupo pumice \citep[73.5 wt\% SiO\textsubscript{2}; P2166C;][]{barker2015} and a Kilauea basalt
+\citep[51 wt\% SiO\textsubscript{2}; KS08-108E;][]{wooten2009, rowe2015}, respectively.
+In addition, Mars Scientific Laboratory (MSL) Curiosity rover CheMin instrument reduced X-ray diffraction patterns
+for sediments from the Bagnold Sand Dunes \citep[Gobabeb;][]{achilles2017, lapotre2017}
+and Buckskin \citep{morris2016} localities, both which have significant amorphous
+materials, have been processed for comparison (CheMin instrument data\footnote{\tt http://pds-geosciences.wustle.edu/missions/msl/chmin.htm}). MSL diffraction patterns have been
+recalculated from a Co K$\alpha$ ($1.7890 \dot{A}$) to Cu K$\alpha$ ($1.5406 \dot{A}$) X-ray source using Bragg's law, to be visually comparable to terrestrial
+examples analysed for this study.
+Results are summarized in Figure~\ref{fig:volcanic_glasses}.
+The most notable differences in the amorphous component characteristics calculated for diffraction patterns for basalt and rhyolite glass include a positive skewness and shift to lower centre of mass
+$2\theta$ values for the rhyolite glass compared to the basalt glass. Although, CheMin
+diffraction analyses were run on a different X-ray diffractometer from the terrestrial
+samples shown here, similar relative changes in amorphous characteristics can be observed
+in X-ray diffraction data from Mars.
+In particular, the Buckskin 1 analysis of the amorphous component indicates a positive skewness $\sim 4\times$ greater than in the Gobabeb 2 sample.
+These results suggest a distinct change in the
+composition of the amorphous material. This interpretation, is consistent with prior
+published results which suggest that at the Gobabeb locality (Namib Dune, Bagnold Dune Field),
+the sample is dominated by a basaltic mineralogy, while at the Buckskin locality,
+the amorphous component has been calculated to contain ~77 wt\% SiO\textsubscript{2}
+\citep[i.e. rhyolitic;][]{morris2016, achilles2017}.
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=1\textwidth]{figures/volcanic_glasses.jpg}
+\caption{\it Four examples of changing characteristics of amorphous material based on composition.
+Kilauea basalt glass (upper left) and taupo rhyolite glass (upper right) were analyzed on a PANalytical Empyrean XRD at University of Auckland.
+Gobabeb 2 (lower left) and Buckskin 1 (lower right) results from the CheMin instrument on the MSL Curiosity rover (data sourced as above).
+Calculated statistics for each analysis are provided below the model results.
+Note in particular the change in skewness from Basalt to Rhyolite, and from Gobabeb 2 to Buckskin 1, supporting compositional change.\label{fig:volcanic_glasses}}
+\end{figure}
+
+\section{Conclusions}\label{sec:conclusions}
+The AMORPH program’s statistical approach to interpreting amorphous materials in X-ray
+diffraction patterns provides a new and unique methodology.
+By explicitly specifying a set of assumptions and computing their
+implications, we can automatically produce outputs separating the
+amorphous and crystalline components and quantifying their properties,
+along with a corresponding uncertainty estimate for any inferred
+quantity.
+A significant outcome of this research is that it reduces intra- and inter-user variability
+by eliminating the need for manual background fitting, the largest source of error in the
+calibration methodology for quantifying the amorphous component. Results demonstrate that
+this approach 1) accurately reproduces values of known crystallinity and 2) is consistent
+with prior manual approaches to the calibration method. Quantification of the amorphous
+content however still requires a calibration curve to correct for x-ray absorption/emission
+in heterogeneous materials, and to remove systematic biases in both background fitting and instrumentation.
+
+In addition to the quantification of amorphous and crystalline components, the AMORPH
+program calculates statistical parameters of the amorphous component, including the centre
+of mass, width of amorphous component, skewness, and nongaussianity. Characterization of the amorphous component requires no calibration and outputs show 
+clear distinctions, in particular in terms of skewness, as a function of changing composition of the amorphous material in the demonstrated examples of
+geologic materials from Earth and Mars.
+
+\section*{Acknowledgments}
+
+BJB was supported by a Marsden Fast-Start grant from the Royal Society of
+New Zealand from 2013 to 2016, and this work is a spinoff from that
+project.
+Data used in Figure~\ref{fig:volcanic_glasses} analyzed and processed by J.
+Li, M.
+Jugo, A.
+Serrano.
+Manual calibration for Figure~\ref{fig:Comparison} was conducted by Y.
+Heled.
+This project made use of the resources of the Centre for
+eResearch at the University of Auckland.
+
+\section*{References}
+\begin{thebibliography}{99}
+\bibliographystyle{plainnat}
+
+\bibitem[Achilles et al.(2017)]{achilles2017}
+Achilles, C.N., et al. 2017. Mineralogy of an active eolian sediment from the Namib Dune, Gale Crater, Mars. Journal of Geophysical Research Planets, doi: 10.1002/2017JE005262.
+
+\bibitem[Achilles et al.(2013)]{achilles2013}
+Achilles, C.N., Morris, R.V., Chipera, S.J., Ming, D.W., Rampe, E.B.
+2013.
+X-ray diffraction reference intensity ratios of amorphous and poorly crystalline phases: Implications for CheMin on the Mars Science Laboratory mission.
+44th Lunar and Planetary Science Conference.
+
+\bibitem[Barker et al.(2015)]{barker2015}
+Barker, S.J., Wilson, C.J.N., Allan, A.S.R., Schipper, C.I.
+2015.
+Fine-scale temporal recovery, reconstruction and evolution of a post-supereruption magmatic system.
+Contributions to Mineralogy and Petrology, 170:5; doi: 10.1007/s00410-015-1155-2.
+
+\bibitem[Berger et al.(2016)]{berger2016}
+Berger, J.A., et al. 2016. A global Mars dust composition refined by the Alpha-Particle X-ray Spectrometer in Gale Crater. Geophysical Research Letters, 43, 67-75.
+
+\bibitem[Bish et al.(2013)]{bish2013}
+Bish, D. L. et al. 2013. X-ray diffraction results from Mars Science Laboratory: mineralogy of Rocknest at Gale Crater. Science, 341, 1238932.
+
+\bibitem[Blake et al.(2013)]{blake2013}
+Blake, D. F. et al. 2013. Curiosity at Gale Crater, Mars: characterization and analysis of the Rocknest sand shadow. Science, 341, 1239505.
+
+\bibitem[Brewer et al.(2011)]{dns}
+Brewer, B. J., Pártay, L. B., \& Csányi, G. (2011). Diffusive nested sampling. Statistics and Computing, 21(4), 649-656.
+
+\bibitem[{Brewer} and {Foreman-Mackey}(2016)]{dnest4}
+Brewer, B. J., Foreman-Mackey, D.
+\newblock {DNest4: Diffusive Nested Sampling in C++ and Python}.
+\newblock \emph{Journal of Statistical Software, accepted. arxiv: 1606.03757},
+  June 2016.
+
+\bibitem[Chipera and Bish(2002)]{chipera2002}
+Chipera, S.J., Bish, D.L.
+2002.
+FULLPAT: a full-pattern quantitative analysis program for X-ray powder diffraction using measured and calculated patterns.
+Journal of Applied Crystallography, 35, 744-749.
+
+\bibitem[Dehouck et al.(2014)]{dehouck2014}
+Dehouck, E., McLennan, S.M., Meslin, P-Y., Cousin, A.
+2014.
+Constraints on abundance, composition and nature of X-ray amorphous components of soils and rocks at Gale crater, Mars.
+Journal of Geophysical Research Planets, 119, 2640-2657.
+
+\bibitem[Ellis et al.(2015)]{ellis2015}
+Ellis, B.S., Cordonnier, B., Rowe, M.C., Szymanowski, D., Bachmann, O., Andrews, G.D.M.
+2015.
+Groundmass crystallisation and cooling rates of lava-like ignimbrites: the Grey’s Landing ignimbrite, southern Idaho, USA.
+Bulletin of Volcanology, doi 10.1007/s00445-015-0972-5.
+
+\bibitem[Gregory(2005)]{gregory2005bayesian}
+Gregory, P. C.
+\newblock \emph{Bayesian Logical Data Analysis for the Physical Sciences: A
+  Comparative Approach with Mathematica{\textregistered} Support}.
+\newblock Cambridge University Press, 2005.
+
+\bibitem[Gualtieri(1996)]{gualtieri1996}
+Gualtieri, A. 1996. Modal analysis of pyroclastic rocks by combined Rietveld and RIR methods. Powder Diffraction, 11(2), 97-106.
+
+\bibitem[Gualtieri(2000)]{gualtieri2000}
+Gualtieri, A. 2000. Accuracy of XRPD QPA using the combined Rietveld-RIR method. Journal of Applied Crystallography, 33, 267-278.
+
+\bibitem[Kanakiya et al.(2017)]{kanakiya2017}
+Kanakiya, S., Adam, L., Esteban, L., Rowe, M.C., Shane, P.
+2017.
+Dissolution and secondary mineral precipitation in basalts due to reactions with carbonic acid, Journal of Geophysical Research Solid Earth, 122, 4312-4327, doi:10.1002/2017JB014019.
+
+\bibitem[Knuth and Skilling(2012)]{knuth2012foundations}
+Knuth, K. H., Skilling, J.
+\newblock Foundations of inference.
+\newblock \emph{Axioms}, 1\penalty0 (1):\penalty0 38--73, 2012.
+
+\bibitem[Lapotre et al.(2017)]{lapotre2017}
+Lapotre, M.G.A., Ehlmann, B.L., Minson, S.E., Arvidson, R.E., Ayoug, F., Fraeman, A.A., Ewing, R.C., Bridges, N.T.
+2017.
+Compositional variations in sands of the Bagnold Dunes, Gale Crater, Mars, from Visible-Shortwave Infrared Spectroscopy and comparison with ground truth from the Curiosity Rover, Journal of Geophysical Research Planets; doi: 10.1002/2016JE005133.
+
+\bibitem[Mackay(2003)]{mackay2003}
+MacKay, D. J. C. (2003). Information theory, inference and learning algorithms. Cambridge university press.
+
+\bibitem[Morris et al.(2016)]{morris2016}
+Morris, R.V., et al. 2016. Silicic volcanism on Mars evidenced by tridymite in high-SiO2 sedimentary rock at Gale crater. PNAS, 113 (26), 7071-7076.
+
+\bibitem[O'Hagan and Forster(2004)]{o2004kendall}
+O'Hagan, A., Forster, J. J.
+\newblock \emph{Kendall's advanced theory of statistics, volume 2B: Bayesian
+  inference}, volume~2.
+\newblock Arnold, 2004.
+
+\bibitem[Rowe et al.(2012)]{rowe2012}
+Rowe, M.C., Ellis, B.S., Lindeberg, A.
+2012.
+Quantifying crystallization and devitrification of rhyolites by means of X-ray diffraction and electron microprobe analysis.
+American Mineralogist, 97, 1685-1699.
+
+\bibitem[Rowe et al.(2015)]{rowe2015}
+Rowe, M.C., Thornber, C.R., Orr, T.R.
+2015.
+Primitive components, crustal assimilation, and magmatic degassing during the early 2008 Kilauea summit eruptive activity.
+Hawaiian Volcanoes: From Source to Surface, ed(s) R.
+Carey, V.
+Cayol, M.
+Poland, and D.
+Weis.
+p.
+439-455.
+
+\bibitem[Schmidt et al.(2009)]{schmidt2009}
+Schmidt, M.
+E.
+et al.
+2009.
+Spectral, mineralogical, and geochemical variations across Home Plate, Gusev Crater, Mars indicate high and low temperature alteration.
+Earth Planet.
+Sci.
+Lett.
+281, 258–268.
+
+\bibitem[Sivia and Skilling(2006)]{sivia2006data}
+Sivia, D., Skilling, J.
+\newblock \emph{Data analysis: a Bayesian tutorial}.
+\newblock OUP Oxford, 2006.
+
+\bibitem[Wall et al.(2014)]{wall2014}
+Wall, K.T., Rowe, M.C., Ellis, B.S., Schmidt, M.E., Eccles, J.D.
+2014.
+Determining volcanic eruption styles on Earth and Mars from crystallinity measurements.
+Nature Communications, doi: 10.1038/ncomms6090.
+
+\bibitem[Wooten et al.(2009)]{wooten2009}
+Wooten, K.M., Thornber, C.R., Orr, T.R., Ellis, J.F., and Trusdell, F.A.
+2009.
+Catalog of tephra samples from Kilauea's summit eruption, March-December 2008: U.S.
+Geological Survey Open-File Report 2009-1134, 26p.
+
+\bibitem[Zorn et al.(2017)]{zorn2017}
+Zorn, E.U., Rowe, M.C., Cronin, S.J., Ryan, A.G., Kennedy, L., Russell, J.K.
+2017.
+Influence of porosity and groundmass crystallinity on dome rock strength; a case study from Mt.
+Taranaki, New Zealand.
+Bulletin of Volcanology, in review.
+
+\end{thebibliography}
+
+\appendix
+\section{Installation and usage}\label{sec:program}
+
+The entire repository, including C++ and Python source code, and a 
+precompiled executable file for Microsoft Windows,
+is hosted on Bitbucket at the following URL:
+
+\vspace{1em}
+{\tt https://bitbucket.org/eggplantbren/amorph}
+\vspace{1em}
+
+AMORPH is free software, released under the terms of the GNU General Public
+Licence, version 3.
 For most users, we suggest simply using the pre-compiled Windows executable
-{\tt main.exe} from the repository.
+{\tt AMORPH.exe} from the repository. To execute the program,
+run {\tt AMORPH.exe}. You will be prompted to enter the filename of the text
+file containing the data.
+All files to be processed by AMORPH need to be in .txt file format, space delimited, with no headers.
+For simplicity, text files for processing should be located in the same folder as the {\tt AMORPH.exe} program.
+The program is set to run until 10,000 saved parameter sets have been generated.
+Data outputs may be viewed at any point, however, closing the program before reaching  10,000 will reduce the accuracy of the final calculations.
+After running for a while, the output can be viewed
+by running the Python script {\tt showresults.py}:
 
-The repository can simply be downloaded from
+\vspace{1em}
+{\tt python showresults.py}
+\vspace{1em}
 
-{\tt https://bitbucket.org/eggplantbren/crystals/get/c78decf570f0.zip}
-
-and unzipped into any location on your computer. To execute the program,
-run {\tt main.exe}. You will be prompted to enter the filename of the text
-file containing the data.
-
-\bibliographystyle{plainnat}
-\bibliography{references}
+This script makes use of the packages {\tt Numpy} and {\tt matplotlib},
+and has only been tested under Python 3. Anaconda\footnote{\tt https://www.anaconda.com/download/} is a convenient distribution
+of Python which comes with scientific packages pre-installed.
+Since AMORPH is really just a specific data analysis situation implemented
+for DNest4 \citep{dnest4}, that paper provides much more detail about
+AMORPH's output and further available options.
 

### 2017-09-09 doc/figures/pgm.py
Updated figures

@@ -1,0 +1,54 @@
+"""
+PGM for the general problem using daft (http://daft-pgm.org/)
+"""
+
+import matplotlib
+from matplotlib import rc
+rc("font", family="serif", size=12)
+rc("text", usetex=True)
+matplotlib.rcParams['text.latex.preamble']=[r"\usepackage{amsmath}"]
+
+import daft
+
+pgm = daft.PGM([9., 5.], origin=[-2.5, -2.5])
+
+# Create the nodes
+pgm.add_node(daft.Node('alpha', r'$\boldsymbol{\alpha}_{\rm amorph}$', 0., 2.))
+pgm.add_node(daft.Node('x', r'$\boldsymbol{\theta}_i^{\rm amorph}$', 0., 0.))
+pgm.add_node(daft.Node('data', r'$\boldsymbol{y}$', 2., 0., observed=True))
+pgm.add_node(daft.Node('N', r'$N_{\rm amorph}$', -2., 0.))
+pgm.add_node(daft.Node('hidden', r'', -1., 0., scale=0, plot_params={'alpha':0}))
+
+# Right hand side
+pgm.add_node(daft.Node('x2', r'$\boldsymbol{\theta}_i^{\rm xtal}$', 4., 0.))
+pgm.add_node(daft.Node('alpha2', r'$\boldsymbol{\alpha}_{\rm xtal}$', 4., 2.))
+pgm.add_node(daft.Node('N2', r'$N_{\rm xtal}$', 6., 0.))
+pgm.add_node(daft.Node('hidden2', r'', 5., 0., scale=0, plot_params={'alpha':0}))
+
+# Top and bottom
+pgm.add_node(daft.Node('sigma', r'$\boldsymbol{\sigma}$', 2., 2.))
+pgm.add_node(daft.Node('sie', r'$\boldsymbol{\theta}_{\rm bg}$', 2., -2.))
+
+# Add the edges
+pgm.add_edge('alpha', 'x')
+pgm.add_edge('x', 'data')
+pgm.add_edge('N', 'hidden')
+
+pgm.add_edge('alpha2', 'x2')
+pgm.add_edge('x2', 'data')
+pgm.add_edge('N2', 'hidden2')
+
+pgm.add_edge('sigma', 'data')
+pgm.add_edge('sie', 'data')
+
+# Add the plates
[…]

### 2017-09-09 doc/ms.tex
Arxivable formatting
Removed journal name
Updated references
Geometry tweak
Couple of minor edits

@@
-\documentclass[review]{elsarticle}
+\documentclass[preprint, a4paper]{elsarticle}
 \usepackage{amsmath}
@@
 \usepackage{graphicx}
-% \usepackage[left=2cm, right=2cm, bottom=3cm, top=2cm]{geometry}
+\usepackage[left=2.2cm, right=2.2cm, bottom=3cm, top=2.5cm]{geometry}
 \usepackage{microtype}
@@
 \usepackage{gensymb} % For degree symbol
-\usepackage{lineno,hyperref}
-\modulolinenumbers[5]
+%\usepackage{lineno,hyperref}
+%\modulolinenumbers[5]
 
-\journal{Computers and Geosciences}
+\journal{}
 
@@
 
-\linenumbers
+%\linenumbers
 
@@
 The AMORPH program calculates a fraction of amorphous material ($A_{\rm m}$) for each posterior sample of the parameters,
-using Equation~\ref{eqn:crystallinity}.Thus measured crystallinity ($C_{\rm m}$) is easily determined from the relationship $C_{\rm m}$ = 1-$A_{\rm m}$, as defined in Equation 5.
-The measured crystallinity is not equivalent to an "actual" or "true" crystallinity, although the values are often very close to one another.
+using Equation~\ref{eqn:crystallinity}.
+Thus, measured crystallinity ($C_{\rm m}$) is easily determined from the relationship $C_{\rm m} = 1-A_{\rm m}$, as defined in Equation 5.
+The measured crystallinity is not equivalent to an ``actual'' or ``true'' crystallinity, although the values are often very close to one another.
 It should therefore be noted that this procedure does not negate the need for calibration standards
@@
 \bibitem[Brewer et al.(2011)]{dns}
-Brewer, B. J., Pártay, L. B., \& Csányi, G. (2011). Diffusive nested sampling. Statistics and Computing, 21(4), 649-656.
+Brewer, B. J., Pártay, L. B., \& Csányi, G. 2011. Diffusive nested sampling. Statistics and Computing, 21(4), 649-656.
 
 \bibitem[{Brewer} and {Foreman-Mackey}(2016)]{dnest4}
-Brewer, B. J., Foreman-Mackey, D.
+Brewer, B. J., Foreman-Mackey, D. 2016.
 \newblock {DNest4: Diffusive Nested Sampling in C++ and Python}.
-\newblock \emph{Journal of Statistical Software, accepted. arxiv: 1606.03757},
-  June 2016.
+\newblock \emph{Journal of Statistical Software, accepted. arxiv: 1606.03757}.
 
@@
 \bibitem[Gregory(2005)]{gregory2005bayesian}
-Gregory, P. C.
+Gregory, P. C. 2005.
 \newblock \emph{Bayesian Logical Data Analysis for the Physical Sciences: A
   Comparative Approach with Mathematica{\textregistered} Support}.
-\newblock Cambridge University Press, 2005.
+\newblock Cambridge University Press, 468pp.
 
@@
 \bibitem[Knuth and Skilling(2012)]{knuth2012foundations}
-Knuth, K. H., Skilling, J.
+Knuth, K. H., Skilling, J. 2012.
 \newblock Foundations of inference.
-\newblock \emph{Axioms}, 1\penalty0 (1):\penalty0 38--73, 2012.
+\newblock Axioms, 1\penalty0 (1), \penalty0 38--73.
 
@@
 \bibitem[Mackay(2003)]{mackay2003}
-MacKay, D. J. C. (2003). Information theory, inference and learning algorithms. Cambridge university press.
+MacKay, D. J. C. 2003. Information theory, inference and learning algorithms. Cambridge University Press, 628pp.
 
@@
 \bibitem[O'Hagan and Forster(2004)]{o2004kendall}
-O'Hagan, A., Forster, J. J.
+O'Hagan, A., Forster, J. J. 2004.
 \newblock \emph{Kendall's advanced theory of statistics, volume 2B: Bayesian
   inference}, volume~2.
-\newblock Arnold, 2004.
+\newblock Arnold, 496pp.
 
@@
 Spectral, mineralogical, and geochemical variations across Home Plate, Gusev Crater, Mars indicate high and low temperature alteration.
-Earth Planet.
-Sci.
-Lett.
-281, 258–268.
+Earth and Planetary Science Letters, 281, 258–268.
 
 \bibitem[Sivia and Skilling(2006)]{sivia2006data}
-Sivia, D., Skilling, J.
+Sivia, D., Skilling, J. 2006.
 \newblock \emph{Data analysis: a Bayesian tutorial}.
-\newblock OUP Oxford, 2006.
+\newblock Oxford University Press, 246pp.
 

### 2018-03-23 src/display.py
Fixed typo

@@ -150,8 +150,8 @@ def display():
 
     plt.hist(wide_fraction, 100, color=[0.3, 0.3, 0.3])
     plt.xlim([0, 1])
-    plt.xlabel("(amporph)/(amorph + crystal)")
-    print("(amporph)/(amorph + crystal) = {a} +- {b}".format(a=wide_fraction.mean(),\
+    plt.xlabel("(amorph)/(amorph + crystal)")
+    print("(amorph)/(amorph + crystal) = {a} +- {b}".format(a=wide_fraction.mean(),\
            b=wide_fraction.std()))
     plt.show()
 

### 2018-03-23 src/MyModel.cpp
Potential speedup

@@ -63,12 +63,24 @@ double MyModel::perturb(DNest4::RNG& rng)
         // Perturb spikes
         logH += narrow_peaks.perturb(rng);
 
+        // Save computational time by handling prior-based rejection here
+        if(rng.rand() >= exp(logH))
+            return -1E300;
+        else
+            logH = 0.0;
+            
         compute_narrow(narrow_peaks.get_removed().size() == 0);
     }
     else if(choice == 1)
     {
         // Perturb spikes
         logH += wide_peaks.perturb(rng);
+
+        // Save computational time by handling prior-based rejection here
+        if(rng.rand() >= exp(logH))
+            return -1E300;
+        else
+            logH = 0.0;
 
         compute_wide(wide_peaks.get_removed().size() == 0);
     }

### 2018-03-25 doc/ms.tex
Downloaded manuscript from overleaf

@@
-\documentclass[preprint, a4paper]{elsarticle}
+\documentclass[review]{elsarticle}
 \usepackage{amsmath}
@@
 \usepackage{graphicx}
-\usepackage[left=2.2cm, right=2.2cm, bottom=3cm, top=2.5cm]{geometry}
+% \usepackage[left=2cm, right=2cm, bottom=3cm, top=2cm]{geometry}
 \usepackage{microtype}
@@
 \usepackage{gensymb} % For degree symbol
-%\usepackage{lineno,hyperref}
-%\modulolinenumbers[5]
+\usepackage{lineno,hyperref}
+\modulolinenumbers[5]
 
-\journal{}
+\journal{Computers and Geosciences}
 
@@
 \begin{abstract}
-AMORPH utilizes a new Bayesian statistical approach to interpreting X-ray diffraction results of partially amorphous materials. AMORPH fits X-ray diffraction patterns with a mixture of gaussians,
+AMORPH utilizes a new Bayesian statistical approach to interpreting X-ray diffraction results of samples with both crystalline and amorphous components.
+AMORPH fits X-ray diffraction patterns with a mixture of gaussians,
 which allows us to simultaneously infer all of the model parameters and quantify their
@@
 
-%\linenumbers
+\linenumbers
 
@@
 Quantification and characterization of amorphous materials is an important area of research in
-terrestrial and planetary sciences \citep[e.g.,][]{schmidt2009, dehouck2014, wall2014, morris2016, zorn2017}.
+terrestrial and planetary sciences \citep[e.g.,][]{schmidt2009, dehouck2014, wall2014, morris2016, zorn2018}.
 Powder X-ray diffraction (XRD), although primarily a tool for characterization of crystalline materials,
-also provides a means to investigate amorphous materials \citep[e.g.,][]{rowe2012, achilles2013}. X-ray diffraction of rocks/soils containing amorphous materials (e.g. volcanic
+also provides a means to investigate amorphous materials \citep[e.g.,][]{rowe2012, achilles2013}.
+X-ray amorphous materials lack long-range crystallographic order and produce broad humps with low intensities in XRD patterns.
+X-ray diffraction of rocks/soils containing amorphous materials (e.g.
+volcanic
 glass, amorphous silica, and organics) produces a diffraction pattern which is a combination of the
@@
 applications to terrestrial volcanics have presented an avenue of expansion for XRD methodologies
-for amorphous material analysis \citep{ellis2015, kanakiya2017, zorn2017}.
+for amorphous material analysis \citep{ellis2015, kanakiya2017, zorn2018}.
 Here we present a new statistical approach and computer program, AMORPH, which utilizes Bayesian inference to determine
@@
 technique \citep[FULLPAT;][]{chipera2002} is routinely utilized for characterizing remotely acquired X-ray
-diffraction patterns, such as those recovered from the Mars Science Laboratory (MSL) Curiosity rover \citep{bish2013}. The problem faced by RIR approaches is they are based on modeling of an observed 
+diffraction patterns, such as those measured by the CheMin instrument on the Mars Science Laboratory (MSL) Curiosity rover \citep{blake2012, bish2013}.
+The problem faced by RIR approaches is they are based on modeling of an observed
 diffraction pattern compared to library measurements, and therefore do not allow an independent
-assessment of the characteristics of the amorphous material. In comparison, the calibration method
+assessment of the characteristics of the amorphous material.
+Furthermore, natural X-ray amorphous phases have a wide variety of compositions such that the library measurements may not include the amorphous phase(s) present in the sample.
+In comparison, the calibration method
 focuses solely on the quantification of the amorphous material, utilizing the integrated counts
@@
 Diffusive Nested Sampling, an advanced MCMC method \citep{dns, dnest4}.
+There have been many applications of Bayesian inference to
+problems of separating background from compact signals
+when noise is present \citep[e.g.][]{}. AMORPH can be viewed as another
+case of this, with assumptions customized to the specific application.
 
@@
 the parameters. This describes uncertainty about what data will be observed,
-but encodes some knowledge about the fact that the data has something to do with the
+but encodes some knowledge about the fact that the data have something to do with the
 parameters $\params$. For example, $p(\data | \params, \info)$ often encodes
@@
 Bayesian inference via MCMC involves finding a sampling of plausible values
-for the parameters.
+for the parameters. Since AMORPH uses \citet{dnest4}, it also calculates
+the `marginal likelihood' or `evidence' quantity $p(\data | \info)$, which
+quantifies how well the model fits the data overall, averaged over all
+possible values of its parameters. This allows straightforward model comparison
+with any alternatives that may be proposed in the future
+\citep{skilling2006nested}.
+
+% BJB: Hopefully this addresses reviewer 1, point a.
+The computational cost of running AMORPH on a data set of around 1000
+points varies, as the posterior distribution depends on the data.
+Another important factor is the number of MCMC samples desired.
+On a modern PC, some useful posterior samples can be obtained in about
+half an hour.
+With 300 posterior samples, typical analyses take approximately 7 hours to complete, depending on computing capabilities and the number of data points.
+However, the complete runs presented in this paper took
+about a day each to run, as we wanted to ensure the results contained
+about 1000 posterior samples.
 
@@
 and between $\x=40$\degree~and $\x=\x_{\rm max}$}.
-These values are selected to maintain a linear background over the interval from 10--40\degree, the predominant range of amorphous materials analyzed using a Cu K$\alpha$ x-ray source, while still allowing for some flexibility in the structure of the background outside this range.
+These values are selected to maintain a linear background over the interval from 10--40\degree, the predominant range of amorphous materials (8.8 -- 2.25 \AA) analyzed using a Cu K$\alpha$ x-ray source (Table~\ref{tab:anode}), while still allowing for some flexibility in the structure of the background outside this range.
 
@@
 \end{align}
+Control points may be modified, depending on the X-ray source in use and changes in the required range for the linear background as a function of the nature of the amorphous material (Table~\ref{tab:anode}).
 The value of the background component at any other point is found by
@@
 values of the parameters.
+
+\begin{table}
+\footnotesize
+\centering
+\begin{tabular}{|llll|}
+\hline
+{\bf X-ray Anode}      &   {\bf X-ray wavelength}   &  {\bf Control point 2} & {\bf Control point 3}\\
+ &   {\bf (\AA)}   &  {\bf(\degree 2 theta)} & {\bf(\degree 2 theta)}\\
+\hline
+Cr & 2.29 & 15.0 & 61.2\\
+Fe & 1.94 & 12.7 & 51.1\\
+Co & 1.79 & 11.7 & 46.9\\
+{\bf Cu} & {\bf 1.54} & {\bf 10.0} & {\bf 40.0}\\
+Mo & 0.71 & 4.6 & 18.2\\
+Ag & 0.56 & 3.6 & 14.3\\
+\hline
+\end{tabular}
+\caption{\it Typical XRD anodes with recommended background control points based on a predominant 8.8-2.25 {\AA}  amorphous peak.\label{tab:anode}}
+\end{table}
 
@@
 \subsection{The crystalline component}
-We assume that there is some number $N_{\rm xtal}$ of narrow peaks,
-each of which has a gaussian shape. The $i$th peak is parameterized
+
+% BJB: Reponse to Reviewer 2, comment about Lorentzian shape vs Gaussian
+% We experimented with the Lorentzian shape instead of the Gaussian and found
+% that for same datasets, it was preferred, whereas for other datasets, the
+% Gaussian is preferred (as adjudicated by the marginal likelihood values
+% for the two different assumptions). Rather than having to run each model
+% twice, we decided it would be more elegant to replace the gaussian assumption
+% with some family of shapes that can smoothly interpolate between gaussian
+% and lorentzian as necessary. Specifically, we chose the Student's
+% t-distribution which can do this straightforwardly. This introduces one
+% more parameter to the model, which controls the shape of the peaks.
+
+% I (BJB) also changed the parameterisation of the heights of the peaks, so
+% they are described by the total integrated mass rather than the peak
+% amplitude. This makes the Metropolis exploration a bit more efficient.
+
+We assume that there is some number $N_{\rm xtal}$ of narrow peaks
+due to crystalline material,
+each of which has the same shape. The shapes of the narrow
+peaks are controlled by a single shape parameter $\nu_{\rm xtal}$,
+assumed common to all of the narrow peaks. The chosen shape is based
+on the Student's $t$-distribution, which allows to peaks to be heavy-tailed
+if $\nu_{\rm xtal}$ is low, and to be Lorentzian or Cauchy in shape if
+$\nu_{\rm xtal}$ is exactly 1. As $\nu_{\rm xtal}$ is high, the shape
+approximates a gaussian.
+
+The $i$th peak is parameterized
 by its central position $X_{i, \rm xtal}$,
-amplitude $A_{i, \rm xtal}$, and width $W_{i, \rm xtal}$.
+total integrated `mass' $M_{i, \rm xtal}$, and width $W_{i, \rm xtal}$.
 The parameter vector for crystalline component is
@@
 \params_{\rm xtal} &=
-  \{N_{\rm xtal}, X_1, X_2, ..., X_{N, \rm xtal},
-    A_1, A_2, ..., A_{N, \rm xtal},
-    W_1, W_2, ..., W_{N, \rm xtal}\}
+  \{N_{\rm xtal}, \nu_{\rm xtal}, X_{1, \rm xtal}, ..., X_{N, \rm xtal},
+    M_{1, \rm xtal}, ..., M_{N, \rm xtal},
+    W_{1, \rm xtal}, ..., W_{N, \rm xtal}\}
 \end{align}
 and the total shape contributed by the
-crystalline component is a sum of the $N_{\rm xtal}$ gaussians:
+crystalline component is a sum of the $N_{\rm xtal}$ peaks:
 \begin{align}
 f_{\rm xtal}(\x; \params_{\rm xtal}) &=
-    \sum_{i=1}^{N_{\rm xtal}} A_{i, \rm xtal}
- \exp\left[-\frac{1}{2W_{i, \rm xtal}^2}\left(x - X_{i, \rm xtal}\right)^2\right].
+	\frac{\Gamma(\frac{\nu_{\rm xtal} + 1}{2})}{\sqrt{\nu_{\rm xtal}\pi}
+                 		\Gamma(\frac{\nu_{\rm xtal}}{2})}
+    \sum_{i=1}^{N_{\rm xtal}}
+    		\frac{M_{i, \rm xtal}}
+            	 {W_{i, \rm xtal}}
+	\left(1 + \frac{1}{\nu}\left(\frac{x - X_{i, \rm xtal}}{W_{i, \rm xtal}}\right)^2\right)
+    			^{-\frac{\nu_{\rm xtal} + 1}{2}}. \label{eqn:t_shape}
 \end{align}
+The terms inside the sum have the form of the probability density function
+of the $t$-distribution, adjusted by shifting according to the central
+position and widening by the width, and the total integral is scaled by
+the mass parameters. $\Gamma()$ is the gamma function.
+
+For computational efficiency, evaluation of Equation~\ref{eqn:t_shape}
+is performed using a lookup table.
 
 \subsection{The amorphous component}
-The amorphous component is also composed of gaussians, like the crystalline component.
-In fact, the only real difference is the prior distributions for the positions,
-amplitudes, and widths, given in Section~\ref{sec:priors}.
+The amorphous component is also composed of a sum of peaks, which we chose to be
+gaussians. The parameterization is the same as for the crystalline peaks but with a fixed,
+high value for the shape parameter.
+Apart from the gaussian shape, the only other difference is the prior distributions for the positions, masses, and widths, given in Section~\ref{sec:priors}.
+In short, the wide gaussians representing the
+amorphous component are more likely to be located near the middle of the
+dataset, and to be wider.
 
@@
 by its central position $X_{i, \rm amorph}$,
-amplitude $A_{i, \rm amorph}$, and width $W_{i, \rm amorph}$.
+mass (total integral) $M_{i, \rm amorph}$, and width $W_{i, \rm amorph}$.
 The parameter vector for wide gaussians is
@@
 \params_{\rm amorph} &=
-  \{N_{\rm amorph}, X_1, X_2, ..., X_{N, \rm amorph},
-    A_1, A_2, ..., A_{N, \rm amorph},
-    W_1, W_2, ..., W_{N, \rm amorph}\}
+  \{N_{\rm amorph}, X_{1, \rm amorph}, ..., X_{N, \rm amorph},
+    M_{1, \rm amorph}, ..., M_{N, \rm amorph},
+    W_{1, \rm amorph}, ..., W_{N, \rm amorph}\}
 \end{align}
@@
 f_{\rm amorph}(\x; \params_{\rm amorph}) &=
-    \sum_{i=1}^{N_{\rm amorph}} A_{i, \rm amorph}
+	\frac{1}{\sqrt{2\pi}}
+    \sum_{i=1}^{N_{\rm amorph}} \frac{M_{i, \rm amorph}}{W_{i, \rm amorph}}
  \exp\left[-\frac{1}{2W_{i, \rm amorph}^2}\left(x - X_{i, \rm amorph}\right)^2\right].
@@
 {\em Hyperparameters} & &\\
-$N_{\rm xtal}$   &   Number of narrow peaks    &  $\propto 1/(N_{\rm xtal}+1)$, $N_{\rm xtal} \in \{0, 1, ..., 300\}$ \\
+$N_{\rm xtal}$   &   Number of narrow peaks    &  $\propto 1/(N_{\rm xtal}+1)$, $N_{\rm xtal} \in \{0, 1, ..., 1000\}$ \\
 $N_{\rm amorph}$   &   Number of wide peaks    &  $\propto 1/(N_{\rm amorph}+1)$,
-$N_{\rm amorph} \in \{0, 1, ..., 300\}$ \\
-$a_{\ln A, \rm xtal}$ & Typical log-amplitude of narrow peaks & Laplace(0, 5)\\
-$a_{\ln A, \rm amorph}$ & Typical log-amplitude of wide peaks & Laplace(0, 5)\\
-$b_{\ln A, \rm xtal}$ & Diversity of log-amplitude of narrow peaks & Uniform(0, 5)\\
-$b_{\ln A, \rm amorph}$ & Diversity of log-amplitude of wide peaks & Uniform(0, 5)\\
+$N_{\rm amorph} \in \{0, 1, ..., 1000\}$ \\
+$a_{\ln M, \rm xtal}$ & Typical log-mass of narrow peaks & Laplace(0, 5)\\
+$a_{\ln M, \rm amorph}$ & Typical log-mass of wide peaks & Laplace(0, 5)\\
+$b_{\ln M, \rm xtal}$ & Diversity of log-mass of narrow peaks & Uniform(0, 5)\\
+$b_{\ln M, \rm amorph}$ & Diversity of log-mass of wide peaks & Uniform(0, 5)\\
 $a_{\ln W, \rm xtal}$ & Typical log-width of narrow peaks & LogUniform$(10^{-3}x_{\rm range}, 0.05x_{\rm range})$\\
@@
                             Uniform$(x_{\rm min}, x_{\rm max})$ \\
-$A_{i, \rm xtal}$ & Amplitudes of narrow peaks &
- $\ln A_{i, \rm xtal} \sim \textnormal{Laplace}\left(a_{\ln A, \rm xtal}, b_{\ln A, \rm xtal}\right)$ \\
+$A_{i, \rm xtal}$ & Masses of narrow peaks &
+ $\ln M_{i, \rm xtal} \sim \textnormal{Laplace}\left(a_{\ln M, \rm xtal}, b_{\ln M, \rm xtal}\right)$ \\
 $W_{i, \rm xtal}$ & Widths of narrow peaks &
  $\ln W_{i, \rm xtal} \sim \textnormal{Laplace}\left(a_{\ln W, \rm xtal}, b_{\ln W, \rm xtal}\right)$ \\
+$\nu_{\rm xtal}$ & Shape of narrow peaks (common) & LogUniform$(0.1, 100)$\\
 \hline
@@
           Uniform$(x_{\rm min}+0.3x_{\rm range}, x_{\rm max}-0.3x_{\rm range})$ \\
-$A_{i, \rm amorph}$ & Amplitudes of wide peaks &
- $\ln A_{i, \rm amorph} \sim \textnormal{Laplace}\left(a_{\ln A, \rm amorph}, b_{\ln A, \rm amorph}\right)$ \\
+$A_{i, \rm amorph}$ & Masses of wide peaks &
+ $\ln M_{i, \rm amorph} \sim \textnormal{Laplace}\left(a_{\ln M, \rm amorph}, b_{\ln M, \rm amorph}\right)$ \\
 $W_{i, \rm amorph}$ & Widths of wide peaks &
@@
 \subsection{Crystallinity}
+To date, applications of crystallinity determinations via XRD methodologies on terrestrial samples have primarily focused on volcanic materials \citep[e.g.,][]{wall2014, ellis2015, andrade2017, zorn2018}, linking crystallinity of volcanic glass to timescales of geologic processes.
+The primary advantage of this technique being its ability to determine crystallinities for finer grain sizes than typically possible through standard petrologic techniques \citep{rowe2012}.
+In contrast, planetary science usages for determining crystallinity have largely focused on quantification of the amorphous and crystalline components for mass balance determination of physical and chemical components in diverse geologic materials \citep[e.g.,][]{blake2013, dehouck2014}.
+
 The AMORPH program calculates a fraction of amorphous material ($A_{\rm m}$) for each posterior sample of the parameters,
-using Equation~\ref{eqn:crystallinity}.
-Thus, measured crystallinity ($C_{\rm m}$) is easily determined from the relationship $C_{\rm m} = 1-A_{\rm m}$, as defined in Equation 5.
-The measured crystallinity is not equivalent to an ``actual'' or ``true'' crystallinity, although the values are often very close to one another.
-It should therefore be noted that this procedure does not negate the need for calibration standards
-to translate measured crystallinity to an actual crystallinity.
-Transmission and absorption of X-rays in the material will vary as the ratio of mineral to glass changes in the sample, causing a deviation from a 1:1 line.
+using Equation~\ref{eqn:crystallinity}.Thus measured crystallinity ($C_{\rm m}$) is easily determined from the relationship $C_{\rm m}$ = 1-$A_{\rm m}$, as defined in Equation 5. 
+To test the applicability of AMORPH crystallinity results we have analysed the Mars Scientific Laboratory (MSL)  CheMin instrument X-ray diffraction patterns of loose basaltic sand from the Bagnold Sand Dunes \citep[Gobabeb;][]{achilles2017, lapotre2017}
+and sediments drilled from a silicic mudstone \citep[Buckskin sample;][]{morris2016}, both which have significant amorphous
+materials (CheMin instrument data\footnote{\tt http://pds-geosciences.wustl.edu/msl/msl-m-chemin-4-rdr-v1/mslcmn{\textunderscore}1xxx.data.rdr4/}).
+CheMin diffraction patterns have been processed using the original Co K$\alpha$ ($1.7890 \dot{A}$) data as well as data transformed from a Co K$\alpha$ to Cu K$\alpha$ ($1.5406 \dot{A}$) X-ray source using Bragg's law.
+This recalculation simply allows for an easier visual comparison to terrestrial samples analysed with the more standard Cu K$\alpha$ X-ray source (see below).
+Note that for Co K$\alpha$ results, background critical points have been adjusted to 11.7-46.9 $2\theta$ to maintain the linear background over the amorphous region from ~8.8 to 2.25 $\dot{A}$ (Table ~\ref{tab:anode}).
+Although the results are uncalibrated, measured crystallinites provide a good first approximation of actual crystallinity, and relative differences are robust.
+Measured amorphous content of the Gobabeb 2 sample is 36.7 $\pm$ 2.9$\%$ (63.3$\%$ crystallinity) while the Buckskin 2 sediment contains 45.3 $\pm$ 2.9$\%$ amorphous phases (54.7 $\%$ crystallinity; Fig.~\ref{fig:Mars}).
+Comparatively, published FULLPAT calculations of amorphous contents are ~35 $\pm$ 15$\%$ for Gobabeb and 50 $\pm$ 15$\%$ for Buckskin \citep{achilles2017, morris2016}.
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=1\textwidth]{figures/Mars.jpg}
+\caption{\it Analysis of Gobabeb 2 (left) and Buckskin 2 (right) results from the CheMin instrument on the MSL Curiosity rover (data sourced as above).
+Upper panels  illustrate initial Co K$\alpha$ processed results while lower panels show AMORPH results after translating data to a Cu K$\alpha$ source (see text for details).
+Calculated statistics provided by AMORPH for each analysis are indicated below the model results.\label{fig:Mars}}
+\end{figure}
+
+The measured crystallinity is not equivalent to an "actual" or "true" crystallinity, although the values are often very close to one another as noted above.
+Therefore, this procedure does not negate the need for calibration standards
+to translate measured crystallinity to an actual crystallinity.Calibration standards can be easily created by mixing
+powdered minerals (e.g. olivine, pyroxene, plagioclase) with powdered glass in known
+proportions to simulate volcanic material of varying crystallinities \citep[e.g.,][]{rowe2012, wall2014}.
+
+Transmission and absorption of X-rays in the material will vary as the ratio of mineral to glass changes in the sample, causing a deviation from a 1:1 line (Fig.~\ref{fig:instruments}).
 In addition to theoretical considerations, practical challenges also exist for background fitting.
 At low crystallinity ($\lesssim 10\%$) diffraction patterns are dominated by the low amplitude amorphous peak such that removing the effects of background noise becomes problematic, resulting in overestimated crystallinities.
-At high crystallinity ($\gtrsim 90\%$) the problem arises that peak overlaps in multi-mineral assemblages cause broadening effects, and may be subsequently modeled as ``amorphous'' and thus crystallinities  underestimated.
+At high crystallinity ($\gtrsim 90\%$) the problem arises that peak overlaps in multi-mineral assemblages cause apparent broadening effects, and may be subsequently modeled as ``amorphous'' and thus crystallinities  underestimated.
 The calibration curve therefore is utilized to remove these effects.
 This is also an important step as
-different X-ray diffractometers will provide slightly different calibration curves (Fig.~\ref{fig:intruments}).
+different X-ray diffractometers will provide slightly different calibration curves (Fig.~\ref{fig:instruments}).
 Differences between instruments largely reflect changes in background intensity and shape,
@@
 
-	Calibration standards can be easily created by mixing
-powdered minerals (e.g. olivine, pyroxene, plagioclase) with powdered glass in known
-proportions to simulate volcanic material of varying crystallinities \citep[e.g.,][]{rowe2012, wall2014}.
+Since the wide peak for many amorphous materials does not extend
+appreciably out of the range from 10--40\degree~$2\theta$ (for a Cu K$\alpha$ source), peak area integration
+outside of this range will decrease proportionally the relative contribution of the amorphous material
+(only peaks for crystalline materials are outside this range in the example depicted in Figure ~\ref{fig:Outputs}).
+It should be noted of course that this only effects the measured crystallinity and as long as the calibration standards are processed under the same conditions as the unknown, the "actual" crystallinity is not significantly affected by the choice in selected analysis interval.
 Results of the AMORPH calculations show good agreement to calibration curves using the manual
-calibration approach (Fig.~\ref{fig:Comparison}). The offset between the AMORPH and manual calibration methods
-in Figure~\ref{fig:Comparison} is consistent with the change in analyzed $2\theta$ range, 10--40\degree~and 5--50\degree,
-respectively (Fig.~\ref{fig:Outputs}). Since the wide peak for the amorphous material does not extend
-appreciably out of the range from 10--40\degree~$2\theta$ (for a Cu K$\alpha$ source), peak integration
-outside of this range will decrease the relative contribution of the amorphous material
-(only peaks for crystalline materials are outside this range).
-Experimentation demonstrates also that data processing only over the 10--40\degree~$2\theta$ range is also significantly faster, as this reduces the total number of data points for processing.
-Thus we recommend that the optimal range for data processing in terms of time and accuracy is from 10--40\degree~$2\theta$.
+calibration approach for the same set of standards analyzed on the same PANalytical Empyrean X-ray diffractometer (Fig.~\ref{fig:Comparison}).
 
@@
 \includegraphics[width=0.7\textwidth]{figures/instruments.jpg}
-\caption{\it Two calibration curves of the same standard material (rhyolite) are compared from different instrumentation, including a Siemens D500 (Washington State University; red circles) and a PANalytical Empyrean (University of Auckland; blue diamonds).\label{fig:intruments}}
-\end{figure}
-
-\begin{figure}[!ht]
-\centering
-\includegraphics[width=0.7\textwidth]{figures/Comparison.jpg}
-\caption{\it Comparison of calibration results for rhyolite calculated from AMORPH compared to a calibration calculated using the manual procedure described by \citet{rowe2012}.\label{fig:Comparison}}
+\caption{\it Two calibration curves of the same standard material (rhyolite) are compared from different instrumentation, including a Siemens D500 (Washington State University; red circles) and a PANalytical Empyrean (University of Auckland; blue diamonds).\label{fig:instruments}}
 \end{figure}
@@
 
+For some materials, e.g. allophane, the amorphous region extends beyond the 10--40\degree~$2\theta$ range and thus would require analysis over a larger $2\theta$ interval. 
+The trade-off in extending the $2\theta$ range is increased processing time and further deviation from a 1:1 calibration line.
+Since the analysis time is, in part, a function of the number of data points, if greater range in $2\theta$ is required, decreasing the total number of points (larger step size during analysis or manual removal of points in $2\theta$ regions not of interest) will offset additional analysis time.
+Thus we recommend that the optimal range for data processing in terms of time and accuracy is from 8.8 to 2.2 {\AA} or 10--40\degree~$2\theta$ for a Cu K{$\alpha$} X-ray source (see Table~\ref{tab:anode} for other X-ray sources) for crystallinity measurements of most volcanics.
+Importantly, reducing the data range from 5--50\degree~$2\theta$ to 10--40\degree~$2\theta$, does not increase the error associated with the peak fitting computational process with respect to other sources of error in sample selection and processing.
+In fact, for some analyses the variance in calculated crystallinity is appreciably reduced by shortening the range in analysed $2\theta$ and thereby reducing potential error associated with modeling background positions (Fig ~\ref{fig:Outputs}).
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=0.7\textwidth]{figures/Comparison.jpg}
+\caption{\it Comparison of calibration results for rhyolite calculated from AMORPH compared to a calibration calculated using the manual procedure described by \citet{rowe2012}.\label{fig:Comparison}}
+\end{figure}
+
 \subsection{Amorphous Characteristics}
@@
 independently model characteristics of the amorphous component.  In particular, centre of mass,
-skewness, and nongaussianity are distinctive amongst different amorphous materials. To test
-this, we present amorphous characteristics of rhyolite and basalt glass picked from a
+skewness, and nongaussianity are distinctive amongst different amorphous materials.
+In contrast to crystallinity determinations, the focus of applications looking at the characteristics of the amorphous material are for indentification of the amorphous phase(s) \citep[e.g.,][]{dehouck2014}.
+To test
+the ability of AMORPH to characterise amorphous phases, we present amorphous characteristics of rhyolite and basalt glass picked from a
 Taupo pumice \citep[73.5 wt\% SiO\textsubscript{2}; P2166C;][]{barker2015} and a Kilauea basalt
 \citep[51 wt\% SiO\textsubscript{2}; KS08-108E;][]{wooten2009, rowe2015}, respectively.
-In addition, Mars Scientific Laboratory (MSL) Curiosity rover CheMin instrument reduced X-ray diffraction patterns
-for sediments from the Bagnold Sand Dunes \citep[Gobabeb;][]{achilles2017, lapotre2017}
-and Buckskin \citep{morris2016} localities, both which have significant amorphous
-materials, have been processed for comparison (CheMin instrument data\footnote{\tt http://pds-geosciences.wustle.edu/missions/msl/chmin.htm}). MSL diffraction patterns have been
-recalculated from a Co K$\alpha$ ($1.7890 \dot{A}$) to Cu K$\alpha$ ($1.5406 \dot{A}$) X-ray source using Bragg's law, to be visually comparable to terrestrial
-examples analysed for this study.
-Results are summarized in Figure~\ref{fig:volcanic_glasses}.
 The most notable differences in the amorphous component characteristics calculated for diffraction patterns for basalt and rhyolite glass include a positive skewness and shift to lower centre of mass
-$2\theta$ values for the rhyolite glass compared to the basalt glass. Although, CheMin
+$2\theta$ values for the rhyolite glass compared to the basalt glass (Fig~\ref{fig:terrestrial}).
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=1\textwidth]{figures/terrestrial.jpg}
+\caption{\it Examples of changing characteristics of amorphous material based on composition.
+Kilauea basalt glass (left) and taupo rhyolite glass (right) were analyzed on a PANalytical Empyrean XRD at University of Auckland.
+Calculated statistics for each analysis are provided below the model results.
+Note in particular the change in skewness from basalt to rhyolite.\label{fig:terrestrial}}
+\end{figure}
+
+Although CheMin
 diffraction analyses were run on a different X-ray diffractometer from the terrestrial
 samples shown here, similar relative changes in amorphous characteristics can be observed
-in X-ray diffraction data from Mars.
-In particular, the Buckskin 1 analysis of the amorphous component indicates a positive skewness $\sim 4\times$ greater than in the Gobabeb 2 sample.
-These results suggest a distinct change in the
-composition of the amorphous material. This interpretation, is consistent with prior
+in X-ray diffraction data from Mars (Fig ~\ref{fig:Mars}).
+It is unknown if/how the X-ray source (e.g.
+Cu versus Co) fundamentally changes the peak shape characteristics of the amorphous content in this study.
+However, qualitative assessment suggests similar overall peak shapes between terrestrial samples analysed with a Cu X-ray anode (Fig.~\ref{fig:terrestrial}) and those analysed with a Co X-ray anode (Fig.~\ref{fig:Mars}).
+In particular, the Buckskin 2 analysis of the amorphous component indicates a greater positive skewness than the Gobabeb 2 sample.
+In addition, the nongaussianity of the Buckskin 2 sample is $\sim 3\times$ greater compared to the Gobabeb 2 amorphous component.
+These results suggest a change in the
+composition of the amorphous material between the two Martian samples. This interpretation, is consistent with prior
 published results which suggest that at the Gobabeb locality (Namib Dune, Bagnold Dune Field),
@@
 
-\begin{figure}[!ht]
-\centering
-\includegraphics[width=1\textwidth]{figures/volcanic_glasses.jpg}
-\caption{\it Four examples of changing characteristics of amorphous material based on composition.
-Kilauea basalt glass (upper left) and taupo rhyolite glass (upper right) were analyzed on a PANalytical Empyrean XRD at University of Auckland.
-Gobabeb 2 (lower left) and Buckskin 1 (lower right) results from the CheMin instrument on the MSL Curiosity rover (data sourced as above).
-Calculated statistics for each analysis are provided below the model results.
-Note in particular the change in skewness from Basalt to Rhyolite, and from Gobabeb 2 to Buckskin 1, supporting compositional change.\label{fig:volcanic_glasses}}
-\end{figure}
+While results may imply a qualitative relationship to composition, further testing and an understanding of the program's limitations is necessary to better quantify compositional correlations.
+In particular, the AMORPH program can quantify the peak shape and proportion of amorphous material but it cannot deconvolve the diffraction pattern of multiple amorphous phases, nor can it distinguish between non-crystalline materials and X-ray amorphous phases such as nanophase oxides \citep{blake2013}.
 
@@
 project.
-Data used in Figure~\ref{fig:volcanic_glasses} analyzed and processed by J.
+Data used in Figure~\ref{fig:Mars} analyzed and processed by J.
 Li, M.
@@
 
+\bibitem[Andrade et al.(2017)]{andrade2017}
+Andrade, F.R.D., Polo, L.A., Janasi, V.A., Caralho, M.S.
+2017.
+Volcanic glass in Cretaceous dacites and rhyolites of the Parana Magmatic Province, southern Brazil: Characterization and quantification by XRD-Reitveld.
+Journal of Volcanology and Geothermal Research, doi: 10.1016/j.jvolgeores.2017.08.008.
+
 \bibitem[Barker et al.(2015)]{barker2015}
@@
 Bish, D. L. et al. 2013. X-ray diffraction results from Mars Science Laboratory: mineralogy of Rocknest at Gale Crater. Science, 341, 1238932.
+
+\bibitem[Blake et al.(2012)]{blake2012}
+Blake, D. et al. 2012. Characterization and calibration of the CheMin mineralogical instrument on Mars Science Laboratory. Space Science Reviews, 170, 341-399. doi: 10.1007/s11214-012-9905-1.
 
@@
 
+\bibitem[Skilling(2006)]{skilling2006nested}
+Skilling, J. 2006. Nested sampling for general Bayesian computation.
+Bayesian analysis, 1, 4, 833--859.
+
 \bibitem[Wall et al.(2014)]{wall2014}
@@
 
-\bibitem[Zorn et al.(2017)]{zorn2017}
+\bibitem[Zorn et al.(2018)]{zorn2018}
 Zorn, E.U., Rowe, M.C., Cronin, S.J., Ryan, A.G., Kennedy, L., Russell, J.K.
-2017.
+2018.
 Influence of porosity and groundmass crystallinity on dome rock strength; a case study from Mt.
 Taranaki, New Zealand.
-Bulletin of Volcanology, in review.
+Bulletin of Volcanology, 80:35, doi: 10.1007/s00445-018-1210-8.
 

### 2018-03-25 doc/ms.tex
Minor edits

@@
 {\bf X-ray Anode}      &   {\bf X-ray wavelength}   &  {\bf Control point 2} & {\bf Control point 3}\\
- &   {\bf (\AA)}   &  {\bf(\degree 2 theta)} & {\bf(\degree 2 theta)}\\
+ &   {(\AA)}   &  {\bf(\degree $2\theta$)} & {\bf(\degree $2\theta$)}\\
 \hline
@@
 \end{tabular}
-\caption{\it Typical XRD anodes with recommended background control points based on a predominant 8.8-2.25 {\AA}  amorphous peak.\label{tab:anode}}
+\caption{\it Typical XRD anodes with recommended background control points based on a predominant 8.8--2.25 {\AA}  amorphous peak.\label{tab:anode}}
 \end{table}

### 2018-03-25 doc/ms.tex
Writing in appendix about command line options

@@
 For simplicity, text files for processing should be located in the same folder as the {\tt AMORPH.exe} program.
-The program is set to run until 10,000 saved parameter sets have been generated.
+
+For speed, you can pass the command line options {\tt -o OPTIONS_AGGRESSIVE} and
+{\tt -t 4} to {\tt AMORPH.exe}. The first of these options sets aggressive
+DNest4 parameters, and the second allows the program to use four threads
+simultaneously.
+
+The program is set to run until 10,000 saved parameter sets
+(`particles') have been generated.
 Data outputs may be viewed at any point, however, closing the program before reaching  10,000 will reduce the accuracy of the final calculations.

### 2018-04-16 doc/figures.py
New figures

@@ -81,9 +81,9 @@ f1 = np.exp(-0.5*(x - 25)**2/width**2)
 width = 3 + 3.05*(x > 25)
 f2 = np.exp(-0.5*(x - 25)**2/width**2)
 
-plt.plot(x, f0/f0.max(), alpha=0.6, label="Skewness = 0.0")
-plt.plot(x, f1/f1.max(), alpha=0.6, label="Skewness = 0.25")
-plt.plot(x, f2/f2.max(), alpha=0.6, label="Skewness = 0.50")
+plt.plot(x, f0/f0.max(), "-",  alpha=0.6, label="Skewness = 0.0")
+plt.plot(x, f1/f1.max(), "--", alpha=0.6, label="Skewness = 0.25")
+plt.plot(x, f2/f2.max(), ":",  alpha=0.6, label="Skewness = 0.50")
 plt.xlabel("$x$")
 plt.ylabel("$y$")
 plt.title("Skewness")
@@ -105,9 +105,9 @@ f2 = (1.0 + (x - 30.0)**2/nu/3**2)**(-0.5*(nu + 1.0))
 f2 /= f2.max()
 
 #plt.figure(figsize=(9, 6))
-plt.plot(x, f0, alpha=0.6, label="Non-gaussianity=0")
-plt.plot(x, f1, alpha=0.6, label="Non-gaussianity=0.08")
-plt.plot(x, f2, alpha=0.6, label="Non-gaussianity=0.15")
+plt.plot(x, f0, "-",  alpha=0.6, label="Non-gaussianity=0")
+plt.plot(x, f1, "--", alpha=0.6, label="Non-gaussianity=0.08")
+plt.plot(x, f2, ":",  alpha=0.6, label="Non-gaussianity=0.15")
 plt.xlabel("$x$")
 plt.ylabel("$y$")
 plt.xlim([-5, 55])

### 2018-04-17 doc/ms.tex
Escape _
Many edits for resubmission

--- 2018-03-25 doc/ms.tex
@@
 \newcommand{\todo}{\color{orange} \bf}
-
+\newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society}
+\newcommand{\changed}{\color{blue}}
+\renewcommand{\thefootnote}{\arabic{footnote}}
 % http://support.river-valley.com/wiki/index.php?title=Elsarticle.cls
@@
 
-\title{AMORPH: A statistical program for characterizing amorphous materials by X-ray diffraction}
+\title{AMORPH: A statistical program for characterizing amorphous materials by X-ray diffraction$^1$\footnotetext{\changed $^1$MR conceived the project, developed geologic applications, and drafted the manuscript with BB.
+BB developed the statistical model, wrote the program code, and co-wrote the manuscript.}}
 
@@
 \begin{abstract}
-AMORPH utilizes a new Bayesian statistical approach to interpreting X-ray diffraction results of samples with both crystalline and amorphous components.
-AMORPH fits X-ray diffraction patterns with a mixture of gaussians,
-which allows us to simultaneously infer all of the model parameters and quantify their
+AMORPH utilizes a new Bayesian statistical approach to interpreting X-ray diffraction results of {\changed samples with both crystalline and amorphous components}.
+AMORPH fits X-ray diffraction patterns with a mixture of {\changed narrow and wide components},
+{\changed simultaneously inferring} all of the model parameters and {\changed quantifying} their
 uncertainties.
-The program simulates background patterns previously manually applied, providing reproducible results, and significantly reducing inter- and intra-user biases.
-This approach allows for the quantification of amorphous and crystalline materials and for the characterization of the amorphous component, including parameters of centre of mass, width, skewness, and nongaussianity of the amorphous component.
+The program simulates background patterns previously {\changed applied manually}, providing reproducible results, and significantly reducing inter- and intra-user biases.
+This approach allows for the quantification of amorphous and crystalline materials and for the characterization of the amorphous component, including {\changed properties such as the} centre of mass, width, skewness, and nongaussianity of the amorphous component.
 Results demonstrate the applicability of this program for calculating amorphous contents of volcanic materials and independently modeling their properties in compositionally variable materials.
@@
 \begin{keyword}
-  Amorphous materials \sep X-ray diffraction \sep Bayesian inference \sep Markov Chain Monte Carlo \sep Mixture models \sep crystallinity
+  Amorphous materials \sep X-ray diffraction \sep Bayesian inference \sep Markov chain monte carlo \sep mixture models \sep crystallinity
 \end{keyword}
@@
 also provides a means to investigate amorphous materials \citep[e.g.,][]{rowe2012, achilles2013}.
-X-ray amorphous materials lack long-range crystallographic order and produce broad humps with low intensities in XRD patterns.
-X-ray diffraction of rocks/soils containing amorphous materials (e.g.
+X-ray {\changed amorphous materials lack long-range crystallographic order and produce broad humps with low intensities in XRD patterns.
+X-ray} diffraction of rocks/soils containing amorphous materials (e.g.
 volcanic
@@
 as it reduces the observed intensity of the amorphous material thus reducing the accuracy of
-pattern matching to determine abundances. An Excel based adaptation of this pattern matching
+pattern matching to determine abundances. An Excel-based adaptation of this pattern matching
 technique \citep[FULLPAT;][]{chipera2002} is routinely utilized for characterizing remotely acquired X-ray
-diffraction patterns, such as those measured by the CheMin instrument on the Mars Science Laboratory (MSL) Curiosity rover \citep{blake2012, bish2013}.
+diffraction patterns, such as those {\changed measured by the CheMin instrument on} the Mars Science Laboratory (MSL) Curiosity rover \citep{blake2012, bish2013}.
 The problem faced by RIR approaches is they are based on modeling of an observed
@@
 assessment of the characteristics of the amorphous material.
-Furthermore, natural X-ray amorphous phases have a wide variety of compositions such that the library measurements may not include the amorphous phase(s) present in the sample.
-In comparison, the calibration method
-focuses solely on the quantification of the amorphous material, utilizing the integrated counts
-associated with the amorphous and crystalline componentry.
-
-\begin{align}
-C_{\rm m} \% = [C_{\rm pa}]/[A_{\rm pa} + C_{\rm pa}] \times 100
+{\changed Furthermore, natural X-ray amorphous phases have a wide variety of compositions such that the library measurements may not include the amorphous phase(s) present in the sample.} In comparison, the calibration method
+focuses solely on the quantification of the amorphous material. {\changed This method utilizes} the integrated counts
+associated with the amorphous and crystalline componentry:
+
+\begin{align}
+C_{\rm m} \% = [C_{\rm pa}]/[A_{\rm pa} + C_{\rm pa}] \times 100\label{eqn:definition}
 \end{align}
@@
 point of the calibration curve is to reduce the inter-user variation, determinations still tend to
-be associated with moderate user error (3-5 vol \%). As every user must independently practice
+be associated with moderate user error (3--5 vol \%). As every user must independently practice
 background fittings using the calibration method, the results ultimately are only as good as the
-user’s ability to manually model the data. This technique, while not constrained by a
+user’s ability to manually {\changed fit} the data. This technique, while not constrained by a
 necessary library of diffraction patterns, also does not characterize the XRD properties of the
@@
 This paper is structured as follows. In Section~\ref{sec:bayes}, we outline the Bayesian
-framework of inference in general terms, and in Section~\ref{sec:model} we describe the
-specific model assumptions made by AMORPH (i.e., the mixture of gaussians), and
-Section~\ref{sec:priors} lists the probabilistic assumptions (i.e., the prior distributions) of the program. Section~\ref{sec:applications} presents some
+framework of inference in general terms. In Section~\ref{sec:model} we describe the
+specific modelling assumptions made by AMORPH, and
+Section~\ref{sec:priors} lists the probabilistic assumptions (i.e., the prior distributions)
+{\changed assumed by} the program. Section~\ref{sec:applications} presents some
 example applications and comparisons with standard techniques, and finally we conclude
@@
 Section~\ref{sec:model}, and the computation is performed with
-Diffusive Nested Sampling, an advanced MCMC method \citep{dns, dnest4}.
-There have been many applications of Bayesian inference to
+Diffusive Nested Sampling {\changed \citep[an advanced MCMC method][]{dns}
+as implemented in the software package DNest4 \citep{dnest4}.
+This method is robust to complicated posterior distributions.
+
+There have been many approaches to
 problems of separating background from compact signals
-when noise is present \citep[e.g.][]{}. AMORPH can be viewed as another
-case of this, with assumptions customized to the specific application.
+when noise is present, such as the famous CLEAN algorithm \citep{clean},
+through to more explicitly Bayesian approaches
+\citep[e.g.][]{padayachee, massinf, object, hobson, worpel}.
+AMORPH can be viewed as another
+case of this, with assumptions customized to this specific application.}
 
@@
 (the probability distribution for $\params$ {\em given} prior information
-$\info$). Throughout this paper, $\params$ is used to denote a
+$\info$). Throughout this paper, the bold $\params$ is used to denote a
 collection of unknown parameters, whereas $2\theta$ (with the factor of two, and the
@@
 if we knew the parameters (and the prior information), as a function of
-the parameters. This describes uncertainty about what data will be observed,
+the parameters. {\changed The sampling distribution} describes uncertainty about what data will be observed,
 but encodes some knowledge about the fact that the data have something to do with the
@@
 \end{align}
-which takes the specific data $\data_{\rm obs}$ into account. With the specific dataset,
+which takes the specific data $\data_{\rm obs}$ into account {\changed via conditioning}.
+With the specific dataset,
 $p(\data_{\rm obs} | \params,\info)$ becomes a function of the
@@
 best fitting values of the parameters (i.e., a maximum likelihood estimate),
-Bayesian inference via MCMC involves finding a sampling of plausible values
-for the parameters. Since AMORPH uses \citet{dnest4}, it also calculates
+Bayesian inference via MCMC involves {\changed exploring a range of plausible values
+for the parameters. Since AMORPH uses Diffusive Nested Sampling, it also calculates
 the `marginal likelihood' or `evidence' quantity $p(\data | \info)$, which
@@
 
-% BJB: Hopefully this addresses reviewer 1, point a.
-The computational cost of running AMORPH on a data set of around 1000
-points varies, as the posterior distribution depends on the data.
+The computational cost of running AMORPH
+varies, as the posterior distribution depends on the data.
+The size of the dataset is important, of course.
 Another important factor is the number of MCMC samples desired.
-On a modern PC, some useful posterior samples can be obtained in about
-half an hour.
-With 300 posterior samples, typical analyses take approximately 7 hours to complete, depending on computing capabilities and the number of data points.
+To quantify the uncertainty about a few scalar parameters, about
+100 posterior samples is usually sufficient. Indeed, \citet[][p. 380]{mackay2003} suggests
+that only about twelve are needed for the uncertainty about a single
+quantity!
+However, if one wants to produce
+smooth histograms or look at details of the posterior distribution,
+many more samples are required.
+
+On a modern desktop PC and using aggressive numerical settings
+(see~\ref{sec:program} for details),
+some useful posterior samples can be obtained in about
+15 minutes when the dataset contains about 1500 points.
+Typical analyses take approximately 5 hours to fully complete and yield
+a few hundred posterior samples.
 However, the complete runs presented in this paper took
-about a day each to run, as we wanted to ensure the results contained
-about 1000 posterior samples.
+about a day each to run, as we ran with more conservative settings
+to ensure the number of posterior samples was about 1000 for each dataset.}
 
@@
 \end{align}
-and the measured amorphous content ($A_{\rm m}$ or $1-C_{\rm m}$) is defined, similar to Equation 1, as the proportion of the area in the
+and the measured amorphous content ($A_{\rm m}$ or $1-C_{\rm m}$) is defined,
+similar to Equation~\ref{eqn:definition}, as the proportion of the area in the
 amorphous and crystalline components which is accounted for by the amorphous component:
@@
 From each sample from the posterior distribution for all of the parameters,
-it is possible to compute the proportion of amorphous material, making it straightforward to
-calculate the marginal posterior distribution of the amorphous content, describing
-the remaining uncertainty about its value in the light of the data.
+it is straightforward to compute the proportion of amorphous material, and hence
+calculate the marginal posterior distribution of the amorphous content.
 In the following subsections, we describe the specific parameterizations of the
@@
 and between $\x=40$\degree~and $\x=\x_{\rm max}$}.
-These values are selected to maintain a linear background over the interval from 10--40\degree, the predominant range of amorphous materials (8.8 -- 2.25 \AA) analyzed using a Cu K$\alpha$ x-ray source (Table~\ref{tab:anode}), while still allowing for some flexibility in the structure of the background outside this range.
+These values are selected to maintain a linear background over the interval from 10--40\degree, the predominant range of amorphous materials {\changed (8.8 -- 2.25 \AA)} analyzed using a Cu K$\alpha$ x-ray source {\changed (Table~\ref{tab:anode})}, while still allowing for some flexibility in the structure of the background outside this range.
 
@@
 \end{align}
-Control points may be modified, depending on the X-ray source in use and changes in the required range for the linear background as a function of the nature of the amorphous material (Table~\ref{tab:anode}).
-The value of the background component at any other point is found by
+{\changed Control points may be modified by editing the file {\tt control\_points.in},
+depending on the X-ray source in use and changes in the required range for the linear background as a function of the nature of the amorphous material (Table~\ref{tab:anode}).} The value of the background component at any other point is found by
 linear interpolation of these control points.
@@
 \centering
+{\changed
 \begin{tabular}{|llll|}
@@
 \caption{\it Typical XRD anodes with recommended background control points based on a predominant 8.8--2.25 {\AA}  amorphous peak.\label{tab:anode}}
+} % \changed
 \end{table}
@@
 \subsection{The crystalline component}
-
-% BJB: Reponse to Reviewer 2, comment about Lorentzian shape vs Gaussian
-% We experimented with the Lorentzian shape instead of the Gaussian and found
-% that for same datasets, it was preferred, whereas for other datasets, the
-% Gaussian is preferred (as adjudicated by the marginal likelihood values
-% for the two different assumptions). Rather than having to run each model
-% twice, we decided it would be more elegant to replace the gaussian assumption
-% with some family of shapes that can smoothly interpolate between gaussian
-% and lorentzian as necessary. Specifically, we chose the Student's
-% t-distribution which can do this straightforwardly. This introduces one
-% more parameter to the model, which controls the shape of the peaks.
-
-% I (BJB) also changed the parameterisation of the heights of the peaks, so
-% they are described by the total integrated mass rather than the peak
-% amplitude. This makes the Metropolis exploration a bit more efficient.
-
 We assume that there is some number $N_{\rm xtal}$ of narrow peaks
-due to crystalline material,
-each of which has the same shape. The shapes of the narrow
+{\changed due to crystalline material},
+each of which has {\changed the same shape. The shapes of the narrow
 peaks are controlled by a single shape parameter $\nu_{\rm xtal}$,
 assumed common to all of the narrow peaks. The chosen shape is based
-on the Student's $t$-distribution, which allows to peaks to be heavy-tailed
+on the Student's $t$-distribution, which allows the peaks to be heavy-tailed
 if $\nu_{\rm xtal}$ is low, and to be Lorentzian or Cauchy in shape if
 $\nu_{\rm xtal}$ is exactly 1. As $\nu_{\rm xtal}$ is high, the shape
-approximates a gaussian.
+approximates a gaussian.}
 
@@
 by its central position $X_{i, \rm xtal}$,
-total integrated `mass' $M_{i, \rm xtal}$, and width $W_{i, \rm xtal}$.
-The parameter vector for crystalline component is
+{\changed total integrated `mass' $M_{i, \rm xtal}$}, and width $W_{i, \rm xtal}$.
+The parameter vector for crystalline component is {\changed
 \begin{align}
@@
     W_{1, \rm xtal}, ..., W_{N, \rm xtal}\}
-\end{align}
+\end{align} } % changed
 and the total shape contributed by the
-crystalline component is a sum of the $N_{\rm xtal}$ peaks:
+crystalline component is a sum of the $N_{\rm xtal}$ {\changed peaks:
 \begin{align}
@@
 For computational efficiency, evaluation of Equation~\ref{eqn:t_shape}
-is performed using a lookup table.
+is performed using a lookup table. In addition, we evaluate the peak
+shapes at the $x$-position of each data point, rather than integrating
+the peak shape across a finite bin width. This is faster but must be
+kept in mind when interpreting parameters affected by this assumption,
+such as the widths of the narrow peaks. } %changed
 
 \subsection{The amorphous component}
-The amorphous component is also composed of a sum of peaks, which we chose to be
+The amorphous component is also composed of {\changed a sum of peaks, which we chose to be
 gaussians. The parameterization is the same as for the crystalline peaks but with a fixed,
@@
 dataset, and to be wider.
+The constrainted prior for the positions of the wide amorphous
+peaks prevents the program from attempting to explain background
+variations near the edge of the data using the amorphous component.}
 
@@
 by its central position $X_{i, \rm amorph}$,
-mass (total integral) $M_{i, \rm amorph}$, and width $W_{i, \rm amorph}$.
-The parameter vector for wide gaussians is
-\begin{align}
-\params_{\rm amorph} &=
+{\changed mass (total integral) $M_{i, \rm amorph}$}, and width $W_{i, \rm amorph}$.
+The parameter vector for the wide gaussians is {\changed
+\begin{align}
+\params_{\rm amorph} =
   \{N_{\rm amorph}, X_{1, \rm amorph}, ..., X_{N, \rm amorph},
-    M_{1, \rm amorph}, ..., M_{N, \rm amorph},
-    W_{1, \rm amorph}, ..., W_{N, \rm amorph}\}
-\end{align}
+    M_{1, \rm amorph}, ..., M_{N, \rm amorph},\\
+   \quad\quad W_{1, \rm amorph}, ..., W_{N, \rm amorph}\}
+\end{align} } % changed
 and the total shape contributed by the
-amorphous component is a sum of the $N_{\rm amorph}$ gaussians:
+amorphous component is a sum of the $N_{\rm amorph}$ gaussians: {\changed
 \begin{align}
@@
  \exp\left[-\frac{1}{2W_{i, \rm amorph}^2}\left(x - X_{i, \rm amorph}\right)^2\right].
-\end{align}
-
-Figure~\ref{fig:wide_component} shows several examples of plausible amorphous component shapes. The prior distributions (specified in
+\end{align} } % changed
+
+Figure~\ref{fig:wide_component} shows several examples of possible amorphous component shapes. The prior distributions (specified in
 Section~\ref{sec:priors})
 constrain the central positions of these gaussians much
-more strongly than the central positions of the narrow peaks (crystalline component), which is how AMORPH is able to distinguish the crystalline
-component from the amorphous component.
+more strongly than the central positions of the narrow peaks (crystalline component). AMORPH is primarily able to distinguish the crystalline
+{\changed and amorphous components due to the different priors for the widths
+of the two different sets of peaks.}
 
@@
 \subsection{Derived properties}
-Some derived properties of the amorphous component are used to summarize its shape.
-We have included `skewness', which describes asymmetry, and `nongaussianity'
+Some derived properties of the amorphous component can be used to summarize its shape.
+We have included {\changed the `width' (second central moment, analogous to the standard
+deviation)}, `skewness', which describes asymmetry, and `nongaussianity'
 which quantifies the departure from gaussianity. These are not the same
@@
 
-Defining the normalized version of the amorphous component as
+Defining a normalized version of the amorphous component as
 \begin{align}
@@
 \end{align}
-its center of mass by
-\begin{align}
-\textnormal{center} &= \int_{-\infty}^\infty x g(\x) \, d\x,
-\end{align}
-and its width (second central moment, perhaps better interpreted
-as a half-width) by
+its center of mass is
+\begin{align}
+\textnormal{center} &= \int_{-\infty}^\infty x g(\x) \, d\x.
+\end{align}
+
+The width of the amorphous component
+(its second central moment, perhaps better interpreted
+as a half-width) is defined by
 \begin{align}
@@
 \end{align}
-the skewness is
+and the skewness is
 \begin{align}
@@
 the data must be specified.
-When the number of parameters is large, this is often done `hierarchically', by
+When the number of parameters is large, this is often done hierarchically, by
 specifying the priors conditional on some `hyperparameters', and then assigning
@@
 
-The joint prior distribution for
-the hyperparameters, parameters, and data,
+{\changed In these cases,
+the joint prior distribution for
+the hyperparameters $\hypers$}, parameters, and data,
 is written $p(\hypers, \params, \data | \info)$, and is usually factorized
@@
 \end{align}
-where the first step is true in general by the product rule, and the second
+where the first step is true in general by the product rule
+{\changed of probability}, and the second
 step assumes that knowing the parameters would make the hyperparameters
@@
 
-\begin{landscape}
-
-\begin{table}
+\begin{table}[!ht]
 \footnotesize
@@
 {\em Prior information}&&\\
-\hline
 $n$ & Number of measurements & given\\
 $\{\x_1, \x_2, ..., \x_n\}$  & $\x$-values of measurements & given \\
+{\changed $x_{\rm range}$} & {\changed Range of measurements} & {\changed $\max(x_1, ..., x_n) - \min(x_1, ..., x_n)$} \\
 \hline
@@
 $N_{\rm amorph}$   &   Number of wide peaks    &  $\propto 1/(N_{\rm amorph}+1)$,
-$N_{\rm amorph} \in \{0, 1, ..., 1000\}$ \\
-$a_{\ln M, \rm xtal}$ & Typical log-mass of narrow peaks & Laplace(0, 5)\\
-$a_{\ln M, \rm amorph}$ & Typical log-mass of wide peaks & Laplace(0, 5)\\
-$b_{\ln M, \rm xtal}$ & Diversity of log-mass of narrow peaks & Uniform(0, 5)\\
-$b_{\ln M, \rm amorph}$ & Diversity of log-mass of wide peaks & Uniform(0, 5)\\
+ {\changed $N_{\rm amorph} \in\{0, 1, ..., 1000\}$} \\
+{\changed $a_{\ln M, \rm xtal}$} & Typical {\changed log-mass} of narrow peaks & Laplace(0, 5)\\
+{\changed $a_{\ln M, \rm amorph}$} & Typical {\changed log-mass} of wide peaks & Laplace(0, 5)\\
+{\changed $b_{\ln M, \rm xtal}$} & Diversity of {\changed log-mass} of narrow peaks & Uniform(0, 5)\\
+{\changed $b_{\ln M, \rm amorph}$} & Diversity of {\changed log-mass} of wide peaks & Uniform(0, 5)\\
 $a_{\ln W, \rm xtal}$ & Typical log-width of narrow peaks & LogUniform$(10^{-3}x_{\rm range}, 0.05x_{\rm range})$\\
@@
                             Uniform$(x_{\rm min}, x_{\rm max})$ \\
-$A_{i, \rm xtal}$ & Masses of narrow peaks &
- $\ln M_{i, \rm xtal} \sim \textnormal{Laplace}\left(a_{\ln M, \rm xtal}, b_{\ln M, \rm xtal}\right)$ \\
+$A_{i, \rm xtal}$ & {\changed Masses} of narrow peaks &
+{\changed
+ $\ln M_{i, \rm xtal} \sim \textnormal{Laplace}\left(a_{\ln M, \rm xtal}, b_{\ln M, \rm xtal}\right)$} \\
 $W_{i, \rm xtal}$ & Widths of narrow peaks &
  $\ln W_{i, \rm xtal} \sim \textnormal{Laplace}\left(a_{\ln W, \rm xtal}, b_{\ln W, \rm xtal}\right)$ \\
-$\nu_{\rm xtal}$ & Shape of narrow peaks (common) & LogUniform$(0.1, 100)$\\
+{\changed $\nu_{\rm xtal}$} &  {\changed Shape of narrow peaks (common)} &  {\changed LogUniform$(0.1, 100)$}\\
 \hline
@@
           Uniform$(x_{\rm min}+0.3x_{\rm range}, x_{\rm max}-0.3x_{\rm range})$ \\
-$A_{i, \rm amorph}$ & Masses of wide peaks &
- $\ln M_{i, \rm amorph} \sim \textnormal{Laplace}\left(a_{\ln M, \rm amorph}, b_{\ln M, \rm amorph}\right)$ \\
+{\changed $M_{i, \rm amorph}$} &  {\changed Masses} of wide peaks &
+ {\changed  $\ln M_{i, \rm amorph} \sim \textnormal{Laplace}\left(a_{\ln M, \rm amorph}, b_{\ln M, \rm amorph}\right)$} \\
 $W_{i, \rm amorph}$ & Widths of wide peaks &
@@
 $\sigma_1$ &    Noise proportional to $\sqrt{\textnormal{model}}$   &  $\ln \sigma_1 \sim \textnormal{Laplace}(0,5)$ \\
-$\nu$     &   Shape parameter for $t$-distribution for noise   &   LogUniform$(1, 1000)$\\
+$\nu$     &   Shape parameter for $t$-distribution&LogUniform$(1, 1000)$\\ &for noise   &   \\
 \hline
 {\em Data}&&\\
-\hline
-$\{y_1, y_2, ..., y_n\}$  &   Measurements    & independent Student-$t(f(x_i; \params), \sqrt{\sigma_0^2 + \sigma_1f(x_i; \params)},\nu)$\\
+$\{y_1, y_2, ..., y_n\}$  &   Measurements    & independent \\ && Student-$t\left(f(x_i; \params), \sqrt{\sigma_0^2 + \sigma_1f(x_i; \params)},\nu\right)$\\
 \hline
@@
 
-\end{landscape}
-
 Most of the chosen distributions were intended to be reasonably
-conservative and wide.
+conservative and wide, {\changed but not excessively so}.
 Some of the prior distributions in AMORPH use the log-uniform distribution,
 sometimes (erroneously) called a Jeffreys prior, whose density for a quantity
-$z$ is proportional to $1/z$.
+$z$ is proportional to $1/z$. {\changed This prior is equivalent to a uniform prior
+for the logarithm of the quantity, and encodes the fact that the order of magnitude
+of a positive quantity may be unknown.}
 
@@
 a more conventional normal distribution, because it assigns higher
-prior probability to values far from the center, and can be computed conveniently, i.e., no special functions such as the inverse of the
-error function are needed.
-The Laplace, or biexponential, distribution with location parameter
+prior probability to values far from the center  {\changed (compared to a normal
+distribution)}, and can be computed conveniently. Specifically,
+special functions such as the inverse of the error function are needed.
+The Laplace distribution with location parameter
 (central position) $a$ and scale parameter (width) $b$ has probability
@@
 The priors for the positions and widths of the `amorphous' and `crystalline'
-gaussians are different, to account for the prior knowledge that they
-should result in wide gaussians near the centre of the domain, and
-narrow gaussians throughout the domain respectively.
-
-The prior for the data given all the parameters is a student-$t$ distribution,
-which allows for potentially heavier tails than a normal distribution would.
+{\changed peaks} are different, to account for the prior knowledge that they
+should result in wide {\changed peaks} near the centre of the domain, and
+narrow {\changed peaks} throughout the domain respectively.
+
+The prior for the data given all the parameters  {\changed (i.e., the sampling distribution)}
+is a student-$t$ distribution,  {\changed centered at the model curve $f(x; \boldsymbol{\theta})$}.
+{\changed The $t$ distribution} allows for potentially heavier tails than a normal distribution would.
 Thus, analyses with AMORPH should be able to cope with discrepant datapoints
@@
 
-AMORPH does not assume a known number of gaussians for either the
-amorphous or the crystalline component. The appropriate number of
-gaussians is inferred from the data. The maximum number of gaussians
-in each component is set to 300.
+AMORPH does not assume a known number of {\changed peaks} for either the
+amorphous or the crystalline component --- the appropriate number of
+{\changed peaks} is inferred from the data. The maximum number of {\changed peaks}
+for each component is set to {\changed 1000}.
 
@@
 \subsection{Crystallinity}
-To date, applications of crystallinity determinations via XRD methodologies on terrestrial samples have primarily focused on volcanic materials \citep[e.g.,][]{wall2014, ellis2015, andrade2017, zorn2018}, linking crystallinity of volcanic glass to timescales of geologic processes.
+{\changed To date, applications of crystallinity determinations via XRD methodologies on terrestrial samples have primarily focused on volcanic materials \citep[e.g.,][]{wall2014, ellis2015, andrade2017, zorn2018}, linking crystallinity of volcanic glass to timescales of geologic processes.
 The primary advantage of this technique being its ability to determine crystallinities for finer grain sizes than typically possible through standard petrologic techniques \citep{rowe2012}.
-In contrast, planetary science usages for determining crystallinity have largely focused on quantification of the amorphous and crystalline components for mass balance determination of physical and chemical components in diverse geologic materials \citep[e.g.,][]{blake2013, dehouck2014}.
-
-The AMORPH program calculates a fraction of amorphous material ($A_{\rm m}$) for each posterior sample of the parameters,
-using Equation~\ref{eqn:crystallinity}.Thus measured crystallinity ($C_{\rm m}$) is easily determined from the relationship $C_{\rm m}$ = 1-$A_{\rm m}$, as defined in Equation 5. 
-To test the applicability of AMORPH crystallinity results we have analysed the Mars Scientific Laboratory (MSL)  CheMin instrument X-ray diffraction patterns of loose basaltic sand from the Bagnold Sand Dunes \citep[Gobabeb;][]{achilles2017, lapotre2017}
+In contrast, planetary science usages for determining crystallinity have largely focused on quantification of the amorphous and crystalline components for mass balance determination of physical and chemical components in diverse geologic materials \citep[e.g.,][]{blake2013, dehouck2014}}.
+
+The AMORPH program calculates a fraction of amorphous material ($A_{\rm m}$) for each posterior sample of the parameters
+using Equation~\ref{eqn:crystallinity}. Thus, measured crystallinity ($C_{\rm m}$) is easily determined from the relationship $C_{\rm m} = 1-A_{\rm m}$.
+{\changed To test the applicability of AMORPH crystallinity results, we have analysed the Mars Scientific Laboratory (MSL)  CheMin instrument X-ray diffraction patterns of loose basaltic sand from the Bagnold Sand Dunes \citep[Gobabeb;][]{achilles2017, lapotre2017}
 and sediments drilled from a silicic mudstone \citep[Buckskin sample;][]{morris2016}, both which have significant amorphous
@@
 This recalculation simply allows for an easier visual comparison to terrestrial samples analysed with the more standard Cu K$\alpha$ X-ray source (see below).
-Note that for Co K$\alpha$ results, background critical points have been adjusted to 11.7-46.9 $2\theta$ to maintain the linear background over the amorphous region from ~8.8 to 2.25 $\dot{A}$ (Table ~\ref{tab:anode}).
+Note that for Co K$\alpha$ results, background critical points have been adjusted to 11.7--46.9\degree~$2\theta$ to maintain the linear background over the amorphous region from $\sim$ 8.8 to 2.25 $\dot{A}$ (Table ~\ref{tab:anode}).
 Although the results are uncalibrated, measured crystallinites provide a good first approximation of actual crystallinity, and relative differences are robust.
@@
 Comparatively, published FULLPAT calculations of amorphous contents are ~35 $\pm$ 15$\%$ for Gobabeb and 50 $\pm$ 15$\%$ for Buckskin \citep{achilles2017, morris2016}.
+}
 
@@
 \includegraphics[width=1\textwidth]{figures/Mars.jpg}
-\caption{\it Analysis of Gobabeb 2 (left) and Buckskin 2 (right) results from the CheMin instrument on the MSL Curiosity rover (data sourced as above).
+\caption{\it \changed Analysis of Gobabeb 2 (left) and Buckskin 2 (right) results from the CheMin instrument on the MSL Curiosity rover (data sourced as above).
 Upper panels  illustrate initial Co K$\alpha$ processed results while lower panels show AMORPH results after translating data to a Cu K$\alpha$ source (see text for details).
@@
 
-The measured crystallinity is not equivalent to an "actual" or "true" crystallinity, although the values are often very close to one another as noted above.
-Therefore, this procedure does not negate the need for calibration standards
-to translate measured crystallinity to an actual crystallinity.Calibration standards can be easily created by mixing
+{\changed The measured crystallinity is not equivalent to an ``actual'' or ``true'' crystallinity, although the values are often very close to one another as noted above.
+Therefore,} this procedure does not negate the need for calibration standards
+to translate measured crystallinity to an actual crystallinity.
+{\changed Calibration standards can be easily created by mixing
 powdered minerals (e.g. olivine, pyroxene, plagioclase) with powdered glass in known
-proportions to simulate volcanic material of varying crystallinities \citep[e.g.,][]{rowe2012, wall2014}.
+proportions to simulate volcanic material of varying crystallinities \citep[e.g.,][]{rowe2012, wall2014}.}
 
@@
 At low crystallinity ($\lesssim 10\%$) diffraction patterns are dominated by the low amplitude amorphous peak such that removing the effects of background noise becomes problematic, resulting in overestimated crystallinities.
-At high crystallinity ($\gtrsim 90\%$) the problem arises that peak overlaps in multi-mineral assemblages cause apparent broadening effects, and may be subsequently modeled as ``amorphous'' and thus crystallinities  underestimated.
+At high crystallinity ($\gtrsim 90\%$) the problem arises that peak overlaps in multi-mineral assemblages cause {\changed apparent} broadening effects, and may be subsequently modeled as ``amorphous'' and thus crystallinities  underestimated.
 The calibration curve therefore is utilized to remove these effects.
@@
 
-Since the wide peak for many amorphous materials does not extend
-appreciably out of the range from 10--40\degree~$2\theta$ (for a Cu K$\alpha$ source), peak area integration
-outside of this range will decrease proportionally the relative contribution of the amorphous material
-(only peaks for crystalline materials are outside this range in the example depicted in Figure ~\ref{fig:Outputs}).
-It should be noted of course that this only effects the measured crystallinity and as long as the calibration standards are processed under the same conditions as the unknown, the "actual" crystallinity is not significantly affected by the choice in selected analysis interval.
+{\changed Since the wide peak for many} amorphous materials does not extend
+appreciably out of the range from 10--40\degree~$2\theta$ (for a Cu K$\alpha$ source), peak {\changed area} integration
+outside of this range will decrease {\changed proportionally} the relative contribution of the amorphous material
+(only peaks for crystalline materials are outside this range {\changed in the example depicted in Figure~\ref{fig:Outputs}, with the corresponding fitting residuals and the posterior distributions
+for the number of narrow peaks in Figure~\ref{fig:residuals_etc} and Figure~\ref{fig:num_peaks}, respectively).
+It should be noted of course that this only effects the measured crystallinity and as long as the calibration standards are processed under the same conditions as the unknown, the ``actual'' crystallinity is not significantly affected by the choice in selected analysis interval.
 Results of the AMORPH calculations show good agreement to calibration curves using the manual
-calibration approach for the same set of standards analyzed on the same PANalytical Empyrean X-ray diffractometer (Fig.~\ref{fig:Comparison}).
+calibration approach for the same set of standards analyzed on the same PANalytical Empyrean X-ray diffractometer (Fig.~\ref{fig:Comparison}).}
 
@@
 \includegraphics[width=1\textwidth]{figures/Outputs.jpg}
-\caption{\it Model results for a basalt standard containing 80\% amorphous glass.
-Data was processed from 10--40\degree~$2\theta$ (upper left panel) with a calculated amorphous content of 76.4\% (upper right panel).
-The same data was processed instead from 5--50\degree~$2\theta$ (lower left panel) with a calculated amorphous content of 71.5\% (lower right panel).
-Rationale for differences are discussed in the text.\label{fig:Outputs}}
-\end{figure}
-
-For some materials, e.g. allophane, the amorphous region extends beyond the 10--40\degree~$2\theta$ range and thus would require analysis over a larger $2\theta$ interval. 
+\caption{\it \changed Model results for a basalt standard containing 80\% amorphous glass.
+Data was processed from 5--50\degree~$2\theta$ (upper left panel) with a calculated amorphous content of 68.9\% (upper right panel).
+The same data was processed instead from 10--40\degree~$2\theta$ (lower left panel) with a calculated amorphous content of 72.2\% (lower right panel).
+The reasons for the
+differences are discussed in the text.\label{fig:Outputs}}
+\end{figure}
+
+\begin{figure}[!ht]
+  \begin{minipage}{\textwidth}
+    \centering
+    \includegraphics[width=0.99\textwidth]{figures/residuals2.pdf} \\
+    \includegraphics[width=0.99\textwidth]{figures/residuals1.pdf}
+  \end{minipage}
+\caption{\it \changed Residuals of the model fits shown in Figure~\ref{fig:Outputs}, showing that the features of the data have been captured (upper panel is 5--50\degree~$2\theta$; lower panel is 10--40\degree~$2\theta$ dataset).
+Technically, there is no single set of
+residuals, but a posterior distribution over them, represented by
+the residuals of each curve in the generated sample. Therefore, these
+plots show the residuals of the posterior mean model curve.\label{fig:residuals_etc}}
+\end{figure}
+
+\begin{figure}[!ht]
+  \begin{minipage}{\textwidth}
+    \centering
+    \includegraphics[width=0.99\textwidth]{figures/numpeaks2.pdf} \\
+    \includegraphics[width=0.99\textwidth]{figures/numpeaks1.pdf}
+  \end{minipage}
+\caption{\it \changed Posterior distributions for the number of narrow peaks for the fits shown in Figure~\ref{fig:Outputs} (upper panel is 5--50\degree~$2\theta$; lower panel is 10--40\degree~$2\theta$ dataset).
+A minimum number of peaks is required in both cases, but the data cannot rule out large numbers of peaks (of very small amplitude).
+There are more peaks in the
+5--50\degree case.\label{fig:num_peaks}}
+\end{figure}
+
+{\changed For some materials, e.g. allophane, the amorphous region extends beyond the 10--40\degree~$2\theta$ range and thus would require analysis over a larger $2\theta$ interval. 
 The trade-off in extending the $2\theta$ range is increased processing time and further deviation from a 1:1 calibration line.
@@
 Importantly, reducing the data range from 5--50\degree~$2\theta$ to 10--40\degree~$2\theta$, does not increase the error associated with the peak fitting computational process with respect to other sources of error in sample selection and processing.
-In fact, for some analyses the variance in calculated crystallinity is appreciably reduced by shortening the range in analysed $2\theta$ and thereby reducing potential error associated with modeling background positions (Fig ~\ref{fig:Outputs}).
+In fact, for some analyses the variance in calculated crystallinity is appreciably reduced by shortening the range in analysed $2\theta$ and thereby reducing potential error associated with modeling background positions (Fig ~\ref{fig:Outputs}).}
 
@@
 \includegraphics[width=0.7\textwidth]{figures/Comparison.jpg}
-\caption{\it Comparison of calibration results for rhyolite calculated from AMORPH compared to a calibration calculated using the manual procedure described by \citet{rowe2012}.\label{fig:Comparison}}
+\caption{\it \changed Comparison of calibration results for rhyolite calculated from AMORPH compared to a calibration calculated using the manual procedure described by \citet{rowe2012}.\label{fig:Comparison}}
 \end{figure}
@@
 skewness, and nongaussianity are distinctive amongst different amorphous materials.
-In contrast to crystallinity determinations, the focus of applications looking at the characteristics of the amorphous material are for indentification of the amorphous phase(s) \citep[e.g.,][]{dehouck2014}.
+{\changed In contrast to crystallinity determinations, the focus of applications looking at the characteristics of the amorphous material are for indentification of the amorphous phase(s) \citep[e.g.,][]{dehouck2014}.
 To test
-the ability of AMORPH to characterise amorphous phases, we present amorphous characteristics of rhyolite and basalt glass picked from a
+the ability of AMORPH to characterise amorphous phases}, we present amorphous characteristics of rhyolite and basalt glass picked from a
 Taupo pumice \citep[73.5 wt\% SiO\textsubscript{2}; P2166C;][]{barker2015} and a Kilauea basalt
 \citep[51 wt\% SiO\textsubscript{2}; KS08-108E;][]{wooten2009, rowe2015}, respectively.
-The most notable differences in the amorphous component characteristics calculated for diffraction patterns for basalt and rhyolite glass include a positive skewness and shift to lower centre of mass
-$2\theta$ values for the rhyolite glass compared to the basalt glass (Fig~\ref{fig:terrestrial}).
+{\changed The most notable differences in the amorphous component characteristics calculated for diffraction patterns for basalt and rhyolite glass include a positive skewness and shift to lower centre of mass
+$2\theta$ values for the rhyolite glass compared to the basalt glass (Fig~\ref{fig:terrestrial}).}
 
@@
 \includegraphics[width=1\textwidth]{figures/terrestrial.jpg}
-\caption{\it Examples of changing characteristics of amorphous material based on composition.
+\caption{\it \changed Examples of changing characteristics of amorphous material based on composition.
 Kilauea basalt glass (left) and taupo rhyolite glass (right) were analyzed on a PANalytical Empyrean XRD at University of Auckland.
@@
 samples shown here, similar relative changes in amorphous characteristics can be observed
-in X-ray diffraction data from Mars (Fig ~\ref{fig:Mars}).
+in X-ray diffraction data from Mars {\changed (Fig ~\ref{fig:Mars}).
 It is unknown if/how the X-ray source (e.g.
 Cu versus Co) fundamentally changes the peak shape characteristics of the amorphous content in this study.
-However, qualitative assessment suggests similar overall peak shapes between terrestrial samples analysed with a Cu X-ray anode (Fig.~\ref{fig:terrestrial}) and those analysed with a Co X-ray anode (Fig.~\ref{fig:Mars}).
-In particular, the Buckskin 2 analysis of the amorphous component indicates a greater positive skewness than the Gobabeb 2 sample.
-In addition, the nongaussianity of the Buckskin 2 sample is $\sim 3\times$ greater compared to the Gobabeb 2 amorphous component.
+However, qualitative assessment suggests similar overall peak shapes between terrestrial samples analysed with a Cu X-ray anode (Fig.~\ref{fig:terrestrial}) and those analysed with a Co X-ray anode (Fig.~\ref{fig:Mars})}.
+In particular, the Buckskin 2 analysis of the amorphous component indicates a greater positive skewness {\changed than the Gobabeb 2 sample.
+In addition, the nongaussianity of the Buckskin 2 sample is $\sim 3\times$ greater compared to the Gobabeb 2 amorphous component}.
 These results suggest a change in the
-composition of the amorphous material between the two Martian samples. This interpretation, is consistent with prior
+composition of the amorphous material {\changed between the two Martian samples}. This interpretation, is consistent with prior
 published results which suggest that at the Gobabeb locality (Namib Dune, Bagnold Dune Field),
@@
 
-While results may imply a qualitative relationship to composition, further testing and an understanding of the program's limitations is necessary to better quantify compositional correlations.
-In particular, the AMORPH program can quantify the peak shape and proportion of amorphous material but it cannot deconvolve the diffraction pattern of multiple amorphous phases, nor can it distinguish between non-crystalline materials and X-ray amorphous phases such as nanophase oxides \citep{blake2013}.
+{\changed While results may imply a qualitative relationship to composition, further testing and an understanding of the program's limitations is necessary to better quantify compositional correlations.
+In particular, the AMORPH program can quantify the peak shape and proportion of amorphous material but it cannot deconvolve the diffraction pattern of multiple amorphous phases, nor can it distinguish between non-crystalline materials and X-ray amorphous phases such as nanophase oxides \citep{blake2013}.}
 
@@
 This project made use of the resources of the Centre for
-eResearch at the University of Auckland.
+eResearch at the University of Auckland. {\changed We would like to thank E. Rampe and two anonymous reviewers for their comments.}
 
@@
 
+{\changed 
 \bibitem[Andrade et al.(2017)]{andrade2017}
@@
 Journal of Volcanology and Geothermal Research, doi: 10.1016/j.jvolgeores.2017.08.008.
+}
 
@@
 
+{\changed 
 \bibitem[Blake et al.(2012)]{blake2012}
 Blake, D. et al. 2012. Characterization and calibration of the CheMin mineralogical instrument on Mars Science Laboratory. Space Science Reviews, 170, 341-399. doi: 10.1007/s11214-012-9905-1.
+}
 
@@
 \newblock \emph{Journal of Statistical Software, accepted. arxiv: 1606.03757}.
+
+{\changed 
+\bibitem[Brewer et al.(2013)]{starfield}
+Brewer, Brendon J., Daniel Foreman-Mackey, and David W. Hogg. ``Probabilistic catalogs for crowded stellar fields.'' The Astronomical Journal 146, no. 1 (2013): 7.
+}
 
@@
 
+{\changed 
+\bibitem[Clark(1980)]{clean} Clark, B. G.
+``An efficient implementation of the algorithm `CLEAN'.''
+Astronomy and Astrophysics 89 (1980): 377.
+}
+
 \bibitem[Dehouck et al.(2014)]{dehouck2014}
@@
 
+{\changed 
+\bibitem[Harkness and Green(2000)]{object} Harkness, M.~A. and Green, P.~J., 2000,
+Parallel chains, delayed rejection and reversible jump MCMC for object recognition,
+British Machine Vision Conference Proceedings.
+}
+
+{\changed 
+\bibitem[Hobson \& McLachlan(2003)]{hobson} Hobson, M.~P., \& McLachlan, C.\ 2003, \mnras, 338, 765 
+}
+
 \bibitem[Kanakiya et al.(2017)]{kanakiya2017}
@@
 \newblock Arnold, 496pp.
+
+{\changed 
+\bibitem[Padayachee et al.(1999)]{padayachee}
+Padayachee, J., Prozesky, V., Von Der Linden, W., Nkwinika, M.
+S., \& Dose, V.
+(1999).
+Bayesian PIXE background subtraction.
+Nuclear Instruments and Methods in Physics Research Section B: Beam Interactions with Materials and Atoms, 150(1-4), 129-135.
+}
 
@@
 
+{\changed 
 \bibitem[Skilling(2006)]{skilling2006nested}
@@
 Bayesian analysis, 1, 4, 833--859.
+
+\bibitem[Skilling(1998)]{massinf}
+Skilling J., 1998, Massive Inference and Maximum Entropy, in Maximum Entropy
+and Bayesian Methods, Kluwer Academic Publishers, Dordrecht/Boston/London p.14
+}
 
@@
 
+{\changed 
+\bibitem[Worpel and Schwope(2015)]{worpel}
+Worpel, H., \& Schwope, A. D. (2015). Background subtraction and transient timing with Bayesian Blocks. Astronomy \& Astrophysics, 578, A80.
+}
+
 \bibitem[Zorn et al.(2018)]{zorn2018}
 Zorn, E.U., Rowe, M.C., Cronin, S.J., Ryan, A.G., Kennedy, L., Russell, J.K.
-2018.
+{\changed 2018}.
 Influence of porosity and groundmass crystallinity on dome rock strength; a case study from Mt.
@@
 For simplicity, text files for processing should be located in the same folder as the {\tt AMORPH.exe} program.
-The program is set to run until 10,000 saved parameter sets have been generated.
+
+{\changed 
+For speed, you can pass the command line options {\tt -o OPTIONS\_AGGRESSIVE} and
+{\tt -t 4} to {\tt AMORPH.exe}. The first of these options sets aggressive
+DNest4 parameters, and the second allows the program to use four threads
+simultaneously (you can use {\tt -t 8} if your computer allows 8 threads to run
+at full speed). Without the {\tt -o OPTIONS\_AGGRESSIVE} option, AMORPH uses
+the more conservative DNest4 settings specified in the file OPTIONS.
+}
+
+The program is set to run until 10,000 saved parameter sets
+{\changed (`particles')} have been generated.
 Data outputs may be viewed at any point, however, closing the program before reaching 10,000 will reduce the accuracy of the final calculations.

### 2018-04-17 doc/ms.tex
Resubmitted version (about to be anyway)

@@
 irrelevant when predicting what data would be observed. In the AMORPH model,
-we introduced hyperparameters describing the typical amplitude
-and width of the gaussians for the amorphous component, and
-the degree to which the actual amplitudes and widths are scattered
+we introduced hyperparameters describing the typical {\changed integrated area (`mass')}
+and width of the peaks for the amorphous component, and
+the degree to which the actual {\changed masses} and widths are scattered
 around that typical value. This was also done for the crystalline
@@
 \caption{The PGM. The hyperparameters controlling the prior distributions
-for the amorphous and crystalline gaussians are denoted collectively
+for the amorphous and crystalline peaks are denoted collectively
 by $\boldsymbol{\alpha}_{\rm amorph}$ and
@@
                             Uniform$(x_{\rm min}, x_{\rm max})$ \\
-$A_{i, \rm xtal}$ & {\changed Masses} of narrow peaks &
+$M_{i, \rm xtal}$ & {\changed Masses} of narrow peaks &
 {\changed
@@
 materials (CheMin instrument data\footnote{\tt http://pds-geosciences.wustl.edu/msl/msl-m-chemin-4-rdr-v1/mslcmn{\textunderscore}1xxx.data.rdr4/}).
-CheMin diffraction patterns have been processed using the original Co K$\alpha$ ($1.7890 \dot{A}$) data as well as data transformed from a Co K$\alpha$ to Cu K$\alpha$ ($1.5406 \dot{A}$) X-ray source using Bragg's law.
+CheMin diffraction patterns have been processed using the original Co K$\alpha$ ($1.7890$ \AA) data as well as data transformed from a Co K$\alpha$ to Cu K$\alpha$ ($1.5406$ \AA) X-ray source using Bragg's law.
 This recalculation simply allows for an easier visual comparison to terrestrial samples analysed with the more standard Cu K$\alpha$ X-ray source (see below).
-Note that for Co K$\alpha$ results, background critical points have been adjusted to 11.7--46.9\degree~$2\theta$ to maintain the linear background over the amorphous region from $\sim$ 8.8 to 2.25 $\dot{A}$ (Table ~\ref{tab:anode}).
+Note that for Co K$\alpha$ results, background critical points have been adjusted to 11.7--46.9\degree~$2\theta$ to maintain the linear background over the amorphous region from $\sim$ 8.8 to 2.25 \AA (Table ~\ref{tab:anode}).
 Although the results are uncalibrated, measured crystallinites provide a good first approximation of actual crystallinity, and relative differences are robust.
